### 2020-04-30 manuscript.tex
Add shell of paper

\documentclass[3p,authoryear]{elsarticle}
\journal{Annals of Nuclear Energy}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[charter]{mathdesign} % Type-1 Latin Modern modern
\usepackage[T1]{fontenc}         % Use T1 encoding instead of OT1
\usepackage[utf8]{inputenc}      % Use UTF8 input encoding
\usepackage{microtype}           % Improve typography
\usepackage{booktabs}            % Publication quality tables
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{float}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{siunitx}
\usepackage{multirow}

\sisetup{list-final-separator = {, and }}

% Captions for figures and tables
\usepackage[labelfont=bf]{caption}
\captionsetup[figure]{labelsep=period, name=Fig.}
\captionsetup[table]{labelsep=newline}

\usepackage{color}
\definecolor{aneblue}{RGB}{0,128,173}
\usepackage[colorlinks,breaklinks,bookmarksnumbered,bookmarksopen]{hyperref}
\AtBeginDocument{
  \hypersetup{linkcolor=aneblue, citecolor=aneblue, urlcolor=aneblue}
}

\usepackage[capitalise,nameinlink]{cleveref}

\newtheorem{lemma}{Lemma}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\title{Depletion capabilities in the OpenMC Monte Carlo particle transport code}

\author[anl]{Paul K. Romano\corref{cor1}}
\ead{promano@anl.gov}
\cortext[cor1]{Corresponding author. Tel.: +1 630 252 6779.}

\author[lanl]{Colin Josey}
\ead{cjosey@lanl.gov}

\author[gatech]{Andrew E. Johnson}
\ead{dasindrew@gatech.edu}

\author[tsinghua]{Jingang Liang}
\ead{jingang@tsinghua.edu.cn}

\address[anl]{Argonne National Laboratory, 9700 S. Cass Ave, Lemont, IL 60439, United States}
\address[lanl]{Los Alamos National Laboratory, PO Box 1663, Los Alamos, NM 87545, United States}
\address[gatech]{Georgia Institute of Technology, 770 State St NW, Atlanta, GA 30318, United States}
\address[tsinghua]{Institute of Nuclear and New Energy Technology, Tsinghua University, Beijing, China}

\begin{abstract}

\end{abstract}

\begin{keyword}
  Monte Carlo, depletion
\end{keyword}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Mention isotalo papers

Mention previous depletion coupling with OpenMC (IMPC-Depletion, OpenBU, AMAC,
etc.)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methodology}
\label{sec:methods}

\subsection{Time integration}

Josey's work and comparison to substeps?

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\label{sec:results}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgments}

This research was supported by the Exascale Computing Project (17-SC-20-SC), a
collaborative effort of the U.S. Department of Energy Office of Science and the
National Nuclear Security Administration. The submitted manuscript has been
created by UChicago Argonne, LLC, operator of Argonne National Laboratory under
contract DE-AC02-06CH11357.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{References}

\bibliographystyle{elsarticle-harv}
\bibliography{references}

\clearpage
\vspace*{\fill}
\noindent\fbox{%
  \parbox{\textwidth}{%
    The submitted manuscript has been created by UChicago Argonne, LLC, Operator
    of Argonne \mbox{National} Laboratory (``Argonne'').  Argonne, a
    U.S. Department of Energy Office of Science laboratory, is operated
    under Contract No. \mbox{DE-AC02-06CH11357}.  The U.S. Government retains for
    itself, and others acting on its behalf, a paid-up nonexclusive, irrevocable
    worldwide license in said article to reproduce, prepare derivative works,
    distribute copies to the public, and perform publicly and display publicly,
    by or on behalf of the Government. The Department of Energy will provide
    public access to these results of federally sponsored research in accordance
    with the DOE Public Access
    Plan. \url{http://energy.gov/downloads/doe-public-access-plan.}
  }%
}
\vspace*{\fill}

\end{document}

### 2020-04-30 manuscript.tex
Add references submodule, material from OpenMC documentation

@@
 
+When materials in a system are subject to irradiation over a long period of
+time, nuclides within the material will transmute due to nuclear reactions as
+well as spontaneous radioactive decay. The time-dependent process by which
+nuclides transmute under irradiation is known as \emph{depletion} or
+\emph{burnup}. To accurately analyze nuclear systems, it is often necessary to
+predict how the composition of materials will change since this change results
+in a corresponding change in the solution of the transport equation. The
+equation that governs the transmutation and decay of nuclides inside of an
+irradiated environment can be written as
+\begin{equation}
+  \frac{dN_i(t)}{dt} = \sum\limits_j
+  \underbrace{\left [ \underbrace{f_{j \rightarrow i} \int_0^\infty dE \;
+  \sigma_j (E, t) \phi(E,t)}_\text{transmutation} +
+  \underbrace{\lambda_{j\rightarrow i}}_\text{decay} \right ]
+  N_j(t)}_{\text{Production of nuclide }i\text{ from nuclide }j}
+  - \underbrace{\left [\underbrace{\int_0^\infty dE \; \sigma_i
+  (E,t) \phi(E,t)}_\text{transmutation} +
+  \underbrace{\sum\limits_j \lambda_{i\rightarrow j}}_\text{decay} \right ]
+  N_i(t)}_{\text{Loss of nuclide }i}
+\end{equation}
+where $N_i$ is the density of nuclide $i$ at time $t$, $\sigma_i$ is the
+transmutation cross section for nuclide $i$ at energy $E$, $f_{j \rightarrow i}$
+is the fraction of transmutation reactions in nuclide $j$ that produce nuclide
+$i$, and $\lambda_{j \rightarrow i}$ is the decay constant for decay modes in
+nuclide $j$ that produce nuclide $i$. Note that we have not included the spatial
+dependence of the flux or cross sections. As one can see, the equation simply
+states that the rate of change of $N_i$ is equal to the production rate minus
+the loss rate. Because the equation for nuclide $i$ depends on the nuclide
+density for possibly many other nuclides, we have a system of first-order
+differential equations. To form a proper initial value problem, we also need the
+nuclide densities at time 0:
+\begin{equation}
+    N_i(0) = N_{i,0}.
+\end{equation}
+These equations can be written more compactly in matrix notation as
+\begin{equation}
+  \label{eq:depletion-matrix}
+  \frac{d\mathbf{n}}{dt} = \mathbf{A}(\mathbf{n},t)\mathbf{n}, \quad \mathbf{n}(0) =
+  \mathbf{n}_0
+\end{equation}
+where $\mathbf{n} \in \mathbb{R}^n$ is the nuclide density vector,
+$\mathbf{A}(\mathbf{n},t) \in \mathbb{R}^{n\times n}$ is the burnup matrix
+containing the decay and transmutation coefficients, and $\mathbf{n}_0$ is the
+initial density vector. Note that the burnup matrix depends on $\mathbf{n}$
+because the solution to the transport equation depends on the nuclide densities.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsection{Time integration}
+\label{sec:time_integration}
 
-Josey's work and comparison to substeps?
+A variety of numerical methods exist for solving \cref{eq:depletion-matrix}. The
+simplest such method, known as the "predictor" method, is to divide the overall
+time interval of interest $[0,t]$ into smaller timesteps over which it is
+assumed that the burnup matrix is constant. Let $t \in [t_i, t_i + h]$ be one
+such timestep. Over the timestep, the solution to \cref{eq:depletion-matrix} can
+be written analytically using the matrix exponential
+\begin{equation}
+    \mathbf{A}_i = \mathbf{A}(\mathbf{n}_i, t_i) \\
+    \mathbf{n}_{i+1} = e^{\mathbf{A}_i h} \mathbf{n}_i
+\end{equation}
+where $\mathbf{n}_i \equiv \mathbf{n}(t_i)$. The exponential of a matrix
+$\mathbf{X}$ is defined by the power series expansion
+\begin{equation}
+    e^{\mathbf{X}} = \sum\limits_{k=0}^\infty \frac{1}{k!} \left ( \mathbf{X}
+    \right )^k
+\end{equation}
+where $\mathbf{X}^0 = \mathbf{I}$. A series of so-called predictor-corrector
+methods that use multiple stages offer improved accuracy over the predictor
+method. The simplest of these methods, the CE/CM algorithm, is defined as
+\begin{equation}
+    \mathbf{n}_{i+1/2} = e^{\frac{h}{2}\mathbf{A}(\mathbf{n}_i, t_i)} \mathbf{n}_i \\
+    \mathbf{n}_{i+1} = e^{h \mathbf{A}(\mathbf{n}_{i+1/2},t_{i+1/2})} \mathbf{n}_i
+\end{equation}
+Here, the value of $\mathbf{n}$ at the midpoint is estimated using
+$\mathbf{A}$ evaluated at the beginning of the timestep. Then,
+$\mathbf{A}$ is evaluated using the densities at the midpoint and used to
+integrate over the entire timestep.
+
+Our aim here is not to exhaustively describe all integration methods but rather
+to give a few examples that elucidate the main considerations one must take into
+account when choosing a method. Generally, there is a tradeoff between the
+accuracy of the method and its computational expense. The expense is driven
+almost entirely by the time to compute a transport solution, i.e., to evaluate
+$\mathbf{A}$ for a given $\mathbf{n}$. Thus, the cost of a method scales with
+the number of $\mathbf{A}$ evaluations that are performed per timestep. On the
+other hand, methods that require more evaluations generally achieve higher
+accuracy. The predictor method only requires one evaluation and its error
+converges as $\mathcal{O}(h)$. The CE/CM method requires two evaluations and is
+thus twice as expensive as the predictor method, but achieves an error of
+$\mathcal{O}(h^2)$. An exhaustive description of time integration methods and
+their merits can be found in the thesis of \citet{josey2017phd}.
+
+OpenMC does not rely on a single time integration method but rather has several
+classes that implement different algorithms. For example, the
+openmc.deplete.PredictorIntegrator class implements the predictor
+method, and the openmc.deplete.CECMIntegrator class implements the
+CE/CM method. A full list of the integrator classes available can be found in
+the documentation for the openmc.deplete module.
+
+\subsection{Matrix Exponential}
+
+As we saw in \cref{sec:time_integration}, numerically integrating
+\cref{eq:depletion-matrix} requires evaluating one or more matrix exponentials.
+OpenMC uses the Chebyshev rational approximation method (CRAM), which was
+introduced in a series of papers by Pusa~\citep{pusa2010nse,pusa2011nse}, to
+evaluate matrix exponentials. In particular, OpenMC utilizes an incomplete
+partial fraction (IPF) form~\citep{pusa2016nse} of CRAM that provides a good
+balance of numerical stability and efficiency. In this representation the matrix
+exponential is approximated as
+\begin{equation}
+    e^{\mathbf{A}t} \approx \alpha_0 \prod\limits_{\ell=1}^{k/2} \left (
+    \mathbf{I} + 2 \text{Re} \left ( \widetilde{\alpha}_\ell \left (\mathbf{A}t
+    - \theta_\ell \mathbf{I} \right )^{-1} \right ) \right )
+\end{equation}
+where $k$ is the order of the approximation and $\alpha_0$,
+$\widetilde{\alpha}_\ell$, and $\theta_\ell$ are coefficients that have been
+tabulated for orders up to $k=48$. Rather than computing the full approximation
+and then multiplying it by a vector, the following algorithm\footnote{The
+original description of the algorithm presented by \citet{pusa2016nse} contains
+a typo.} is used to incrementally apply the terms within the product:
+\begin{enumerate}
+\item $\mathbf{n} \gets \mathbf{n_0}$
+\item For $\ell = 1, 2, \dots, k/2$
+  \begin{itemize}
+   \item $\mathbf{n} \gets \mathbf{n} + 2\text{Re}(\widetilde{\alpha}_\ell
+     (\mathbf{A}t - \theta_\ell)^{-1})\mathbf{n}$
+  \end{itemize}
+\item $\mathbf{n} \gets \alpha_0 \mathbf{n}$
+\end{enumerate}
+The $k$\ th order approximation for CRAM requires solving $k/2$ sparse linear
+systems. OpenMC relies on functionality from scipy.sparse.linalg for solving the
+linear systems.
 

### 2020-05-01 manuscript.tex
Cite previous works coupling OpenMC with depletion
Change IPF CRAM outline to use algorithm environment

@@
 
-Mention isotalo papers
+Mention Isotalo papers
 
-Mention previous depletion coupling with OpenMC (IMPC-Depletion, OpenBU, AMAC,
-etc.)
+Lots of community efforts toward coupled transport-depletion calculations with
+OpenMC~\citep{gul2017ane,lanversin2017icone,lanversin2019phd,liu2019nst,zhuang2020pne,zhao2020ned,zhang2020ane}.
 
@@
 a typo.} is used to incrementally apply the terms within the product:
-\begin{enumerate}
-\item $\mathbf{n} \gets \mathbf{n_0}$
-\item For $\ell = 1, 2, \dots, k/2$
-  \begin{itemize}
-   \item $\mathbf{n} \gets \mathbf{n} + 2\text{Re}(\widetilde{\alpha}_\ell
+\begin{algorithm}[H]
+  \caption{Incomplete partial fraction form of CRAM.}
+  \label{alg:cram}
+  \begin{algorithmic}[1]
+    \State $\mathbf{n} \gets \mathbf{n_0}$
+    \For{$\ell \gets 1$ \textbf{to} $k/2$}
+      \State $\mathbf{n} \gets \mathbf{n} + 2\text{Re}(\widetilde{\alpha}_\ell
         (\mathbf{A}t - \theta_\ell)^{-1})\mathbf{n}$
-  \end{itemize}
-\item $\mathbf{n} \gets \alpha_0 \mathbf{n}$
-\end{enumerate}
+    \EndFor
+    \State $\mathbf{n} \gets \alpha_0 \mathbf{n}$
+  \end{algorithmic}
+\end{algorithm}
 The $k$\ th order approximation for CRAM requires solving $k/2$ sparse linear

### 2020-05-03 manuscript.tex
One more reference

@@
 Lots of community efforts toward coupled transport-depletion calculations with
-OpenMC~\citep{gul2017ane,lanversin2017icone,lanversin2019phd,liu2019nst,zhuang2020pne,zhao2020ned,zhang2020ane}.
+OpenMC~\citep{gul2017ane,lanversin2017icone,lanversin2019phd,liu2019nst,zhuang2020pne,zhao2020ned,zhao2020cpc,zhang2020ane}.
 

### 2020-05-08 manuscript.tex
Work on introduction

--- 2020-05-01 manuscript.tex
@@
 
-Mention Isotalo papers
+When a material in a system is subject to irradiation over a long period of
+time, nuclides within the material will transmute due to nuclear reactions and
+spontaneous radioactive decay. The time-dependent process by which nuclides
+transmute under irradiation is known as \emph{depletion}, \emph{burnup}, or
+\emph{transmutation}. Modeling this process plays an important role in the
+design and licensing of nuclear reactors~\citep{betzler2019ned}. Consequently,
+many codes have been developed that solve the underlying equations for the
+change in material compositions over time.
 
-Lots of community efforts toward coupled transport-depletion calculations with
-OpenMC~\citep{gul2017ane,lanversin2017icone,lanversin2019phd,liu2019nst,zhuang2020pne,zhao2020ned,zhang2020ane}.
+When modeling nuclear reactors, depletion is inherently linked with solving the
+neutron transport equation. Nuclear reaction rates, which determine the rate at
+which nuclides are transmuted, must be provided from a transport code.
+Conversely, the change in material composition directly affects the solution of
+the transport equation. As will be shown in \cref{sec:methods}, this results in
+a non-linear coupling between transport and depletion. The implication is that
+depletion codes are rarely used in isolation---normally, they are directly
+coupled with a neutron transport code.
 
-When materials in a system are subject to irradiation over a long period of
-time, nuclides within the material will transmute due to nuclear reactions as
-well as spontaneous radioactive decay. The time-dependent process by which
-nuclides transmute under irradiation is known as \emph{depletion} or
-\emph{burnup}. To accurately analyze nuclear systems, it is often necessary to
-predict how the composition of materials will change since this change results
-in a corresponding change in the solution of the transport equation. The
-equation that governs the transmutation and decay of nuclides inside of an
+With continued advances in computing, Monte Carlo (MC) particle transport
+simulations are now widely used in reactor design and analysis in a variety of
+roles. The earliest efforts to couple MC particle transport with depletion date
+back to the 1990s~\citep{moore1995inel,trellue1998lanl}. These early efforts
+involved taking existing MC and depletion codes, such as
+MCNP~\citep{goorley2012nt} and ORIGEN~\citep{croff1983nt}, and putting a
+``wrapper'' around them that orchestrated the transfer of solutions from one
+code to the other. Thus, the actual transport/depletion codes themselves
+required no source code changes. In the last decade, many MC code developers
+have chosen to directly include a depletion solver as a first-rate feature,
+including MCNPX~\citep{waters2007aip} (which has since been merged into
+MCNP6~\citep{goorley2012nt}), Serpent~\citep{leppanen2015ane}, and
+MC21~\citep{griesheimer2015ane}.
+
+In the present work, we describe a new depletion solver in
+OpenMC~\citep{romano2015ane1}, an open source, community-developed MC particle
+transport code. To date, OpenMC has not had a built-in depletion solver. This
+has led to numerous efforts in the community toward developing coupled
+transport-depletion calculations with
+OpenMC~\citep{gul2017ane,lanversin2017icone,lanversin2019phd,liu2019nst,zhuang2020pne,zhao2020ned,zhao2020cpc,zhang2020ane}.
+While these efforts have proven the viability of performing such coupled
+calculations with OpenMC, none of them has resulted in a validated,
+openly-available, and generalized capability that OpenMC users can take
+advantage of. The capability that is described herein is part of the OpenMC
+project and can be found in version 0.11 of the code or later.
+
+The paper is organized as follows. In \cref{sec:methods}, we describe the
+methods and algorithms used to solve the depletion equations, the implementation
+within OpenMC, and how nuclear data is used to create depletion chains. Then, in
+\cref{sec:results}, we present results on two problems that demonstrate the
+accuracy of OpenMC when compared to Serpent~\citep{leppanen2015ane}. In
+\cref{sec:conclusions}, we summarize the findings, discuss limitations, and
+provide commentary on future avenues of inquiry that may be promising.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Methodology}
+\label{sec:methods}
+
+The equation that governs the transmutation and decay of nuclides inside of an
 irradiated environment can be written as
@@
 where $N_i$ is the density of nuclide $i$ at time $t$, $\sigma_i$ is the
-transmutation cross section for nuclide $i$ at energy $E$, $f_{j \rightarrow i}$
-is the fraction of transmutation reactions in nuclide $j$ that produce nuclide
-$i$, and $\lambda_{j \rightarrow i}$ is the decay constant for decay modes in
-nuclide $j$ that produce nuclide $i$. Note that we have not included the spatial
-dependence of the flux or cross sections. As one can see, the equation simply
-states that the rate of change of $N_i$ is equal to the production rate minus
-the loss rate. Because the equation for nuclide $i$ depends on the nuclide
-density for possibly many other nuclides, we have a system of first-order
-differential equations. To form a proper initial value problem, we also need the
-nuclide densities at time 0:
+transmutation cross section for nuclide $i$ at energy $E$ and time $t$, $\phi$
+is the neutron flux at energy $E$ and time $t$, $f_{j \rightarrow i}$ is the
+fraction of transmutation reactions in nuclide $j$ that produce nuclide $i$, and
+$\lambda_{j \rightarrow i}$ is the decay constant for decay modes in nuclide $j$
+that produce nuclide $i$. Note that we have not included the spatial dependence
+of the flux or cross sections. As one can see, the equation simply states that
+the rate of change of $N_i$ is equal to the production rate minus the loss rate.
+Because the equation for nuclide $i$ depends on the nuclide density for possibly
+many other nuclides, we have a system of first-order differential equations. To
+form a proper initial value problem, we also need the nuclide densities at time
+0:
 \begin{equation}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Methodology}
-\label{sec:methods}
-
 \subsection{Time integration}
 \label{sec:time_integration}
+
+Mention Isotalo papers
 
@@
 \section{Conclusions}
+\label{sec:conclusions}
 

### 2020-05-09 manuscript.tex
Start adjusting methodology

@@
 \address[anl]{Argonne National Laboratory, 9700 S. Cass Ave, Lemont, IL 60439, United States}
-\address[lanl]{Los Alamos National Laboratory, PO Box 1663, Los Alamos, NM 87545, United States}
+\address[lanl]{Los Alamos National Laboratory, P.O. Box 1663, Los Alamos, NM 87545, United States}
 \address[gatech]{Georgia Institute of Technology, 770 State St NW, Atlanta, GA 30318, United States}
@@
 has led to numerous efforts in the community toward developing coupled
-transport-depletion calculations with
-OpenMC~\citep{gul2017ane,lanversin2017icone,lanversin2019phd,liu2019nst,zhuang2020pne,zhao2020ned,zhao2020cpc,zhang2020ane}.
-While these efforts have proven the viability of performing such coupled
-calculations with OpenMC, none of them has resulted in a validated,
-openly-available, and generalized capability that OpenMC users can take
-advantage of. The capability that is described herein is part of the OpenMC
+transport-depletion calculations with OpenMC~\citep{gul2017ane,
+lanversin2017icone, lanversin2019phd, liu2019nst, zhuang2020pne, zhao2020ned,
+zhao2020cpc, zhang2020ane}. While these efforts have proven the viability of
+performing such coupled calculations with OpenMC, none of them has resulted in a
+validated, openly-available, and generalized capability that OpenMC users can
+take advantage of. The capability that is described herein is part of the OpenMC
 project and can be found in version 0.11 of the code or later.
@@
 \begin{equation}
-  \frac{dN_i(t)}{dt} = \sum\limits_j
-  \underbrace{\left [ \underbrace{f_{j \rightarrow i} \int_0^\infty dE \;
-  \sigma_j (E, t) \phi(E,t)}_\text{transmutation} +
-  \underbrace{\lambda_{j\rightarrow i}}_\text{decay} \right ]
-  N_j(t)}_{\text{Production of nuclide }i\text{ from nuclide }j}
-  - \underbrace{\left [\underbrace{\int_0^\infty dE \; \sigma_i
-  (E,t) \phi(E,t)}_\text{transmutation} +
-  \underbrace{\sum\limits_j \lambda_{i\rightarrow j}}_\text{decay} \right ]
-  N_i(t)}_{\text{Loss of nuclide }i}
+  \label{eq:depletion}
+  \begin{split}
+  \frac{dN_i(t)}{dt} = &\sum\limits_j \left [ f_{j \rightarrow i}
+  \int_0^\infty dE \; \sigma_j (E, t) \phi(E,t) + \lambda_{j\rightarrow i}
+  \right ] N_j(t) \\ &- \left [\int_0^\infty dE \; \sigma_i (E,t) \phi(E,t) +
+  \sum\limits_j \lambda_{i\rightarrow j} \right ] N_i(t), \quad i=1,\dots,n
+  \end{split}
 \end{equation}
-where $N_i$ is the density of nuclide $i$ at time $t$, $\sigma_i$ is the
-transmutation cross section for nuclide $i$ at energy $E$ and time $t$, $\phi$
-is the neutron flux at energy $E$ and time $t$, $f_{j \rightarrow i}$ is the
-fraction of transmutation reactions in nuclide $j$ that produce nuclide $i$, and
-$\lambda_{j \rightarrow i}$ is the decay constant for decay modes in nuclide $j$
-that produce nuclide $i$. Note that we have not included the spatial dependence
-of the flux or cross sections. As one can see, the equation simply states that
-the rate of change of $N_i$ is equal to the production rate minus the loss rate.
-Because the equation for nuclide $i$ depends on the nuclide density for possibly
-many other nuclides, we have a system of first-order differential equations. To
-form a proper initial value problem, we also need the nuclide densities at time
-0:
+where
+\begin{equation*}
+  \begin{split}
+  N_i(t) &\equiv \text{density of nuclide $i$ at time $t$} \\
+  \sigma_i(E,t) &\equiv \text{transmutation cross section for nuclide $i$ at energy $E$ and time $t$} \\
+  \phi(E,t) &\equiv \text{neutron flux at energy $E$ and time $t$} \\
+  f_{j \rightarrow i} &\equiv \text{fraction of transmutation reactions in nuclide $j$ that produce nuclide $i$} \\
+  \lambda_{j \rightarrow i} &\equiv \text{decay constant for decay modes in nuclide $j$ that produce nuclide $i$} \\
+  n &\equiv \text{total number of nuclides.}
+  \end{split}
+\end{equation*}
+Note that we have not included the spatial dependence of the flux or cross
+sections. For \cref{eq:depletion} to be accurate, the spatial variation of the
+reaction rates over the region which is being depleted should be small. For a
+reactor model, this may entail using a unique material in each fuel pin that is
+depleted separately. Ultimately, the user is responsible for deciding what level
+of discretization is necessary for the problem at hand.
+
+\Cref{eq:depletion} simply states that the rate of change of $N_i$ is equal to
+the production rate minus the loss rate. Because the equation for $N_i$ depends
+on the density other nuclides, $\{N_j \,|\, j \ne i\}$, \cref{eq:depletion}
+represents a system of first-order differential equations. To form a proper
+initial value problem, we also need the nuclide densities at time 0:
 \begin{equation}
@@
 \end{equation}
-where $\mathbf{n} \in \mathbb{R}^n$ is the nuclide density vector,
-$\mathbf{A}(\mathbf{n},t) \in \mathbb{R}^{n\times n}$ is the burnup matrix
-containing the decay and transmutation coefficients, and $\mathbf{n}_0$ is the
-initial density vector. Note that the burnup matrix depends on $\mathbf{n}$
-because the solution to the transport equation depends on the nuclide densities.
+where
+\begin{equation}
+  \mathbf{n} = \begin{pmatrix} N_1 \\ N_2 \\ \vdots \\ N_n \end{pmatrix}, \quad
+  \mathbf{n}_0 = \begin{pmatrix} N_{1,0} \\ N_{2,0} \\ \vdots \\ N_{n,0} \end{pmatrix}
+\end{equation}
+and $\mathbf{A}(\mathbf{n},t) \in \mathbb{R}^{n\times n}$ is the burnup matrix
+containing the decay and transmutation coefficients. Note that the burnup matrix
+depends on $\mathbf{n}$ because the solution to the transport equation depends
+on the nuclide densities.
 

### 2020-05-19 manuscript.tex
Add reference for OpenMC 0.11
Work on methodology section

@@
 take advantage of. The capability that is described herein is part of the OpenMC
-project and can be found in version 0.11 of the code or later.
+project and can be found in version 0.11~\citep{openmc-0110} of the code or
+later.
 
@@
 depleted separately. Ultimately, the user is responsible for deciding what level
-of discretization is necessary for the problem at hand.
+of discretization is adequate for the problem at hand.
 
@@
 the production rate minus the loss rate. Because the equation for $N_i$ depends
-on the density other nuclides, $\{N_j \,|\, j \ne i\}$, \cref{eq:depletion}
+on the density of other nuclides, $\{N_j \,|\, j \ne i\}$, \cref{eq:depletion}
 represents a system of first-order differential equations. To form a proper
-initial value problem, we also need the nuclide densities at time 0:
+initial value problem, the nuclide densities at time 0 must be specified:
 \begin{equation}
@@
 \begin{equation}
-  \label{eq:depletion-matrix}
+  \label{eq:depletion-matrix-t}
   \frac{d\mathbf{n}}{dt} = \mathbf{A}(\mathbf{n},t)\mathbf{n}, \quad \mathbf{n}(0) =
@@
 depends on $\mathbf{n}$ because the solution to the transport equation depends
-on the nuclide densities.
+on the nuclide densities, thereby making \cref{eq:depletion-matrix-t} nonlinear.
 
-\subsection{Time integration}
+Although the neutron transport equation is time-dependent, the timescale over
+which material compositions change is sufficiently long such that the transport
+equation can be solved as a time-independent equation. Without loss of
+generality, let us write the burnup matrix solely as a function of $\mathbf{n}$
+so that \cref{eq:depletion-matrix-t} becomes
+\begin{equation}
+  \label{eq:depletion-matrix}
+  \frac{d\mathbf{n}}{dt} = \mathbf{A}(\mathbf{n})\mathbf{n}, \quad \mathbf{n}(0) =
+  \mathbf{n}_0.
+\end{equation}
+This better reflects the fact that the time, $t$, is generally not an input to
+the transport equation, and hence the burnup matrix is uniquely determined by
+the value of $\mathbf{n}$.
+
+The principle difficulty of solving \cref{eq:depletion-matrix} is the fact that
+the elements of $\mathbf{A}$ are not constant with respect to time. If
+$\mathbf{A}$ is constant, \cref{eq:depletion-matrix} admits a closed-form
+solution,
+\begin{equation}
+  \mathbf{n}(t) = \exp \left (\mathbf{A} t \right ) \mathbf{n}_0.
+\end{equation}
+The exponential of a matrix $\mathbf{X}$ is defined by the power series
+expansion
+\begin{equation}
+    \exp(\mathbf{X}) = \sum\limits_{k=0}^\infty \frac{1}{k!} \mathbf{X}^k
+\end{equation}
+where $\mathbf{X}^0 = \mathbf{I}$. Most numerical methods for solving
+\cref{eq:depletion-matrix} involve two separate components:
+\begin{enumerate}
+  \item Approximating \cref{eq:depletion-matrix} in a form that involves taking
+  one or more matrix exponentials, and
+  \item Evaluating the matrix exponential or, equivalently, the action of a
+  matrix exponential on a vector.
+\end{enumerate}
+We will henceforth refer to the first component as \emph{time integration}. Let
+us now consider how each of these components are treated in OpenMC.
+
+\subsection{Time Integration}
 \label{sec:time_integration}
-
-Mention Isotalo papers
 
@@
 be written analytically using the matrix exponential
-\begin{equation}
-    \mathbf{A}_i = \mathbf{A}(\mathbf{n}_i, t_i) \\
-    \mathbf{n}_{i+1} = e^{\mathbf{A}_i h} \mathbf{n}_i
-\end{equation}
-where $\mathbf{n}_i \equiv \mathbf{n}(t_i)$. The exponential of a matrix
-$\mathbf{X}$ is defined by the power series expansion
-\begin{equation}
-    e^{\mathbf{X}} = \sum\limits_{k=0}^\infty \frac{1}{k!} \left ( \mathbf{X}
-    \right )^k
-\end{equation}
-where $\mathbf{X}^0 = \mathbf{I}$. A series of so-called predictor-corrector
+\begin{align}
+    \mathbf{A}_i &= \mathbf{A}(\mathbf{n}_i) \\
+    \mathbf{n}_{i+1} &= \exp(\mathbf{A}_i h) \mathbf{n}_i
+\end{align}
+where $\mathbf{n}_i \equiv \mathbf{n}(t_i)$. A series of so-called predictor-corrector
 methods that use multiple stages offer improved accuracy over the predictor
 method. The simplest of these methods, the CE/CM algorithm, is defined as
-\begin{equation}
-    \mathbf{n}_{i+1/2} = e^{\frac{h}{2}\mathbf{A}(\mathbf{n}_i, t_i)} \mathbf{n}_i \\
-    \mathbf{n}_{i+1} = e^{h \mathbf{A}(\mathbf{n}_{i+1/2},t_{i+1/2})} \mathbf{n}_i
-\end{equation}
+\begin{align}
+    \mathbf{n}_{i+1/2} &= \exp \left (\frac{h}{2}\mathbf{A}(\mathbf{n}_i) \right) \mathbf{n}_i \\
+    \mathbf{n}_{i+1} &= \exp \left(h \mathbf{A}(\mathbf{n}_{i+1/2}) \right) \mathbf{n}_i
+\end{align}
 Here, the value of $\mathbf{n}$ at the midpoint is estimated using
@@
 \begin{equation}
-    e^{\mathbf{A}t} \approx \alpha_0 \prod\limits_{\ell=1}^{k/2} \left (
+    \exp(\mathbf{A}t) \approx \alpha_0 \prod\limits_{\ell=1}^{k/2} \left (
     \mathbf{I} + 2 \text{Re} \left ( \widetilde{\alpha}_\ell \left (\mathbf{A}t
@@
 \end{algorithm}
-The $k$\ th order approximation for CRAM requires solving $k/2$ sparse linear
+The $k$th order approximation for CRAM requires solving $k/2$ sparse linear
 systems. OpenMC relies on functionality from scipy.sparse.linalg for solving the
 linear systems.
+
+\subsection{Implementation}
+
+In-memory handling through Python API. This allows time loading nuclear data to
+be amortized since it only needs to be done once.
 
@@
 \label{sec:results}
+
+\subsection{Data Preparation}
+
+Using ENDF/B-VII.1 files from MCNP distribution. However, for light water
+$S(\alpha,\beta)$, Serpent is not capable of handling continuous format, so the
+file from ENDF/B-VII.0 was used. The evaluation didn't change between VII.0 and
+VII.1, but the processing was different. For Serpent, decay and neutron fission
+yield sublibraries are concatenated.
+
+\subsection{}
 

### 2020-05-19 manuscript.tex
Work on time integration section

@@
 \begin{equation}
+  \label{eq:constant-A}
   \mathbf{n}(t) = \exp \left (\mathbf{A} t \right ) \mathbf{n}_0.
@@
 
-A variety of numerical methods exist for solving \cref{eq:depletion-matrix}. The
-simplest such method, known as the "predictor" method, is to divide the overall
-time interval of interest $[0,t]$ into smaller timesteps over which it is
-assumed that the burnup matrix is constant. Let $t \in [t_i, t_i + h]$ be one
-such timestep. Over the timestep, the solution to \cref{eq:depletion-matrix} can
-be written analytically using the matrix exponential
-\begin{align}
-    \mathbf{A}_i &= \mathbf{A}(\mathbf{n}_i) \\
-    \mathbf{n}_{i+1} &= \exp(\mathbf{A}_i h) \mathbf{n}_i
-\end{align}
-where $\mathbf{n}_i \equiv \mathbf{n}(t_i)$. A series of so-called predictor-corrector
-methods that use multiple stages offer improved accuracy over the predictor
-method. The simplest of these methods, the CE/CM algorithm, is defined as
+Because an analytical solution to \cref{eq:depletion-matrix} is not available,
+one must resort to using a numerical method that generates approximation
+solutions in discrete increments. In general, the problem we need to solve then
+is: given $\mathbf{n}_i \equiv \mathbf{n}(t_i)$, what is the value of
+$\mathbf{n}_{i+1} \equiv \mathbf{n}(t_{i+1}) = \mathbf{n}(t_i + h)$, with $h$
+being the step size? Without loss of generality, we shall choose $t_i=0$ as it
+makes the presentation of various methods simpler.
+
+OpenMC doesn't rely on a single time integration method but instead has several
+methods available that the user can choose. In the following sections, we
+outline the available methods in OpenMC, several of which have not been
+described in the literature.
+
+\subsubsection{Predictor Method}
+
+The simplest numerical method for solving \cref{eq:depletion-matrix} over a
+timestep is known as the \emph{predictor} method. It proceeds by assuming that
+$\mathbf{A}$ is constant over the timestep and assumes the value at the
+beginning of the step. The solution then follows directly from
+\cref{eq:constant-A}:
+\begin{equation}
+    \mathbf{n}_{i+1} = \exp\left(h\mathbf{A}(\mathbf{n}_i) \right) \mathbf{n}_i.
+\end{equation}
+The end-of-step composition is thus ``predicted'' by the reaction rates at the
+beginning of the step. This method has the virtue of requiring a single
+transport solution and a single matrix exponential but is only first-order
+accurate.
+
+\subsubsection{Predictor-Corrector Methods}
+
+A series of \emph{predictor-corrector} methods that use multiple stages offer
+improved accuracy over the predictor method. The simplest of these methods is
+the CE/CM algorithm, which is the default algorithm used in
+MCNP6~\citep{fensin2006tans}. In this method, $\mathbf{A}$ is evaluated at the
+beginning of the step and used to deplete the composition over half the step.
+Then, $\mathbf{A}$ is reevaluated using the predicted middle-of-step composition
+and used to deplete over the full step. The algorithm can be expressed as
 \begin{align}
     \mathbf{n}_{i+1/2} &= \exp \left (\frac{h}{2}\mathbf{A}(\mathbf{n}_i) \right) \mathbf{n}_i \\
-    \mathbf{n}_{i+1} &= \exp \left(h \mathbf{A}(\mathbf{n}_{i+1/2}) \right) \mathbf{n}_i
+    \mathbf{n}_{i+1} &= \exp \left(h \mathbf{A}(\mathbf{n}_{i+1/2}) \right) \mathbf{n}_i.
 \end{align}
-Here, the value of $\mathbf{n}$ at the midpoint is estimated using
-$\mathbf{A}$ evaluated at the beginning of the timestep. Then,
-$\mathbf{A}$ is evaluated using the densities at the midpoint and used to
-integrate over the entire timestep.
+This method requires two transport solutions and two matrix exponential solves
+(twice as expensive as the predictor method) and is second-order accurate.
+
+More sophisticated predictor-corrector methods have been introduced by
+\citet{isotalo2011ane2,isotalo2011ane3}. In the CE/LI method, the end-of-step
+composition is first predicted using $\mathbf{A}(\mathbf{n}_i)$:
+\begin{equation}
+  \mathbf{n}_{i+1}^p = \exp \left( h\mathbf{A}(\mathbf{n}_i) \right) \mathbf{n}_i
+\end{equation}
+Then, $\mathbf{n}_{i+1}^p$, is used to reevaluate $\mathbf{A}$. An approximation
+of $\mathbf{A}$ that depends on time is formed by linearly interpolating between
+$\mathbf{A}(\mathbf{n}_i)$ and $\mathbf{A}\left(\mathbf{n}_{i+1}^p\right)$. This
+contitutes a new differential equation for an approximation of $\mathbf{n}$ that
+we shall denote $\hat{\mathbf{n}}$,
+\begin{equation}
+  \frac{d\hat{\mathbf{n}}}{dt} = \left [ \left ( 1 - \frac{t}{h} \right)
+  \mathbf{A}(\mathbf{n}_i) + \frac{t}{h} \mathbf{A}\left(\mathbf{n}_{i+1}^p \right) \right ] \hat{\mathbf{n}}
+\end{equation}
 

### 2020-05-21 manuscript.tex
Continued work on methodology

@@
 \newtheorem{lemma}{Lemma}
+
+\newcommand{\vect}[1]{\mathbf{#1}} % vectors and matrices
 
@@
 on the density of other nuclides, $\{N_j \,|\, j \ne i\}$, \cref{eq:depletion}
-represents a system of first-order differential equations. To form a proper
-initial value problem, the nuclide densities at time 0 must be specified:
+represents a system of first-order ordinary differential equations (ODEs). To
+form a proper initial value problem, the nuclide densities at time 0 must be
+specified:
 \begin{equation}
@@
   \label{eq:depletion-matrix-t}
-  \frac{d\mathbf{n}}{dt} = \mathbf{A}(\mathbf{n},t)\mathbf{n}, \quad \mathbf{n}(0) =
-  \mathbf{n}_0
+  \frac{d\vect{n}}{dt} = \vect{A}(\vect{n},t)\vect{n}, \quad \vect{n}(0) =
+  \vect{n}_0
 \end{equation}
@@
 \begin{equation}
-  \mathbf{n} = \begin{pmatrix} N_1 \\ N_2 \\ \vdots \\ N_n \end{pmatrix}, \quad
-  \mathbf{n}_0 = \begin{pmatrix} N_{1,0} \\ N_{2,0} \\ \vdots \\ N_{n,0} \end{pmatrix}
+  \vect{n} = \begin{pmatrix} N_1 \\ N_2 \\ \vdots \\ N_n \end{pmatrix}, \quad
+  \vect{n}_0 = \begin{pmatrix} N_{1,0} \\ N_{2,0} \\ \vdots \\ N_{n,0} \end{pmatrix}
 \end{equation}
-and $\mathbf{A}(\mathbf{n},t) \in \mathbb{R}^{n\times n}$ is the burnup matrix
+and $\vect{A}(\vect{n},t) \in \mathbb{R}^{n\times n}$ is the burnup matrix
 containing the decay and transmutation coefficients. Note that the burnup matrix
-depends on $\mathbf{n}$ because the solution to the transport equation depends
+depends on $\vect{n}$ because the solution to the transport equation depends
 on the nuclide densities, thereby making \cref{eq:depletion-matrix-t} nonlinear.
@@
 equation can be solved as a time-independent equation. Without loss of
-generality, let us write the burnup matrix solely as a function of $\mathbf{n}$
+generality, let us write the burnup matrix solely as a function of $\vect{n}$
 so that \cref{eq:depletion-matrix-t} becomes
@@
   \label{eq:depletion-matrix}
-  \frac{d\mathbf{n}}{dt} = \mathbf{A}(\mathbf{n})\mathbf{n}, \quad \mathbf{n}(0) =
-  \mathbf{n}_0.
+  \frac{d\vect{n}}{dt} = \vect{A}(\vect{n})\vect{n}, \quad \vect{n}(0) =
+  \vect{n}_0.
 \end{equation}
@@
 the transport equation, and hence the burnup matrix is uniquely determined by
-the value of $\mathbf{n}$.
+the value of $\vect{n}$.
 
 The principle difficulty of solving \cref{eq:depletion-matrix} is the fact that
-the elements of $\mathbf{A}$ are not constant with respect to time. If
-$\mathbf{A}$ is constant, \cref{eq:depletion-matrix} admits a closed-form
+the elements of $\vect{A}$ are not constant with respect to time. If
+$\vect{A}$ is constant, \cref{eq:depletion-matrix} admits a closed-form
 solution,
@@
   \label{eq:constant-A}
-  \mathbf{n}(t) = \exp \left (\mathbf{A} t \right ) \mathbf{n}_0.
+  \vect{n}(t) = \exp \left (\vect{A} t \right ) \vect{n}_0.
 \end{equation}
-The exponential of a matrix $\mathbf{X}$ is defined by the power series
+The exponential of a matrix $\vect{X}$ is defined by the power series
 expansion
 \begin{equation}
-    \exp(\mathbf{X}) = \sum\limits_{k=0}^\infty \frac{1}{k!} \mathbf{X}^k
+    \exp(\vect{X}) = \sum\limits_{k=0}^\infty \frac{\vect{X}^k}{k!}
 \end{equation}
-where $\mathbf{X}^0 = \mathbf{I}$. Most numerical methods for solving
+where $\vect{X}^0 = \vect{I}$. Most numerical methods for solving
 \cref{eq:depletion-matrix} involve two separate components:
@@
 solutions in discrete increments. In general, the problem we need to solve then
-is: given $\mathbf{n}_i \equiv \mathbf{n}(t_i)$, what is the value of
-$\mathbf{n}_{i+1} \equiv \mathbf{n}(t_{i+1}) = \mathbf{n}(t_i + h)$, with $h$
+is: given $\vect{n}_i \equiv \vect{n}(t_i)$, what is the value of
+$\vect{n}_{i+1} \equiv \vect{n}(t_{i+1}) = \vect{n}(t_i + h)$, with $h$
 being the step size? Without loss of generality, we shall choose $t_i=0$ as it
@@
 timestep is known as the \emph{predictor} method. It proceeds by assuming that
-$\mathbf{A}$ is constant over the timestep and assumes the value at the
+$\vect{A}$ is constant over the timestep and assumes the value at the
 beginning of the step. The solution then follows directly from
@@
 \begin{equation}
-    \mathbf{n}_{i+1} = \exp\left(h\mathbf{A}(\mathbf{n}_i) \right) \mathbf{n}_i.
+    \vect{n}_{i+1} = \exp\left(h\vect{A}(\vect{n}_i) \right) \vect{n}_i.
 \end{equation}
@@
 beginning of the step. This method has the virtue of requiring a single
-transport solution and a single matrix exponential but is only first-order
-accurate.
+transport solution and a single matrix exponential per timestep but is only
+first-order accurate.
 
@@
 the CE/CM algorithm, which is the default algorithm used in
-MCNP6~\citep{fensin2006tans}. In this method, $\mathbf{A}$ is evaluated at the
+MCNP6~\citep{fensin2006tans}. In this method, $\vect{A}$ is evaluated at the
 beginning of the step and used to deplete the composition over half the step.
-Then, $\mathbf{A}$ is reevaluated using the predicted middle-of-step composition
+Then, $\vect{A}$ is reevaluated using the predicted middle-of-step composition
 and used to deplete over the full step. The algorithm can be expressed as
-\begin{align}
-    \mathbf{n}_{i+1/2} &= \exp \left (\frac{h}{2}\mathbf{A}(\mathbf{n}_i) \right) \mathbf{n}_i \\
-    \mathbf{n}_{i+1} &= \exp \left(h \mathbf{A}(\mathbf{n}_{i+1/2}) \right) \mathbf{n}_i.
-\end{align}
+\begin{equation}
+  \begin{split}
+    \vect{n}_{i+1/2} &= \exp \left (\frac{h}{2}\vect{A}(\vect{n}_i) \right) \vect{n}_i \\
+    \vect{n}_{i+1} &= \exp \left(h \vect{A}(\vect{n}_{i+1/2}) \right) \vect{n}_i.
+  \end{split}
+\end{equation}
 This method requires two transport solutions and two matrix exponential solves
-(twice as expensive as the predictor method) and is second-order accurate.
+per timestep (twice as expensive as the predictor method) and is second-order
+accurate.
 
@@
 \citet{isotalo2011ane2,isotalo2011ane3}. In the CE/LI method, the end-of-step
-composition is first predicted using $\mathbf{A}(\mathbf{n}_i)$:
+composition is first predicted using $\vect{A}(\vect{n}_i)$:
 \begin{equation}
-  \mathbf{n}_{i+1}^p = \exp \left( h\mathbf{A}(\mathbf{n}_i) \right) \mathbf{n}_i
+  \vect{n}_{i+1}^p = \exp \left( h\vect{A}(\vect{n}_i) \right) \vect{n}_i.
 \end{equation}
-Then, $\mathbf{n}_{i+1}^p$, is used to reevaluate $\mathbf{A}$. An approximation
-of $\mathbf{A}$ that depends on time is formed by linearly interpolating between
-$\mathbf{A}(\mathbf{n}_i)$ and $\mathbf{A}\left(\mathbf{n}_{i+1}^p\right)$. This
-contitutes a new differential equation for an approximation of $\mathbf{n}$ that
-we shall denote $\hat{\mathbf{n}}$,
+Then, $\vect{n}_{i+1}^p$ is used to reevaluate $\vect{A}$. An approximation of
+$\vect{A}$ that depends on time is formed by linearly interpolating between
+$\vect{A}(\vect{n}_i)$ and $\vect{A}\left(\vect{n}_{i+1}^p\right)$.
+\Cref{eq:depletion-matrix} thus becomes
 \begin{equation}
-  \frac{d\hat{\mathbf{n}}}{dt} = \left [ \left ( 1 - \frac{t}{h} \right)
-  \mathbf{A}(\mathbf{n}_i) + \frac{t}{h} \mathbf{A}\left(\mathbf{n}_{i+1}^p \right) \right ] \hat{\mathbf{n}}
+  \label{eq:linear-ode}
+  \frac{d\vect{n}}{dt} = \hat{\vect{A}}(t) \vect{n}
 \end{equation}
+where
+\begin{equation}
+  \label{eq:linear-interp}
+  \hat{\vect{A}}(t) = \left ( 1 - \frac{t}{h} \right) \vect{A}(\vect{n}_i) +
+  \frac{t}{h} \vect{A}\left(\vect{n}_{i+1}^p \right).
+\end{equation}
+We see that $\hat{\vect{A}}$ no longer depends on $\vect{n}$ thereby making
+\cref{eq:linear-ode} a linear nonautonomous ODE. Because $\hat{\vect{A}}$ is a
+function of time, an approximation is still needed to obtain $\vect{n}_{i+1}$.
+The simplest approach~\citep{isotalo2011ane2} is to average
+\cref{eq:linear-interp} over the step,
+\begin{equation}
+  \bar{\vect{A}} = \int_0^h dt \, \hat{\vect{A}}(t) = \frac{\vect{A}(\vect{n}_i)
+  + \vect{A}(\vect{n}_{i+1}^p)}{2},
+\end{equation}
+and then use $\bar{\vect{A}}$ to deplete over the full step:
+\begin{equation}
+  \vect{n}_{i+1} = \exp \left ( h \bar{\vect{A}} \right ) \vect{n}_i.
+\end{equation}
+\citet{isotalo2011ane3} suggested an improvement over this approach wherein
+$\hat{\vect{A}}$ is averaged over substeps instead of the entire interval. That
+is, if we divide the timestep into $m$ substeps, the solution is given by
+\begin{equation}
+  \begin{split}
+    \label{eq:celi-substeps}
+    \vect{A}_s &= \int_{(s-1)h/m}^{sh/m} dt \; \hat{\vect{A}}(t), \quad s=1,\dots,m \\
+    \vect{n}_{i+1} &= \exp\left(\vect{A}_m \right) \exp \left(\vect{A}_{m-1}\right)
+    \cdots \exp\left(\vect{A}_1\right) \vect{n}_i.
+  \end{split}
+\end{equation}
+While using substeps is a reasonable approach for solving
+\cref{eq:linear-ode}, there are other methods as well. For example, some
+codes use a backward differentiation formula rather than a method based on an
+exponential integrator~\cite{carpenter2009mc,hykes2017mc,sublet2017nds}.
+
+\subsubsection{Magnus expansion integrators}
+
+Any linear approximation of \cref{eq:depletion-matrix} characterized by
+$\vect{A}$ that depends only on time, such as \cref{eq:linear-ode}, has an
+infinite series solution known as the Magnus expansion~\citep{blanes2009pr}.
+This expansion takes the form
+\begin{equation}
+  \begin{split}
+  \vect{n}(t) &= \exp \left( \vect{\Omega}(t) \right ) \vect{n}_0 \\
+  \vect{\Omega}(t) &= \sum\limits_{k=1}^\infty \vect{\Omega}_k(t).
+  \end{split}
+\end{equation}
+The first three terms of the expansion are given by
+\begin{equation}
+  \label{eq:magnus-terms}
+  \begin{split}
+    \vect{\Omega}_1(t) &= \int_0^t dt_1 \vect{A}(t_1) \\
+    \vect{\Omega}_2(t) &= \int_0^t dt_1 \int_0^{t_1} dt_2 [\vect{A}(t_1), \vect{A}(t_2) ] \\
+    \vect{\Omega}_3(t) &= \int_0^t dt_1 \int_0^{t_1} dt_2 \int_0^{t_2} dt_3
+      \left ( [\vect{A}(t_1), [\vect{A}(t_2), \vect{A}(t_3)]] + [\vect{A}(t_3), [\vect{A}(t_2), \vect{A}(t_1)]] \right )
+  \end{split}
+\end{equation}
+where $[\vect{A},\vect{B}] = \vect{A}\vect{B} - \vect{B}\vect{A}$ is a matrix
+commutator. Subsequent terms can be defined through a recurrence relation.
+
+The Magnus expansion can be used as the basis for developing numerical
+integration algorithms and has been widely studied. The typical approach is to
+truncate the $\vect{\Omega}$ series at an appropriate order and apply an
+approximation to the multivariate integrals in \cref{eq:magnus-terms}. In
+particular, OpenMC uses a fourth-order commutator-free exponential
+integrator~\citep{blanes2006anm} that is derived from the Magnus expansion in
+order to solve \cref{eq:linear-ode}. In this approximation, the Magnus expansion
+is truncated at two terms and then expressed as the composition of two
+exponentials. For a system of ODEs in the form of \cref{eq:linear-ode}, the
+resulting method is
+\begin{equation}
+  \label{eq:cf4}
+  \vect{n}_{i+1} = \exp (a_2 \vect{A}_1 + a_1 \vect{A}_2)
+  \exp (a_1 \vect{A}_1 + a_2 \vect{A}_2) \vect{n}_i
+\end{equation}
+where $\vect{A}_1$ and $\vect{A}_2$ correspond $\hat{\vect{A}}$ evaluated at the
+nodes of a two-point Gauss-Legendre quadrature over the interval $[0,h]$:
+\begin{equation}
+  \begin{split}
+    a_1 &= \frac{1}{4} + \frac{\sqrt{3}}{6}, \; a_2 = \frac{1}{4} - \frac{\sqrt{3}}{6} \\
+    c_1 &= \frac{1}{2} - \frac{\sqrt{3}}{6}, \; c_2 = \frac{1}{2} + \frac{\sqrt{3}}{6} \\
+    \vect{A}_1 &= h\hat{\vect{A}}\left ( c_1 h\right ), \; \vect{A}_2 = h\hat{\vect{A}}\left ( c_2 h\right ).
+  \end{split}
+\end{equation}
+Evaluating $\vect{A}_1$ and $\vect{A}_2$ using \cref{eq:linear-interp} and
+substituting in \cref{eq:cf4}, we obtain a CE/LI method based on the
+fourth-order commutator-free Magnus integrator:
+\begin{equation}
+  \vect{n}_{i+1} = \exp \left( \frac{h}{12} \vect{A}(\vect{n}_i) + \frac{5h}{12} \vect{A}(\vect{n}_{i+1}^p) \right)
+  \exp \left( \frac{5h}{12} \vect{A}(\vect{n}_i) + \frac{h}{12} \vect{A}(\vect{n}_{i+1}^p) \right) \vect{n}_i.
+\end{equation}
+
+OpenMC does include
+both CE/LI and LE/QI methods~\cite{isotalo2011ane2}, but the approach for
+solving \cref{eq:linear-ode} is a bit different.
 
@@
 almost entirely by the time to compute a transport solution, i.e., to evaluate
-$\mathbf{A}$ for a given $\mathbf{n}$. Thus, the cost of a method scales with
-the number of $\mathbf{A}$ evaluations that are performed per timestep. On the
+$\vect{A}$ for a given $\vect{n}$. Thus, the cost of a method scales with
+the number of $\vect{A}$ evaluations that are performed per timestep. On the
 other hand, methods that require more evaluations generally achieve higher
@@
 \begin{equation}
-    \exp(\mathbf{A}t) \approx \alpha_0 \prod\limits_{\ell=1}^{k/2} \left (
-    \mathbf{I} + 2 \text{Re} \left ( \widetilde{\alpha}_\ell \left (\mathbf{A}t
-    - \theta_\ell \mathbf{I} \right )^{-1} \right ) \right )
+    \exp(\vect{A}t) \approx \alpha_0 \prod\limits_{\ell=1}^{k/2} \left (
+    \vect{I} + 2 \text{Re} \left ( \widetilde{\alpha}_\ell \left (\vect{A}t
+    - \theta_\ell \vect{I} \right )^{-1} \right ) \right )
 \end{equation}
@@
   \begin{algorithmic}[1]
-    \State $\mathbf{n} \gets \mathbf{n_0}$
+    \State $\vect{n} \gets \vect{n_0}$
     \For{$\ell \gets 1$ \textbf{to} $k/2$}
-      \State $\mathbf{n} \gets \mathbf{n} + 2\text{Re}(\widetilde{\alpha}_\ell
-        (\mathbf{A}t - \theta_\ell)^{-1})\mathbf{n}$
+      \State $\vect{n} \gets \vect{n} + 2\text{Re}(\widetilde{\alpha}_\ell
+        (\vect{A}t - \theta_\ell)^{-1})\vect{n}$
     \EndFor
-    \State $\mathbf{n} \gets \alpha_0 \mathbf{n}$
+    \State $\vect{n} \gets \alpha_0 \vect{n}$
   \end{algorithmic}

### 2020-05-22 manuscript.tex
More or less complete description of CE/LI and LE/QI

@@
 More sophisticated predictor-corrector methods have been introduced by
-\citet{isotalo2011ane2,isotalo2011ane3}. In the CE/LI method, the end-of-step
-composition is first predicted using $\vect{A}(\vect{n}_i)$:
+\citet{isotalo2011ane2,isotalo2011ane3}. In the CE/LI (constant extrapolation,
+linear interpolation) method, the end-of-step composition is first predicted
+using $\vect{A}(\vect{n}_i)$:
 \begin{equation}
@@
 approximation to the multivariate integrals in \cref{eq:magnus-terms}. In
-particular, OpenMC uses a fourth-order commutator-free exponential
+particular, OpenMC uses a fourth-order commutator-free\footnote{The integrator
+is called commutator-free because it does not require computing any matrix
+commutators like those that appear in \cref{eq:magnus-terms}. Other methods
+based on the Magnus expansion do require computing matrix commutators, some of
+which were explored by \citet{josey2017phd}.} exponential
 integrator~\citep{blanes2006anm} that is derived from the Magnus expansion in
@@
 \begin{equation}
-  \label{eq:cf4}
+  \label{eq:compose4}
   \vect{n}_{i+1} = \exp (a_2 \vect{A}_1 + a_1 \vect{A}_2)
@@
 Evaluating $\vect{A}_1$ and $\vect{A}_2$ using \cref{eq:linear-interp} and
-substituting in \cref{eq:cf4}, we obtain a CE/LI method based on the
+substituting in \cref{eq:compose4}, we obtain a CE/LI method based on the
 fourth-order commutator-free Magnus integrator:
 \begin{equation}
-  \vect{n}_{i+1} = \exp \left( \frac{h}{12} \vect{A}(\vect{n}_i) + \frac{5h}{12} \vect{A}(\vect{n}_{i+1}^p) \right)
-  \exp \left( \frac{5h}{12} \vect{A}(\vect{n}_i) + \frac{h}{12} \vect{A}(\vect{n}_{i+1}^p) \right) \vect{n}_i.
+  \label{eq:celi-cfq4}
+  \begin{split}
+  \vect{n}_{i+1}^p &= \exp \left ( h \vect{A}(\vect{n}_i ) \right ) \\
+  \vect{n}_{i+1} &= \exp \left( \frac{h}{12} \vect{A}(\vect{n}_i) +
+    \frac{5h}{12} \vect{A}(\vect{n}_{i+1}^p) \right)
+  \exp \left( \frac{5h}{12} \vect{A}(\vect{n}_i) +
+  \frac{h}{12} \vect{A}(\vect{n}_{i+1}^p) \right) \vect{n}_i.
+  \end{split}
 \end{equation}
+\Cref{eq:celi-cfq4} requires two transport solutions and two matrix
+exponentials, whereas the CE/LI method with substeps, \cref{eq:celi-substeps},
+requires two transport solutions and $m$ matrix exponentials. A comparison of
+these two approaches has shown that \cref{eq:celi-cfq4} achieves better accuracy
+than the use of five substeps~\citep{josey2017phd}. OpenMC also includes an
+integrator based on the LE/QI (linear extrapolation, quadratic interpolation)
+method~\citep{isotalo2011ane2} where the integration of \cref{eq:linear-ode} (in
+which case $\hat{\vect{A}}(t)$ is represented by a quadratic polynomial in $t$)
+is performed using the fourth-order commutator-free Magnus integrator based on
+\cref{eq:compose4}.
 
-OpenMC does include
-both CE/LI and LE/QI methods~\cite{isotalo2011ane2}, but the approach for
-solving \cref{eq:linear-ode} is a bit different.
+\subsubsection{Lie Group Integrator}
 
-Our aim here is not to exhaustively describe all integration methods but rather
-to give a few examples that elucidate the main considerations one must take into
-account when choosing a method. Generally, there is a tradeoff between the
-accuracy of the method and its computational expense. The expense is driven
-almost entirely by the time to compute a transport solution, i.e., to evaluate
-$\vect{A}$ for a given $\vect{n}$. Thus, the cost of a method scales with
-the number of $\vect{A}$ evaluations that are performed per timestep. On the
-other hand, methods that require more evaluations generally achieve higher
-accuracy. The predictor method only requires one evaluation and its error
-converges as $\mathcal{O}(h)$. The CE/CM method requires two evaluations and is
-thus twice as expensive as the predictor method, but achieves an error of
-$\mathcal{O}(h^2)$. An exhaustive description of time integration methods and
-their merits can be found in the thesis of \citet{josey2017phd}.
+\citep{celledoni2004fgcs}
+
+\subsubsection{Stochastic Implicit Euler Integrators}
+
+\subsubsection{Summary}
 
 OpenMC does not rely on a single time integration method but rather has several
-classes that implement different algorithms. For example, the
-openmc.deplete.PredictorIntegrator class implements the predictor
-method, and the openmc.deplete.CECMIntegrator class implements the
-CE/CM method. A full list of the integrator classes available can be found in
-the documentation for the openmc.deplete module.
+classes that implement different algorithms. Generally, there is a tradeoff
+between the accuracy of the method, the computational expense, and the memory
+requirements. When a MC code is used, the expense is almost entirely by the time
+to compute the transport solution, i.e., to evaluate $\vect{A}$ for a given
+$\vect{n}$. Thus, the cost of a method scales with the number of $\vect{A}$
+evaluations that are performed per timestep. On the other hand, methods that
+require more evaluations generally achieve higher accuracy. The predictor method
+only requires one evaluation and its error converges as $\mathcal{O}(h)$. The
+CE/CM method requires two evaluations and is thus twice as expensive as the
+predictor method, but achieves an error of $\mathcal{O}(h^2)$. An exhaustive
+description of time integration methods and their merits can be found in the
+thesis of \citet{josey2017phd}.
 

### 2020-05-22 manuscript.tex
Add section on CF4 method

@@
 
-\citep{celledoni2004fgcs}
+OpenMC also includes an integrator based on a commutator-free Lie group
+method~\citep{celledoni2004fgcs}. In particular, a fourth-order scheme called
+CF4 is used that is given as
+\begin{equation}
+  \begin{split}
+    \vect{A}_1 &= h\vect{A}(\vect{n}_0) \\
+    \hat{\vect{n}}_1 &= \exp \left ( \frac{\vect{A}_1}{2} \right ) \\
+    \vect{A}_2 &= h\vect{A}(\hat{\vect{n}}_1) \\
+    \hat{\vect{n}}_2 &= \exp \left ( \frac{\vect{A}_2}{2} \right ) \\
+    \vect{A}_3 &= h \vect{A}(\hat{\vect{n}}_2) \\
+    \hat{\vect{n}}_3 &= \exp \left ( -\frac{\vect{A}_1}{2}  + \vect{A}_3 \right ) \\
+    \vect{A}_4 &= h\vect{A}(\hat{\vect{n}}_3) \\
+    \vect{n}_{i+1} &= \exp \left ( \frac{\vect{A}_1}{4} + \frac{\vect{A}_2}{6} + \frac{\vect{A}_3}{6} - \frac{\vect{A}_4}{12} \right )
+    \exp \left ( -\frac{\vect{A}_1}{12} + \frac{\vect{A}_2}{6} + \frac{\vect{A}_3}{6} - \frac{\vect{A}_4}{4} \right ) \vect{n}_i.
+  \end{split}
+\end{equation}
+This method requires four transport solutions and five matrix exponentials per
+timestep. However, because it is fourth-order, it allows large timesteps to be
+used without sacrificing accuracy. For example, on a test problem based on a
+3$\times$3 fuel pin array with gadolinum as a burnable absorber, the CF4 method
+allowed 20 day timesteps while still achieving an accuracy of of 0.1\% on the
+$^{157}$Gd number density~\citep{josey2017phd}.
 

### 2020-05-27 manuscript.tex
Start adding EPC-RK4 and SIE integrators

@@
 
-The equation that governs the transmutation and decay of nuclides inside of an
-irradiated environment can be written as
+The equation that governs the transmutation of nuclides by nuclear reactions and
+radioactive decay inside of an irradiated environment can be written as
 \begin{equation}
@@
 depleted separately. Ultimately, the user is responsible for deciding what level
-of discretization is adequate for the problem at hand.
+of spatial discretization is adequate for the problem at hand.
 
@@
 \end{equation}
-where $\vect{X}^0 = \vect{I}$. Most numerical methods for solving
-\cref{eq:depletion-matrix} involve two separate components:
+where $\vect{X}^0 = \vect{I}$. Solving \cref{eq:depletion-matrix} usually
+involves two separate components:
 \begin{enumerate}
-  \item Approximating \cref{eq:depletion-matrix} in a form that involves taking
-  one or more matrix exponentials, and
+  \item Use of a numerical method to integrate \cref{eq:depletion-matrix}
+  forward in time, usually in a form that involves taking one or more matrix
+  exponentials, and
   \item Evaluating the matrix exponential or, equivalently, the action of a
@@
 Because an analytical solution to \cref{eq:depletion-matrix} is not available,
-one must resort to using a numerical method that generates approximation
-solutions in discrete increments. In general, the problem we need to solve then
-is: given $\vect{n}_i \equiv \vect{n}(t_i)$, what is the value of
-$\vect{n}_{i+1} \equiv \vect{n}(t_{i+1}) = \vect{n}(t_i + h)$, with $h$
-being the step size? Without loss of generality, we shall choose $t_i=0$ as it
-makes the presentation of various methods simpler.
+one must resort to using a numerical method that generates approximate solutions
+in discrete increments. The problem we need to solve then is: given $\vect{n}_i
+\equiv \vect{n}(t_i)$, what is the value of $\vect{n}_{i+1} \equiv
+\vect{n}(t_{i+1}) = \vect{n}(t_i + h)$, with $h$ being the step size? Without
+loss of generality, we shall choose $t_i=0$ as it makes the presentation of
+various methods simpler.
 
@@
 outline the available methods in OpenMC, several of which have not been
-described in the literature.
+previously described in the literature.
 
@@
 \begin{equation}
+  \label{eq:predictor}
   \vect{n}_{i+1} = \exp\left(h\vect{A}(\vect{n}_i) \right) \vect{n}_i.
@@
 
-\subsubsection{Predictor-Corrector Methods}
+\subsubsection{predictor--Corrector Methods}
 
-A series of \emph{predictor-corrector} methods that use multiple stages offer
+A series of \emph{predictor--corrector} methods that use multiple stages offer
 improved accuracy over the predictor method. The simplest of these methods is
@@
 
-More sophisticated predictor-corrector methods have been introduced by
+More sophisticated predictor--corrector methods have been introduced by
 \citet{isotalo2011ane2,isotalo2011ane3}. In the CE/LI (constant extrapolation,
@@
 codes use a backward differentiation formula rather than a method based on an
-exponential integrator~\cite{carpenter2009mc,hykes2017mc,sublet2017nds}.
+exponential integrator~\citep{carpenter2009mc,hykes2017mc,sublet2017nds}.
 
@@
 OpenMC also includes an integrator based on a commutator-free Lie group
-method~\citep{celledoni2004fgcs}. In particular, a fourth-order scheme called
-CF4 is used that is given as
+integration method~\citep{celledoni2004fgcs}. In particular, a fourth-order
+scheme called CF4 is used that is given as
 \begin{equation}
@@
 
+\subsubsection{Extended Predictor--Corrector Methods}
+
+One option to improve on the traditional predictor--corrector methods is to add
+intermediate steps at which $\vect{A}$ is evaluated and combine them in such a
+way as to reduce the error of the solution. The general method would take the
+following form:
+\begin{equation}
+  \label{eq:epc}
+  \begin{split}
+    \hat{\vect{n}}_1 &= \vect{n}_i \\
+    \hat{\vect{n}}_k &= \exp \left( h \sum_{j=1}^{k-1} a_{ij} \vect{A}(\hat{\vect{n}}_j) \right) \vect{n}_i \\
+    \vect{n}_{i+1} &= \exp \left( h \sum_{j=1}^{s} b_{j} \vect{A}(\hat{\vect{n}}_j) \right) \vect{n}_i
+  \end{split}
+\end{equation}
+Noting that the coefficients in \cref{eq:epc} can be arranged in a Butcher
+tableau, \citet{josey2016jcp} proposed using the coefficients for the classic
+fourth Runge--Kutta method:
+\begin{equation}
+  \begin{split}
+    \vect{A}_1 &= h\vect{A}(\vect{n}_0) \\
+    \hat{\vect{n}}_1 &= \exp \left ( \frac{\vect{A}_1}{2} \right ) \\
+    \vect{A}_2 &= h\vect{A}(\hat{\vect{n}}_1) \\
+    \hat{\vect{n}}_2 &= \exp \left ( \frac{\vect{A}_2}{2} \right ) \\
+    \vect{A}_3 &= h \vect{A}(\hat{\vect{n}}_2) \\
+    \hat{\vect{n}}_3 &= \exp \left ( \vect{A}_3 \right ) \\
+    \vect{A}_4 &= h\vect{A}(\hat{\vect{n}}_3) \\
+    \vect{n}_{i+1} &= \exp \left ( \frac{\vect{A}_1}{6} + \frac{\vect{A}_2}{3}
+      + \frac{\vect{A}_3}{3} + \frac{\vect{A}_4}{6} \right ) \vect{n}_i.
+  \end{split}
+\end{equation}
+
 \subsubsection{Stochastic Implicit Euler Integrators}
+
+All of the integration methods discussed thus far are explicit methods that are
+only conditionally stable. There are many well-documented cases where
+oscillatory behavior and diverging solutions have been observed. To address
+this, \citet{dufek2013ane} proposed an implicit-Euler-like method that is
+unconditionally stable. The main idea is to replace $\vect{A}(\vect{n}_i)$ with
+$\vect{A}(\vect{n}_{i+1})$ in \cref{eq:predictor}:
+\begin{equation}
+  \label{eq:implicit}
+  \vect{n}_{i=1} = \exp \left( h\vect{A}(\vect{n}_{i+1}) \right) \vect{n}_i.
+\end{equation}
+This results in a non-linear equation for $\vect{n}_{i+1}$ that can't be solved
+directly. However, it is possible to solve for $\vect{n}_{i+1}$ using a
+fixed-point iteration. \citet{dufek2013ane} utilized a stochastic approximation
+based on the Robbins-Monro algorithm~\citep{robbins1951ams} in order to
+guarantee convergence of the fixed-point iteration while accounting for
+stochasticity in the MC transport solver.
+
+The major downside of an algorithm based on \cref{eq:implicit} is that, like the
+predictor method, it is only first-order accurate. Implicit methods have also
+been applied to higher-order
+algorithms~\citep{kotlyar2014ane,kotlyar2016ane,cosgrove2020ane}, thereby
+obtaining methods that are unconditionally stable and accurate with longer
+timesteps. In OpenMC, stochastic implicit versions of the CE/LI and LE/QI
+methods have also been implemented. The basis for the algorithm is to iterate on
+$\vect{n}_{i+1}^p$ used in \cref{eq:linear-interp}, obtaining progressively
+better estimates that are then used to solve \cref{eq:linear-ode}. A full
+discussion of the algorithm and its performance and tradeoffs can be found in
+\citep{josey2017phd}.
 
@@
 between the accuracy of the method, the computational expense, and the memory
-requirements. When a MC code is used, the expense is almost entirely by the time
+requirements. When a MC code is used, the expense is dominated by the time
 to compute the transport solution, i.e., to evaluate $\vect{A}$ for a given
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{References}
+%\section*{References}
 

### 2020-05-28 manuscript.tex
Start writing data section

@@
 \usepackage{multirow}
+\usepackage{upgreek}
 
@@
 tabulated for orders up to $k=48$. Rather than computing the full approximation
-and then multiplying it by a vector, the following algorithm\footnote{The
-original description of the algorithm presented by \citet{pusa2016nse} contains
-a typo.} is used to incrementally apply the terms within the product:
+and then multiplying it by a vector, \cref{alg:cram}\footnote{The original
+description of the algorithm presented by \citet{pusa2016nse} contains a typo.}
+is used to incrementally apply the terms within the product. The $k$th order
+approximation for CRAM requires solving $k/2$ sparse linear systems. OpenMC
+relies on functionality from scipy.sparse.linalg for solving the linear system.
 \begin{algorithm}[H]
@@
 \end{algorithm}
-The $k$th order approximation for CRAM requires solving $k/2$ sparse linear
-systems. OpenMC relies on functionality from scipy.sparse.linalg for solving the
-linear systems.
+
+\subsection{Nuclear Data Considerations}
+
+In principle, using CRAM for evaluating matrix exponentials is fairly simple:
+construct the burnup matrix at various times and solve a set of sparse linear
+systems. However, constructing the burnup matrix, $\vect{A}$, involves not only
+solving the transport equation to estimate nuclear reaction rates but also a
+series of choices about what data to include.
+
+\subsubsection{Depletion Chain Generation}
+\label{sec:depletion-chain}
+
+In OpenMC, the burnup matrix is constructed based on data inside of a
+\emph{depletion chain} file, which includes fundamental data gathered from ENDF
+incident neutron, decay, and fission product yield sublibraries. For each
+nuclide, this file includes:
+\begin{itemize}
+  \item What transmutation reactions are possible, their Q values, and their products;
+  \item If a nuclide is not stable, what decay modes are possible, their
+  branching ratios, and their products; and
+  \item If a nuclide is fissionable, the fission products yields at any number
+  of incident neutron energies.
+\end{itemize}
+
+Depletion chains can be automatically created in the Python API in the following
+manner. First, the user supplies a list of ENDF incident neutron, decay, and
+fission product yield sublibrary files. The incident neutron files are used to
+determine what nuclear reactions result in transmutation and what the
+corresponding Q values of the reactions are. The decay files are used to
+determine the half-life and possible decay modes of nuclides that are unstable.
+The fission product yield sublibraries are used to determine what fission
+products are produced when a fission reaction occurs. This information is then
+aggregated and written to an XML file that serves as the depletion chain file
+that is used at runtime to construct a burnup matrix.
+
+In some cases, decay or reaction products appear that do not actually exist in
+the decay sublibrary files. OpenMC uses a set of heuristics to find a suitable
+replacement. First, if the atomic number of the missing product is greater than
+98, it is assumed that the product undergoes alpha decay. Otherwise, for a given
+product, OpenMC will determine the mass number of the isotope of that element
+that has the longest half-life. If the mass number of the missing product is
+less than the mass number of the longest-lived isotope, it is assumed that the
+nuclide will undergo a $\upbeta^+$ decay. Conversely, if the mass number is
+larger than the mass number of the longest-lived isotope, it is assumed that the
+nuclide will undergo a $\upbeta^-$ decay. Once the replacement nuclide has been
+determined, OpenMC will check again if it exists in the decay sublibrary files.
+If not, the process is repeated.
+
+Using all incident neutron, decay, and fission product yield sublibrary files
+from an ENDF-format library, such as ENDF/B-VIII.0, results in a \emph{full}
+depletion chain that typically includes thousands of nuclides. Because the use
+of a chain with $n$ nuclides will result in an $n\times n$ burnup matrix, there
+is an incentive to reduce the number of nuclides in the depletion chain to
+minimize both the size of the burnup matrix and the number of nuclides that are
+tracked during a transport solve. The depletion chain class in OpenMC's Python
+API includes a method for automatically reducing the size of the chain. Starting
+with a set of nuclides that must be included, the method will recursively go
+through all decay and reaction modes to determine what products may be produced.
+Using this method on the ENDF/B-VII.1 library reduces the total number of
+nuclides in the depletion chain from 3,820 nuclides to XXX.
+
+In addition to the full depletion chain, a simplified depletion chain based on
+the VERA depletion benchmark specification~\cite{kim2015casl} was created and
+made available for users at \href{https://openmc.org}. This chain includes 223
+nuclides and has adjusted branching ratios for some nuclides to preserve
+accuracy of nuclide production rates.
+
+\subsubsection{Transmutation Reactions}
+
+OpenMC will setup tallies in a problem based on what transmutation reactions are
+available in a depletion chain file, so any arbitrary number of transmutation
+reactions can be tracked. By default, the automatic depletion chain generation
+described in \cref{sec:depletion-chain} will include the following transmutation
+reactions: fission, (n,$\gamma$), (n,2n), (n,3n), (n,4n), (n,p), and
+(n,$\alpha$). While this is suitable for fission applications, fusion
+applications may require a wider set of transmutation reactions to be tracked.
+
+\subsubsection{Capture Branching Ratios}
+
+(n,$\gamma$) reactions may result in a product nucleus being in either the
+ground state or an excited state. If the excited state is a metastable state
+(having an unusually long half-life), it is important to account for this
+difference during depletion. The most well-known example is capture in
+$^{241}$Am, which can produce either $^{242}$Am or $^{242m}$Am. Because the
+metastable state has a significantly longer half-life than the ground state, it
+is important to account for the branching of the capture reaction in $^{241}$Am
+to accurately predict the production of higher actinides~\citep{haeck2012nse}.
+This is complicated by the fact that the branching ratio may depend on the
+incident neutron energy causing capture.
+
+OpenMC does not currently allow energy-dependent capture branching ratios.
+However, the depletion chain file does allow a reaction to be listed multiple
+times with different branching ratios that result in different products.
+Spectrum-averaged capture branching ratios have been computed in LWR and SFR
+spectra and are available at \href{https://openmc.org/depletion-chains}. Within
+the Python API, the class representing a depletion chain allows branching ratios
+to be easily retrieved and changed. We note that while this treatment doesn't
+accurately account for the change in the capture branching ratio with increasing
+burnup (due to spectral changes), these changes are currently much smaller than
+the differences in capture branching ratios between nuclear data evaluations.
+
+\subsubsection{Fission Product Yields}
+
+Like capture branching ratios, fission product yields (FPYs) also depend on the
+incident neutron energy. ENDF fission product yield sublibraries typically
+include yields tabulated at two or three energies (thermal, fast, and fusion
+energies). Complicating matters is the fact that while these data are specified
+for a single energy, they actually represent the observed FPYs for a spectrum of
+energies (characterized by a single point). For example, FPYs are often given at
+\SI{0.0253}{\electronvolt}, but this actually represents the observed FPY under
+a thermal neutron spectrum. The ENDF-6 formats manual~\citep{trkov2018bnl}
+recommends using linear-linear interpolation between energy points. However, at
+the end of a transport solve, we typically have a one-group or multigroup
+fission reaction rate---it's not clear how one should use this information to
+calculate an effective FPY.
+
+Different codes have adopted different approaches in how to handle the energy
+dependence of FPYs. In ORIGEN, the average neutron energy causing fission is
+calculated, and FPYs are interpolated using this average
+energy~\citep{gauld2011nt}. In MCNP, the fission rate is tallied over a few
+energy groups (thermal, fast, and high energy), and the energy group that
+contains the majority of fissions is used to select appropriate
+FPYs~\citep{fensin2008nt}. In Serpent, a linear interpolation factor is scored
+based on the neutron energy causing fission in relationship to the energy points
+at which FPY are tabulated; the interpolation factor is then used to calculate
+an effective FPY~\citep{kunchev2019mc}.
+
+Rather than choose a single method, OpenMC includes three methods for treating
+the energy dependence of FPY. In the first method, FPY data corresponding to a
+specific energy is manually selected. This approach may be appropriate if it is
+known that neutrons in the system under consideration have a typical thermal,
+fast, or fusion energy spectrum. In the second method, fission rates are tallied
+above and below a specified cutoff energy. Then, thermal FPYs are used for all
+fissions below the cutoff energy, and fast FPYs are used for all fissions above
+the cutoff energy. In the third method, the average energy at which fission
+events occur is tallied,
+\begin{equation}
+  \bar{E} = \frac{\int_0^\infty dE \; E \sigma_f(E)\phi(E)}{\int_0^\infty \sigma_f(E) \phi(E)},
+\end{equation}
+where $\sigma_f$ is the microscopic fission cross section and $\phi$ is the
+neutron flux. An effective FPY is then calculated by linearly interpolating
+between tabulated FPYs using the average energy. This approach is effectively
+the same as that used by ORIGEN~\citep{gauld2011nt}.
+
+\subsubsection{Power Normalization}
+
+The reaction rates calculated by OpenMC are given in units of reactions per
+source particle. For depletion, it is necessary to compute an absolute reaction
+rate in reactions per second. To do so, the total observed heating in the system
+is determined, and then the ratio of the system power to the observed heating is
+used as a normalization factor on the reaction rates. OpenMC allows the total
+heating to be determined by two different methods. In the first method, fission
+reaction rates are multiplied by constant Q values in the depletion chain file
+to estimate the total heating. This method is particularly useful for making
+inter-code comparisons, where one must ensure that the same Q values for fission
+are applied. In the second method, the actual heating rate is tallied over the
+entire system (separate from the tally for transmutation reaction rates). This
+method makes it possible to account for the energy dependence of the fission
+energy release data as well as the fact that neutrons and photons may leak out
+of the system, carrying away their energy with them. The latter effect may be
+particularly important for small, leaky systems.
 
@@
 
-In-memory handling through Python API. This allows time loading nuclear data to
-be amortized since it only needs to be done once.
+Class organization with Operator and Integrators separate. Use of shared library
+and in-memory handling through Python API. This allows time loading nuclear data
+to be amortized since it only needs to be done once.
+
+Implementation of depletion in Python. Relying on scipy sparse solve for the
+CRAM system. Parallelization through Python multiprocessing pool.
 
@@
 yield sublibraries are concatenated.
-
-\subsection{}
 

### 2020-05-29 manuscript.tex
Discuss implementation, add depletion chain plot

@@
 \end{equation}
-While using substeps is a reasonable approach for solving
-\cref{eq:linear-ode}, there are other methods as well. For example, some
-codes use a backward differentiation formula rather than a method based on an
-exponential integrator~\citep{carpenter2009mc,hykes2017mc,sublet2017nds}.
+While using substeps is a reasonable approach for solving \cref{eq:linear-ode},
+there are other methods as well. For example, some codes use a backward
+differentiation formula rather than a method based on an exponential
+integrator~\citep{carpenter2009mc,hykes2017mc,sublet2017nds}.
 
@@
 through all decay and reaction modes to determine what products may be produced.
-Using this method on the ENDF/B-VII.1 library reduces the total number of
-nuclides in the depletion chain from 3,820 nuclides to XXX.
+Using this method on the ENDF/B-VII.1 library starting from a list of all stable
+or very long lived (half-life greater than $10^{15}$ sec) nuclides reduces the
+total number of nuclides in the depletion chain from 3,820 to 1,693.
 
@@
 the VERA depletion benchmark specification~\cite{kim2015casl} was created and
-made available for users at \href{https://openmc.org}. This chain includes 223
+made available for users at \url{https://openmc.org}. This chain includes 223
 nuclides and has adjusted branching ratios for some nuclides to preserve
-accuracy of nuclide production rates.
+accuracy of nuclide production rates. \Cref{fig:chains} shows which nuclides are
+included in each of the depletion chains.
+\begin{figure}
+  \centering
+  \includegraphics[width=4in]{figures/chains.pdf}
+  \caption{Table of nuclides showing which nuclides are included in each of the
+    depletion chains based on ENDF/B-VII.1.}
+  \label{fig:chains}
+\end{figure}
 
@@
 Spectrum-averaged capture branching ratios have been computed in LWR and SFR
-spectra and are available at \href{https://openmc.org/depletion-chains}. Within
+spectra and are available at \url{https://openmc.org/depletion-chains}. Within
 the Python API, the class representing a depletion chain allows branching ratios
@@
 contains the majority of fissions is used to select appropriate
-FPYs~\citep{fensin2008nt}. In Serpent, a linear interpolation factor is scored
-based on the neutron energy causing fission in relationship to the energy points
-at which FPY are tabulated; the interpolation factor is then used to calculate
-an effective FPY~\citep{kunchev2019mc}.
+FPYs~\citep{fensin2008nt,fensin2011lanl}. In Serpent, a linear interpolation
+factor is scored based on the neutron energy causing fission in relationship to
+the energy points at which FPY are tabulated; the interpolation factor is then
+used to calculate an effective FPY~\citep{kunchev2019mc}.
 
@@
 
-Class organization with Operator and Integrators separate. Use of shared library
-and in-memory handling through Python API. This allows time loading nuclear data
-to be amortized since it only needs to be done once.
+OpenMC's core particle transport solver is written in C++ and is built as a
+single executable. Most of the user-facing code, however, is written in Python.
+Generally, to run OpenMC, a user writes a Python script that utilizes classes
+and functions defined as part of an application programming interface (API) to
+generate XML input files that are then read by the executable. On top of this,
+OpenMC is also built as a shared library with a C++ API that can be called by
+external programs. OpenMC's Python API includes bindings to the C++ API; this
+enables a single run to be broken up into individual components. For example, a
+function call from the Python bindings to the C++ API can initialize the solver
+(reading inputs, loading cross sections, etc.) without actually performing a
+simulation. A separate function call can run the transport simulation. Further
+function calls allow a power user to obtain information about tally results
+directly in memory from the C++ API.
 
-Implementation of depletion in Python. Relying on scipy sparse solve for the
-CRAM system. Parallelization through Python multiprocessing pool.
+OpenMC's depletion capability is implemented entirely in Python as the
+\texttt{openmc.deplete} module. This module relies on the Python bindings to the
+C++ API to initialize OpenMC, create tallies necessary for depletion, run
+transport simulations, and obtain tally results. This approach has two primary
+advantages. First, the time necessary to load cross sections is amortized
+because it only has to be performed once, as opposed to once per transport
+simulation. Second, using the Python bindings to the C++ API allows results to
+be read directly in memory rather than using filesystem I/O, which helps to
+minimizing overhead, particularly for large simulations.
+
+The \texttt{openmc.deplete} module is designed with three key abstract classes
+that represent the major components of the depletion solver. Each of the
+abstract classes themselves defines an interface, that is, a set of methods that
+a subclass must implement. First, the abstract \texttt{TransportOperator} class
+represents a particle transport solver. Subclasses of \texttt{TransportOperator}
+are required to implement a method that takes a list of nuclide densities to be
+used in the problem and a system power and returns the $k$-eigenvalue and the
+reaction rates resulting from the transport operator. Currently, there is a
+single subclass of \texttt{TransportOperator} that interfaces with the Python
+bindings to OpenMC's C++ API to call a tranport solve and obtain reaction rates.
+Once reaction rates are obtained, a separate depletion chain class (that is not
+abstract) uses them to construct the burnup matrix, $\vect{A}$, based on the
+data present in the depletion chain.
+
+The abstract \texttt{DepSystemSolver} class represents a solver for
+\cref{eq:depletion-matrix} when $\vect{A}$ is constant (a solver that can
+evaluate \cref{eq:constant-A}). Subclasses of \texttt{DepSystemSolver} are
+required to implement a method that takes the burnup matrix $\vect{A}$, an
+initial density vector $\vect{n}_0$, and a timestep size, and returns the
+resulting nuclide densities at the end of the step. The only subclass of
+\texttt{DepSystemSolver} at present is \texttt{IPFCramSolver}, which implements
+the IPF form of CRAM to evaluate \cref{eq:constant-A}. Because the CRAM method
+requires solving multiple sparse linear systems, this class relies on the
+third-party \texttt{scipy.sparse} package for performing a sparse linear solve
+(REF).
+
+The abstract \texttt{Integrator} class represents a time integration method.
+This class effectively orchestrates the coupled transport-depletion simulation.
+Subclasses of \texttt{Integrator} are required to implement a method that takes
+an initial vector of nuclide densities, reaction rates returned from an
+operator, a time interval, and a power, and returns the resulting nuclide
+densities at the end of the step. Thus, \texttt{Integrator} utilizes
+\texttt{TransportOperator} in order to obtain reaction rates given $\vect{n}$,
+and once the burnup matrix is constructed, it utilizes \texttt{DepSystemSolver}
+to deplete with a constant $\vect{A}$ as many times as is necessary. One example
+of a subclass of \texttt{Integrator} is \texttt{CECMIntegrator}, which
+implements the CE/CM method.
+
+The depletion system in OpenMC is designed to be modular. Although the only
+transport operator that is implemented is the OpenMC transport solver itself,
+the rest of the infrastructure (time integration methods, CRAM solver) does not
+rely on anything specific to OpenMC's transport solver, so one could in
+principle implement a class that utilizes a different transport solver (whether
+that be another Monte Carlo code or a deterministic code). Similarly, although
+\texttt{IPFCramSolver} is the only depletion system solver implemented, one
+could replace it with a different solver, for example, a solver based on
+backward differentiation formulas.
+
+Lastly, we note that when multiple depletable materials are present in a problem
+(thereby requiring multiple independent solutions of
+\cref{eq:depletion-matrix}), the work is fully parallelized over the available
+processes/threads. The third-party \texttt{mpi4py} Python package is used to
+distribute the material compositions among the participating MPI processes.
+Then, within each process, Python's \texttt{multiprocessing} standard-library
+module is used to distribute materials among threads within a single MPI
+process. This matches the distributed- and shared-memory parallelization scheme
+used in OpenMC's transport solver, which relies on MPI and OpenMP.
 

### 2020-05-29 figures/plot_chains.py

@@ -1,0 +1,61 @@
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+from matplotlib.patches import Rectangle, Patch
+import openmc.deplete
+import openmc.data
+
+# customizations
+mpl.rcParams['axes.autolimit_mode'] = 'round_numbers'
+mpl.rcParams['axes.labelsize'] = 'large'
+mpl.rcParams['axes.xmargin'] = 0
+mpl.rcParams['axes.ymargin'] = 0
+mpl.rcParams['axes.axisbelow'] = True
+mpl.rcParams['font.family'] = 'serif'
+mpl.rcParams['pdf.use14corefonts'] = True
+mpl.rcParams['savefig.bbox'] = 'tight'
+mpl.rcParams['font.size'] = 12.0
+
+chain_full = openmc.deplete.Chain.from_xml('/home/romano/openmc-data/depletion/chain_endfb71_pwr.xml')
+chain_reduced = openmc.deplete.Chain.from_xml('/home/romano/openmc-data/depletion/chain_reduced2_pwr.xml')
+chain_casl = openmc.deplete.Chain.from_xml('/home/romano/openmc-data/depletion/chain_casl_pwr.xml')
+
+def get_patch(z, a, **kwargs):
+    n = a - z
+    return Rectangle((n-0.5, z-0.5), 1, 1, **kwargs)
+
+fig, ax = plt.subplots()
+
+style_full = {'alpha': 0.3}
+style_reduced = {}
+style_casl = {'color': 'orange'}
+
+for nuc in chain_full.nuclides:
+    z, a, m = openmc.data.zam(nuc.name)
+    if m > 0: continue
+    if nuc.name not in chain_reduced:
+        ax.add_patch(get_patch(z, a, **style_full))
+
+for nuc in chain_reduced.nuclides:
+    z, a, m = openmc.data.zam(nuc.name)
+    if m > 0: continue
+    if nuc.name not in chain_casl:
+        ax.add_patch(get_patch(z, a, **style_reduced))
+
+for nuc in chain_casl.nuclides:
+    z, a, m = openmc.data.zam(nuc.name)
+    if m > 0: continue
+    ax.add_patch(get_patch(z, a, **style_casl))
+
+ax.set_xlim(0, 175)
+ax.set_ylim(0, 120)
+ax.set_xlabel('N')
+ax.set_ylabel('Z')
+ax.grid()
+legend_elements = [
[…]

### 2020-05-29 figures/plot_chains.py
Remove Er145 from reduced chain

@@ -16,8 +16,14 @@ mpl.rcParams['savefig.bbox'] = 'tight'
 mpl.rcParams['font.size'] = 12.0
 
 chain_full = openmc.deplete.Chain.from_xml('/home/romano/openmc-data/depletion/chain_endfb71_pwr.xml')
-chain_reduced = openmc.deplete.Chain.from_xml('/home/romano/openmc-data/depletion/chain_reduced2_pwr.xml')
 chain_casl = openmc.deplete.Chain.from_xml('/home/romano/openmc-data/depletion/chain_casl_pwr.xml')
+
+stable = [
+    nuc.name
+    for nuc in chain_full.nuclides
+    if nuc.name != 'Er145' and (nuc.half_life is None or nuc.half_life > 1e15)
+]
+chain_reduced = chain_full.reduce(stable)
 
 def get_patch(z, a, **kwargs):
     n = a - z

### 2020-06-04 manuscript.tex
Minor edits

@@
 applications may require a wider set of transmutation reactions to be tracked.
+Currently, reaction rates are determined by directly tallying the reaction rates
+using a tracklength estimator.
 
@@
 difference during depletion. The most well-known example is capture in
-$^{241}$Am, which can produce either $^{242}$Am or $^{242m}$Am. Because the
+$^{241}$Am, which can produce either $^{242}$Am or $^{242\text{m}}$Am. Because the
 metastable state has a significantly longer half-life than the ground state, it
@@
 
-OpenMC does not currently allow energy-dependent capture branching ratios.
-However, the depletion chain file does allow a reaction to be listed multiple
-times with different branching ratios that result in different products.
-Spectrum-averaged capture branching ratios have been computed in LWR and SFR
-spectra and are available at \url{https://openmc.org/depletion-chains}. Within
-the Python API, the class representing a depletion chain allows branching ratios
-to be easily retrieved and changed. We note that while this treatment doesn't
-accurately account for the change in the capture branching ratio with increasing
-burnup (due to spectral changes), these changes are currently much smaller than
-the differences in capture branching ratios between nuclear data evaluations.
+OpenMC does not currently implement a method to handle energy-dependent capture
+branching ratios directly. However, the depletion chain file does allow a
+reaction to be listed multiple times with different branching ratios that result
+in different products. Spectrum-averaged capture branching ratios have been
+computed in LWR and SFR spectra and are available at
+\url{https://openmc.org/depletion-chains}. Within the Python API, the class
+representing a depletion chain allows branching ratios to be easily retrieved
+and changed. We note that while this treatment doesn't accurately account for
+the change in the capture branching ratio with increasing burnup (due to
+spectral changes), these changes are generally much smaller than the differences
+in capture branching ratios between nuclear data evaluations.
 
@@
 \SI{0.0253}{\electronvolt}, but this actually represents the observed FPY under
-a thermal neutron spectrum. The ENDF-6 formats manual~\citep{trkov2018bnl}
-recommends using linear-linear interpolation between energy points. However, at
-the end of a transport solve, we typically have a one-group or multigroup
-fission reaction rate---it's not clear how one should use this information to
-calculate an effective FPY.
+a thermal neutron spectrum that contains a continuum of energies. The ENDF-6
+formats manual~\citep{trkov2018bnl} recommends using linear-linear interpolation
+between energy points. However, at the end of a transport solve, we typically
+have a one-group or multigroup fission reaction rate---it's not clear how one
+should use this information to calculate an effective FPY.
 
@@
 Rather than choose a single method, OpenMC includes three methods for treating
-the energy dependence of FPY. In the first method, FPY data corresponding to a
+the energy dependence of FPYs. In the first method, FPY data corresponding to a
 specific energy is manually selected. This approach may be appropriate if it is
@@
 requires solving multiple sparse linear systems, this class relies on the
-third-party \texttt{scipy.sparse} package for performing a sparse linear solve
-(REF).
+third-party \texttt{scipy.sparse} package~\cite{virtanen2020nm} for performing a
+sparse linear solve.
 
@@
 
+\subsection{PWR Pincell}
+
+\subsection{SFR Assembly}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-06-23 manuscript.tex
Add discussion of PWR results

@@
 \usepackage{float}
+\usepackage{subcaption}
 \usepackage{algorithm}
@@
 
-In addition to the full depletion chain, a simplified depletion chain based on
-the VERA depletion benchmark specification~\cite{kim2015casl} was created and
-made available for users at \url{https://openmc.org}. This chain includes 223
-nuclides and has adjusted branching ratios for some nuclides to preserve
-accuracy of nuclide production rates. \Cref{fig:chains} shows which nuclides are
-included in each of the depletion chains.
+In addition to the full and reduced depletion chains, a simplified depletion
+chain based on the VERA depletion benchmark specification~\citep{kim2015casl}
+was created and made available for users at \url{https://openmc.org}. This chain
+includes 223 nuclides and has adjusted decay branching ratios for some nuclides
+to preserve the accuracy of nuclide production rates. \Cref{fig:chains} shows
+which nuclides are included in each of the depletion chains.
 \begin{figure}
@@
 \subsubsection{Fission Product Yields}
+\label{sec:fpy}
 
@@
 requires solving multiple sparse linear systems, this class relies on the
-third-party \texttt{scipy.sparse} package~\cite{virtanen2020nm} for performing a
-sparse linear solve.
+third-party \texttt{scipy.sparse} package~\citep{virtanen2020nm} for performing
+a sparse linear solve.
 
@@
 
+To validate the depletion solver in OpenMC, we have carried out simulations
+using the \texttt{openmc.deplete} module as well as Serpent 2.1.31 on two simple
+problems. The first problem is a pressurized water reactor (PWR) pincell and the
+second problem is a single assembly of a sodium-cooled fast reactor (SFR). These
+problems were selected to cover both thermal and fast neutron spectra.
+
+While comparison of simulation predictions to an experiment is a better means of
+ensuring fitness for a particular application, it is not without downsides.
+First, the presence of experimental uncertainties makes it hard to ascertain
+whether small differences are attributable to the uncertainties or errors in the
+code itself. In a cross-code comparison, if the same solution method and
+fundamental nuclear data are being used in both codes, one expects to obtain the
+same answer. Thus, even very minute differences may be indicative of a true
+problem in the method implementation in one of the codes.
+
 \subsection{Data Preparation}
 
-Using ENDF/B-VII.1 files from MCNP distribution. However, for light water
-$S(\alpha,\beta)$, Serpent is not capable of handling continuous format, so the
-file from ENDF/B-VII.0 was used. The evaluation didn't change between VII.0 and
-VII.1, but the processing was different. For Serpent, decay and neutron fission
-yield sublibraries are concatenated.
+Given the remarks above, it is of utmost importance to ensure that the same
+nuclear data are being used in a cross-code comparison. For a depletion
+simulation, we must ensure that the incident neutron cross sections and
+secondary angle/energy distributions, radioactive decay data, and fission
+product yields are the same for both codes. In the simulations performed, all
+data was based on the ENDF/B-VII.1 nuclear data library~\citep{chadwick2011nds}.
+Incident neutron cross sections from the ENDF/B-VII.1 library processed by the
+LANL nuclear data team~\citep{conlin2013lanl}, known as endf71x, were used for
+transport simulations in both OpenMC and Serpent. Serpent is able to use the ACE
+files from the endf71x library directly while OpenMC uses an HDF5 format that
+requires conversion of the ACE files first. For the PWR pincell problem,
+$S(\alpha,\beta)$ data is also needed for hydrogen in light water; because
+Serpent does not yet handle the continuous representation used in the ENDF71SaB
+library, the light water data from ENDF/B-VII.0 was used instead. We note that
+the light water evaluation didn't actually change between ENDF/B-VII.0 and
+ENDF/B-VII.1.
+
+In Serpent, the depletion solver relies directly on ENDF decay and FPY
+sublibrary files. Appropriate decay and FPY libraries for Serpent were thus
+formed by concatenating the ENDF files directly from the ENDF/B-VII.1
+distribution. For OpenMC, cases were run with the full, reduced, and CASL
+depletion chains based on ENDF/B-VII.1 decay and FPY sublibrary files.
+
+To ensure that OpenMC and Serpent perform power normalization in the same
+manner, the default fission heating values from Serpent were used in OpenMC. By
+default, the energy release per fission in $^{235}U$ in Serpent is set to
+\SI{202.27}{\mega\electronvolt}, and the values for other actinides is scaled
+based on the Q value for fission listed in MF=3 in the ENDF evaluation. Without
+the same heating values, the power normalization might result in OpenMC
+depleting materials too quickly or too slowly.
+
+Serpent uses energy-independent capture branching ratios by default that are
+based on the JEFF 3.1 activation file using a PWR flux spectrum. For the PWR
+problem, these energy-independent branching ratios were used in OpenMC. For the
+SFR problem, Serpent was used to first generate representative capture branching
+ratios in the fast flux spectrum using Serpent's energy-dependent capture
+branching ratio treatment. Then, the energy-independent branching ratios were
+set in both Serpent and OpenMC. As discussed in \cref{sec:fpy}, OpenMC also has
+several methods for treating the energy dependence of FPYs. The OpenMC
+simulations were carried out with the three different methods to see what effect
+they have on resulting isotopic compositions as compared with Serpent.
 
 \subsection{PWR Pincell}
+
+The PWR pincell problem is a single 2.4\% enriched PWR pincell from the BEAVRS
+benchmark~\citep{horelik2013mc,horelik2018mit}. Material compositions and
+dimensions from the BEAVRS benchmark were used without modification. The pincell
+extends to positive/negative infinity in the $z$ direction and has reflective
+boundary conditions in the $x$ and $y$ directions. A linear heating rate of
+\SI{174}{\watt\per\centi\meter} was used for power normalization. The pin was
+depleted to 50 MWd/kg using the following burnup schedule:
+\begin{itemize}
+  \item A single 0.1 MWd/kg, 0.4 MWd/kg, and 0.5 MWd/kg timesteps;
+  \item 1.0 MWd/kg timesteps out to a cumulative burnup of 10.0 MWd/kg;
+  \item 2.5 MWd/kg timesteps out to a cumulative burnup of 50.0 MWd/kg.
+\end{itemize}
+The short timesteps at the beginning are used to accurately capture the buildup
+of xenon isotopes. In both codes, the predictor method was used for time
+integration. While the predictor method is the least accurate time integration
+method, the purpose of the comparison is to confirm that OpenMC and Serpent
+produce comparable results, irrespective of the absolute values.
+
+Six coupled transport-depletion simulations were carried out with OpenMC to test
+a variety of combinations of depletion chains and methods for handling FPY
+energy dependence:
+\begin{table}
+
+\end{table}
+\begin{enumerate}
+  \item Full depletion chain, thermal FPYs
+  \item Full depletion chain, interpolated FPYs
+  \item Reduced depletion chain, thermal FPYs
+  \item CASL depletion chain, thermal FPYs
+  \item CASL depletion chain, FPYs based on cutoff method
+  \item CASL depletion chain, interpolated FPYs
+\end{enumerate}
+Because Serpent doesn't have an option for changing the depletion chain or the
+method for handling FPY energy dependence, a single coupled transport-depletion
+simulation with Serpent was carried out. For all simulations, 20 inactive
+batches, 200 active batches, and \num{3.6e6} particles per batch were used,
+which resulted in an uncertainty on the effective multiplication factor of about
+3--4 pcm.
+
+\Cref{fig:pwr-keff} shows the difference between $k_\text{eff}$ calculated by
+OpenMC and Serpent. The blue points correspond to results using the full
+depletion chain, the red points correspond to points using the reduced chain,
+and the green points correspond to results using the CASL chain. For each chain,
+the use of interpolated FPYs is denoted by circles and thermal FPYs by x's. At
+zero days, the difference in $k_\text{eff}$ between OpenMC and Serpent is not
+statistically significant. As the burnup increases, the difference for the full
+chain cases rises to about 20 pcm. The reduced chain case exhibits the exact
+same behavior as the full chain as one would expect. Use of the CASL chain
+results in a slightly larger difference between OpenMC and Serpent at 50 MWd/kg,
+about 40 pcm. We also see that for both the full and CASL chains, changing the
+method for handling FPY energy dependence doesn't appear to have much of an
+affect on $k_\text{eff}$, which is primarily driven by the prediction of
+actinide concentrations.
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=4in]{figures/pwr_keff.pdf}
+  \caption{Difference in the effective multiplication factor, $k_\text{eff}$,
+  between OpenMC and Serpent for the PWR pincell model. While not shown, the one
+  standard deviation uncertainty in the difference is about 3.5 pcm.}
+  \label{fig:pwr-keff}
+\end{figure}
+
+\Cref{fig:pwr-actinides-full} shows the relative difference in predicted
+actinide concentrations between OpenMC and Serpent at four different burnup
+values. We see that most of the differences are less than 0.1\%, and all are
+within 0.15\%. The differences in $^{235}$U and $^{239}$Pu concentrations
+increase slightly with increasing burnup, which most likely accounts for the 20
+pcm difference observed for the full chain in \cref{fig:pwr-keff}. While not
+shown, the predicted actinide concentrations for the reduced chain are nearly
+identical to those observed for the full chain in \cref{fig:pwr-actinides-full}.
+\Cref{fig:pwr-actinides-casl} shows the differences in actinide concentrations
+using the CASL depletion chain. Because the CASL depletion chain differs from
+the full depletion chain primarily with respect to how fission products are
+treated, the differences reported in \cref{fig:pwr-actinides-casl} closely
+mirror those in \cref{fig:pwr-actinides-full}.
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=4in]{figures/pwr/pwr_actinides_full.pdf}
+  \caption{Relative difference in actinide concetrations between OpenMC and
+  Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the full
+  depletion chain.}
+  \label{fig:pwr-actinides-full}
+\end{figure}
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=4in]{figures/pwr/pwr_actinides_casl.pdf}
+  \caption{Relative difference in actinide concetrations between OpenMC and
+  Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the CASL
+  depletion chain.}
+  \label{fig:pwr-actinides-casl}
+\end{figure}
+
+The predicted concentration of fission products depends on both the depletion
+chain that is used and the method for handling the energy dependence of the FPY.
+We begin with results for the full depletion chain and interpolated FPYs, which
+most closely matches the method used in Serpent; \cref{fig:pwr-fp-full-average}
+shows the relative difference in fission product concentrations between OpenMC
+and Serpent for this case. Most of the differences are less than 0.2\%, and only
+a few fission products show differences up to 0.5\%.
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=4in]{figures/pwr/pwr_fp_full_average.pdf}
+  \caption{Relative difference in fission product concetrations between OpenMC
+  and Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the
+  full depletion chain and interpolated FPYs.}
+  \label{fig:pwr-fp-full-average}
+\end{figure}
+
+\Cref{fig:pwr-fp-full-thermal} shows the predicted fission product concetrations
+when using the full depletion chain along with thermal FPYs. In this case,
+although most fission products are within 1\% of Serpent, they are
+systematically overpredicted other than a few nuclides. Because Serpent uses the
+average fission energy to interpolate between two sets of FPYs, the systematic
+difference is actually expected when using thermal FPYs in OpenMC. As the
+incident energy of a neutron causing fission increases, the ``saddle'' and
+``wings'' in the FPY curve tend to increase significantly, and the peaks around
+$A=95$ and $A=135$ tend to slightly decrease. This corresponds to exactly what
+is observed in \cref{fig:pwr-fp-full-thermal}. Because thermal FPYs were used in
+OpenMC for this case, we expect that fission products near one of the two peaks
+of the FPY curve will be slightly overpredicted relative to Serpent, and fission
+products away from the peak will be underpredicted. This can be confirmed for
+particular nuclides. Take, for example, $^{109}$Ag; in the $A=109$ mass chain,
+the nuclides with the highest independent yields are $^{109}$Tc and $^{109}$Mo.
+Both these nuclides have independent yields that increase significantly with
+increasing neutron energy. The same situation occurs for $^{129}$I, whose
+predominant precursor is $^{129}$Sn. The independent yields for these three
+precursor nuclides are shown in \cref{tab:fp-yields}.
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=4in]{figures/pwr/pwr_fp_full_thermal.pdf}
+  \caption{Relative difference in fission product concetrations between OpenMC
+  and Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the
+  full depletion chain and thermal FPYs.}
+  \label{fig:pwr-fp-full-thermal}
+\end{figure}
+\begin{table}[htbp]
+  \centering
+  \caption{Independent yields of select fission products due to fission in
+  $^{235}$U (data from the ENDF/B-VII.1 FPY sublibrary).}
+  \label{tab:fp-yields}
+  \begin{tabular}{cccc}
+    \toprule
+    \textbf{Energy Spectrum} & \textbf{$^{109}$Tc} & \textbf{$^{109}$Mo} & \textbf{$^{129}$Sn} \\
+    \midrule
+    Thermal & 0.00013 & 0.00015 & 0.00230 \\
+    Fast & 0.00034 & 0.00042 & 0.00550 \\
+    Fusion & 0.00598 & 0.00109 & 0.01162 \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+\Cref{fig:pwr-fp-casl-average} shows the relative difference in predicted
+fission product concentrations between OpenMC and Serpent when using the CASL
+depletion chain and interpolated FPYs. As in \cref{fig:pwr-fp-full-average},
+almost all fission products are predicted to well within 1\% at 50 MWd/kg.
+However, in this case there is one clear outlier: $^{85}$Kr. Upon further
+investigation, it was determined that the root cause for this difference is the
+fact that the CASL depletion chain uses a cumulative yield for $^{85}$Kr, and
+the cumulative yield of $^{85}$Kr is known to be inconsistent with the
+independent yields of the $A=85$ mass chain~\cite{pigni2015nds}. Serpent relies
+solely on the independent yield in its depletion solver (as does the full
+depletion chain). Therefore, there is nothing inherently wrong with the CASL
+depletion chain or OpenMC; it is simply a matter of the underlying nuclear data
+being inconsistent. The 5\% underprediction of $^{85}$Kr is also likely
+responsible for the slightly higher difference in $k_\text{eff}$ for the CASL
+depletion chain shown in \cref{fig:pwr-keff}.
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=4in]{figures/pwr/pwr_fp_casl_average.pdf}
+  \caption{Relative difference in fission product concetrations between OpenMC
+  and Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the
+  CASL depletion chain and interpolated FPYs.}
+  \label{fig:pwr-fp-casl-average}
+\end{figure}
+
+\Cref{fig:pwr-fp-casl-thermal} shows the relative difference in predicted in
+fission product concentrations when using the CASL depletion chain along with
+thermal FPYs. The results are quite close to those shown
+\cref{fig:pwr-fp-full-thermal}, with the exception of $^{85}$Kr for the
+aforementioned reasons.
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=4in]{figures/pwr/pwr_fp_casl_thermal.pdf}
+  \caption{Relative difference in fission product concetrations between OpenMC
+  and Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the
+  CASL depletion chain and thermal FPYs.}
+  \label{fig:pwr-fp-casl-thermal}
+\end{figure}
 

### 2020-06-23 figures/plot_barcharts.py

@@ -1,0 +1,142 @@
+import matplotlib.pyplot as plt
+from matplotlib import rcParams
+import matplotlib.ticker as mtick
+import numpy as np
+from openmc.data import ATOMIC_SYMBOL, zam
+from openmc.deplete import ResultsList
+import serpentTools
+
+# customizations
+rcParams['axes.autolimit_mode'] = 'round_numbers'
+rcParams['axes.labelsize'] = 'large'
+rcParams['axes.xmargin'] = 0
+rcParams['axes.ymargin'] = 0
+rcParams['axes.axisbelow'] = True
+rcParams['font.family'] = 'serif'
+rcParams['pdf.use14corefonts'] = True
+rcParams['savefig.bbox'] = 'tight'
+rcParams['font.size'] = 12.0
+rcParams['text.usetex'] = True
+rcParams['text.latex.preamble'] = [
+    r'\usepackage{amsmath}',
+]
+
+print('Loading PWR data...')
+results_full_thermal = ResultsList.from_hdf5("pwr/depletion_results_full_thermal.h5")
+results_full_cutoff = ResultsList.from_hdf5("pwr/depletion_results_full_cutoff.h5")
+results_full_average = ResultsList.from_hdf5("pwr/depletion_results_full_average.h5")
+results_reduced = ResultsList.from_hdf5("pwr/depletion_results_reduced_average.h5")
+results_casl_thermal = ResultsList.from_hdf5("pwr/depletion_results_casl_thermal.h5")
+results_casl_cutoff = ResultsList.from_hdf5("pwr/depletion_results_casl_cutoff.h5")
+results_casl_average = ResultsList.from_hdf5("pwr/depletion_results_casl_average.h5")
+fuel = serpentTools.read('pwr/serpent_input_dep.m')['fuel']
+
+print('Loading SFR data...')
+results_sfr_full = ResultsList.from_hdf5('sfr/depletion_results_full_average.h5')
+results_sfr_casl = ResultsList.from_hdf5('sfr/depletion_results_casl_average.h5')
+results_sfr_full_nop = ResultsList.from_hdf5('sfr/no_ptables/depletion_results_full_average.h5')
[…]

### 2020-06-23 figures/plot_keff.py

@@ -1,0 +1,71 @@
+import matplotlib.pyplot as plt
+from matplotlib import rcParams
+import numpy as np
+from openmc.deplete import ResultsList
+import serpentTools
+from uncertainties import unumpy as unp
+
+# customizations
+rcParams['axes.labelsize'] = 'large'
+rcParams['axes.axisbelow'] = True
+rcParams['font.family'] = 'serif'
+rcParams['pdf.use14corefonts'] = True
+rcParams['savefig.bbox'] = 'tight'
+rcParams['font.size'] = 12.0
+rcParams['text.usetex'] = True
+rcParams['text.latex.preamble'] = [
+    r'\usepackage{amsmath}',
+]
+
+# Read Serpent results --------------------------------------------------------
+
+def get_serpent_keff(filename):
+    res = serpentTools.read(filename)
+    return unp.uarray(*res.resdata['absKeff'].T)
+
+serpent_pwr = get_serpent_keff('pwr/serpent_input_res.m')
+serpent_sfr = get_serpent_keff('sfr/serpent_met1000_2d_res.m')
+serpent_sfr_nop = get_serpent_keff('sfr/no_ptables/serpent_met1000_2d_res.m')
+
+# Read OpenMC results ---------------------------------------------------------
+
+results_full_thermal = ResultsList.from_hdf5("pwr/depletion_results_full_thermal.h5")
+results_full_average = ResultsList.from_hdf5("pwr/depletion_results_full_average.h5")
+results_reduced = ResultsList.from_hdf5("pwr/depletion_results_reduced_average.h5")
+results_casl_thermal = ResultsList.from_hdf5("pwr/depletion_results_casl_thermal.h5")
+results_casl_cutoff = ResultsList.from_hdf5("pwr/depletion_results_casl_cutoff.h5")
+results_casl_average = ResultsList.from_hdf5("pwr/depletion_results_casl_average.h5")
+
+results_sfr_full = ResultsList.from_hdf5('sfr/depletion_results_full_average.h5')
+results_sfr_casl = ResultsList.from_hdf5('sfr/depletion_results_casl_average.h5')
+results_sfr_full_nop = ResultsList.from_hdf5('sfr/no_ptables/depletion_results_full_average.h5')
[…]

### 2020-06-23 manuscript.tex
Add discussion of SFR results

@@
 
+TODO: Discuss missing FPY~\cite{fensin2015pne}.
+
 Using all incident neutron, decay, and fission product yield sublibrary files
@@
 energy dependence:
-\begin{table}
-
-\end{table}
 \begin{enumerate}
@@
 simulation with Serpent was carried out. For all simulations, 20 inactive
-batches, 200 active batches, and \num{3.6e6} particles per batch were used,
+batches, 200 active batches, and \num{4.0e6} particles per batch were used,
 which resulted in an uncertainty on the effective multiplication factor of about
@@
 the cumulative yield of $^{85}$Kr is known to be inconsistent with the
-independent yields of the $A=85$ mass chain~\cite{pigni2015nds}. Serpent relies
+independent yields of the $A=85$ mass chain~\citep{pigni2015nds}. Serpent relies
 solely on the independent yield in its depletion solver (as does the full
@@
 
+The SFR assembly model is based on an OECD/NEA benchmark for SFR
+analysis~\citep{nsc2015}. Namely, a 2D model of a driver assembly for the 1000
+MWth metallic core (MET-1000) was used. The fuel composition corresponded to the
+lowest axial region (the 17.16 cm column in Table 2.23 from ~\citep{nsc2015}).
+All dimensions and material compositions can be found in the original OECD/NEA
+report. The assembly model extends to positive/negative infinity in the $z$
+direction and has periodic boundary conditions on the outer hexagonal surfaces.
+A linear heating rate of \SI{23.3}{\kilo\watt\per\meter} per
+pin~\citep{cahalan2007anl} was used for power normalization. The assembly was
+depleted for 360 days in 20 day timesteps. As in the PWR pincell case, the
+predictor method was used in both codes for time integration.
+
+Unlike the PWR pincell model, which only contains materials at a single
+temperature, the SFR assembly model uses a higher temperature in the fuel than
+in other materials. In OpenMC, the temperature-dependence of neutron cross
+sections was handled by interpolating data between neighboring temperatures.
+Serpent uses an explicit temperature treatment known as target motion
+sampling~\cite{viitanen2012nse}. However, we note that for this problem, the
+temperature dependence of unresolved resonance probability tables becomes
+important because of the fast flux spectrum, but the temperature treatment in
+Serpent 2.1.31 does not handle probability tables.
+
+Four coupled transport-depletion simulations were carried out with OpenMC for
+the SFR assembly model:
+\begin{enumerate}
+  \item Full depletion chain, interpolated FPYs, probability tables on
+  \item CASL depletion chain, interpolated FPYs, probability tables on
+  \item Full depletion chain, interpolated FPYs, probability tables off
+  \item CASL depletion chain, interpolated FPYs, probability tables off
+\end{enumerate}
+The results for the PWR pincell model indicated that the reduced depletion chain
+results in identical behavior to the full depletion chain, so we have not
+included a case for the SFR assembly using the reduced depletion chain. Two
+coupled transport-depletion simulation with Serpent was carried out, one with
+probabilty tables turned on and one with them turned off. For all simulations,
+20 inactive batches, 200 active batches, and \num{4.0e6} particles per batch
+were used, which resulted in an uncertainty on the effective multiplication
+factor of about 3--4 pcm.
+
+% Also, we
+% have only included a single case with thermal FPYs to compare with the use of
+% interpolated FPYs using the same depletion chain.
+
+\Cref{fig:sfr-keff} shows the difference between $k_\text{eff}$ over time
+calculated by OpenMC and Serpent. The blue points correspond to results using
+the full depletion chain, and the green points correspond to results using the
+CASL chain. For each chain, simulations with probability tables on are denoted
+by circles and simulations without probability tables are denoted by squares.
+Unlike in the PWR pincell case, there is a statistically significant difference
+in $k_\text{eff}$ at zero days. Let us first consider the case with probability
+tables turned off. In this case, $k_\text{eff}$ calculated by OpenMC is about 20
+pcm higher than that of Serpent. When using the full depletion chain, there is
+no change in the difference between OpenMC and Serpent. When using the CASL
+depletion chain, the difference increases by about 20--30 pcm, similar to the
+behavior observed in \cref{fig:pwr-keff}. With probability tables turned on,
+however, the situation is quite different. At zero days, $k_\text{eff}$
+calculated by OpenMC is 20 pcm lower than the value calculated by Serpent. As
+the burnup increases, there is an increasing positive bias between OpenMC and
+Serpent. To understand the cause of this difference, it is instructive to look
+at the differences in isotopic compositions.
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=4in]{figures/sfr_keff.pdf}
+  \caption{Difference in the effective multiplication factor, $k_\text{eff}$,
+  between OpenMC and Serpent for the SFR assembly model. While not shown, the one
+  standard deviation uncertainty in the difference is about 3.5 pcm.}
+  \label{fig:sfr-keff}
+\end{figure}
+
+\Cref{fig:sfr-actinides} shows the relative difference in predicted actinide
+concentrations between OpenMC and Serpent with probability tables turned on. For
+both the full and CASL depletion chains, the relative differences are all within
+0.2\%. While this may seem close, the observed difference in $^{238}$U and
+$^{239}$Pu concentrations merits special mention. In this model, the fuel
+material is at \SI{807}{\kelvin}, but the endf71x library contains cross
+sections processed at \SI{600}{\kelvin} and \SI{900}{\kelvin}. Thus, Serpent
+ends up using probability tables at \SI{600}{\kelvin} rather than
+\SI{807}{\kelvin}, which results in less parasitic absorption in $^{238}$U and
+consequently less production of $^{239}$Pu (hence a lower $k_\text{eff}$ than
+OpenMC with increasing burnup). Although OpenMC also does not have access to
+probability tables at the exact fuel temperature, it statistically interpolates
+between the probability tables at \SI{600}{\kelvin} and \SI{900}{\kelvin}.
+\begin{figure}[H]
+  \centering
+  \begin{subfigure}[t]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{figures/sfr/sfr_actinides_full.pdf}
+    \caption{Full depletion chain}
+    \label{fig:sfr-actinides-full}
+  \end{subfigure}
+  \hspace{0.05\textwidth}
+  \begin{subfigure}[t]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{figures/sfr/sfr_actinides_casl.pdf}
+    \caption{CASL depletion chain}
+    \label{fig:sfr-actinides-casl}
+  \end{subfigure}
+  \caption{Relative difference in actinide concetrations between OpenMC and
+  Serpent for the SFR assembly model at 80, 180, 260, and 360 days with
+  probability tables on.}
+  \label{fig:sfr-actinides}
+\end{figure}
+
+When probability tables are turned off, very close agreement is observed between
+the predicted actinide concentrations in OpenMC and Serpent, as shown in
+\cref{fig:sfr-actinides-nop}. In this case, the actinide concetrations agree to
+within 0.01\% at all times during depletion for both the full and CASL depletion
+chains.
+\begin{figure}[H]
+  \centering
+  \begin{subfigure}[t]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{figures/sfr/sfr_actinides_full_nop.pdf}
+    \caption{Full depletion chain}
+    \label{fig:sfr-actinides-full-nop}
+  \end{subfigure}
+  \hspace{0.05\textwidth}
+  \begin{subfigure}[t]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{figures/sfr/sfr_actinides_casl_nop.pdf}
+    \caption{CASL depletion chain}
+    \label{fig:sfr-actinides-casl-nop}
+  \end{subfigure}
+  \caption{Relative difference in actinide concetrations between OpenMC and
+  Serpent for the SFR assembly model at 80, 180, 260, and 360 days with
+  probability tables off.}
+  \label{fig:sfr-actinides-nop}
+\end{figure}
+
+\Cref{fig:sfr-fp} shows the relative difference in fission product
+concentrations between OpenMC and Serpent with probability tables turned on.
+Most differences between OpenMC and Serpent are below 1\%. As with the PWR
+pincell model, OpenMC tends to produce a lower concentration of $^{109}$Ag most
+likely because of differences in the treatment of FPY energy dependence. The
+only noticeable difference between the full and CASL depletion chains is for
+$^{85}$Kr, which is again explained by the use of an inconsistent cumulative
+yield in the CASL depletion chain.
+\begin{figure}[H]
+  \centering
+  \begin{subfigure}[t]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{figures/sfr/sfr_fp_full_average.pdf}
+    \caption{Full depletion chain}
+  \end{subfigure}
+  \hspace{0.05\textwidth}
+  \begin{subfigure}[t]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{figures/sfr/sfr_fp_casl_average.pdf}
+    \caption{CASL depletion chain}
+  \end{subfigure}
+  \caption{Relative difference in fission product concetrations between OpenMC
+  and Serpent for the SFR assembly model at 80, 180, 260, and 360 days with
+  probability tables on.}
+  \label{fig:sfr-fp}
+\end{figure}
+
+With probability tables turned on, the differences between OpenMC and Serpent in
+fission product concentrations are mostly the same as when probability tables
+were turned off, as shown in \cref{fig:sfr-fp-nop}. We note that because the
+difference in $^{238}$U and $^{239}$Pu was still less than 0.2\% in
+\cref{fig:sfr-actinides}, it is expected that the use of probability tables
+doesn't have a large effect on the fission product concentrations. Again, most
+fission product concentrations agree to within 1\%, with $^{109}$Ag and
+$^{85}$Kr being the notable outliers.
+\begin{figure}[H]
+  \centering
+  \begin{subfigure}[t]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{figures/sfr/sfr_fp_full_average_nop.pdf}
+    \caption{Full depletion chain}
+  \end{subfigure}
+  \hspace{0.05\textwidth}
+  \begin{subfigure}[t]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{figures/sfr/sfr_fp_casl_average_nop.pdf}
+    \caption{CASL depletion chain}
+  \end{subfigure}
+  \caption{Relative difference in fission product concetrations between OpenMC
+  and Serpent for the SFR assembly model at 80, 180, 260, and 360 days with
+  probability tables off.}
+  \label{fig:sfr-fp-nop}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{sec:conclusions}
+
+Discuss temperature dependence and probability tables in SFR model.
 

### 2020-06-23 figures/plot_barcharts.py

@@ -73,7 +73,7 @@ burnup_pwr = [
     12.5, 15.0, 17.5, 20.0, 22.5, 25.0, 27.5, 30.0, 32.5, 35.0, 37.5,
     40.0, 42.5, 45.0, 50.0
 ]
-burnup_sfr = np.linspace(0.0, 360.0, 73)
+burnup_sfr = np.linspace(0.0, 360.0, 19)
 
 def isotope_bar_plot(results_openmc, results_serpent, nuclides, filename, decimals=2, **kwargs):
     print(f'{filename}...')
@@ -93,9 +93,11 @@ def isotope_bar_plot(results_openmc, results_serpent, nuclides, filename, decima
     if 'pwr' in filename:
         burn_indices = (7, 14, 21, 28)
         burnup = burnup_pwr
+        burnup_units = 'MWd/kg'
     else:
         burn_indices = (4, 9, 13, 18)
         burnup = burnup_sfr
+        burnup_units = 'days'
     ind = np.arange(len(nuclides))
     height = 1/(len(burn_indices)+1)
     for i, index in enumerate(burn_indices):
@@ -104,7 +106,7 @@ def isotope_bar_plot(results_openmc, results_serpent, nuclides, filename, decima
         diff = np.zeros_like(data_openmc)
         nonzero = data_serpent > 0.0
         diff[nonzero] = (data_openmc[nonzero] - data_serpent[nonzero])/data_serpent[nonzero]
-        ax.barh(ind + i*height, diff, height, label=f'{burnup[index]} MWd/kg')
+        ax.barh(ind + i*height, diff, height, label=f'{burnup[index]} {burnup_units}')
 
     locs = ind + height*(len(burn_indices)/2 - 0.5)
     ax.xaxis.set_major_formatter(mtick.PercentFormatter(1.0, decimals))
@@ -124,19 +126,19 @@ isotope_bar_plot(results_casl_average, fuel, actinides, f'pwr/pwr_actinides_casl
 
 isotope_bar_plot(results_sfr_full, fuel_sfr, actinides, f'sfr/sfr_actinides_full.{ext}')
 isotope_bar_plot(results_sfr_casl, fuel_sfr, actinides, f'sfr/sfr_actinides_casl.{ext}')
[…]

### 2020-06-23 figures/plot_keff.py

@@ -48,7 +48,7 @@ def difference(results, serpent_keff):
     diff = 1e5*(openmc_keff - serpent_keff)
     return unp.nominal_values(diff)
 
-# Plot results -----------------------------------------------------------------
+# Plot PWR results -------------------------------------------------------------
 
 burnup = [
     0.0, 0.1, 0.5, 1.0, 2.0, 3.0, 4.0, 5.0, 6.0, 7.0, 8.0, 9.0, 10.0, 11.0,
@@ -68,4 +68,21 @@ ax.set_ylabel(r"$k_\text{OpenMC} - k_\text{Serpent}$ [pcm]")
 ax.grid(True)
 ax.legend()
 plt.savefig('pwr_keff.pdf', bbox_inches='tight')
+plt.close()
+
+# Plot SFR results -------------------------------------------------------------
+
+days = np.linspace(0.0, 360.0, 19)
+
+fig, ax = plt.subplots()
+ax.plot(days, difference(results_sfr_full, serpent_sfr), 'bo', label='Full, ptables')
+ax.plot(days, difference(results_sfr_casl, serpent_sfr), 'go', label='CASL, ptables')
+ax.plot(days, difference(results_sfr_full_nop, serpent_sfr_nop), 'bs', markerfacecolor='none', label='Full, no ptables')
+ax.plot(days, difference(results_sfr_casl_nop, serpent_sfr_nop), 'gs', markerfacecolor='none', label='CASL, no ptables')
+ax.plot(color='k', linestyle='--')
+ax.set_xlabel("Time [days]")
+ax.set_ylabel(r"$k_\text{OpenMC} - k_\text{Serpent}$ [pcm]")
+ax.grid(True)
+ax.legend()
+plt.savefig('sfr_keff.pdf', bbox_inches='tight')
 plt.close()

### 2020-06-23 manuscript.tex
Add conclusions and acknowledgments

@@
 
-\subsubsection{predictor--Corrector Methods}
+\subsubsection{Predictor--Corrector Methods}
 
@@
 
-Discuss temperature dependence and probability tables in SFR model.
+A new depletion solver has been implemented in the OpenMC Monte Carlo particle
+transport code. The depletion solver is implemented primarily in Python 3 and
+interfaces with OpenMC's transport solver through Python bindings to an
+underlying C++ API, which enables data transfers to happen in memory rather than
+through filesystem I/O. The design of the depletion solver relies on a set of
+abstract classes that provide modularity and separation between difference
+components, including the time integration methods, the matrix exponential
+solver, and the transport operator. In principle, this modularity would allow
+the depletion system to be used with another transport code. Rather than relying
+on a single time integration method, OpenMC implements several time integration
+schemes including the predictor method, standard predictor-corrector methods,
+stochastic implicit Euler integrators, and several novel
+methods~\citep{josey2017phd, josey2017jcp}. To evaluate the matrix exponential,
+OpenMC relies on the incomplete partial fraction form of CRAM by default.
+
+To assess the accuracy of the new depletion solver, comparisons were made
+between OpenMC and Serpent on a PWR pincell model and an SFR assembly model.
+When the full depletion chain was used in OpenMC, the predictions of
+$k_\text{eff}$ agreed to within 20--30 pcm of that calculated by Serpent at all
+times in life for both the PWR and SFR model. Actinide concentrations were found
+to agree to a fraction of a percent, and most fission product concentrations
+were found to agree within 1\%. The few cases where fission product
+concentrations differed by a larger amount were found to be related to
+differences in the treatment of the energy dependence of FPYs. It is our sense
+that there is no consensus in the community as to what is the ``correct'' way of
+handling this energy dependence, so it is not possible to opine on whether one
+set of results is right or wrong. When the simplified CASL depletion chain was
+used in OpenMC, the accuracy of actinide and fission product concentrations did
+not change appreciably. The only nuclide for which a significant difference was
+observed was $^{85}$Kr; this difference can be explained by a cumulative yield
+in the ENDF/B-VII.1 FPY sublibrary that is inconsistent with the independent
+yields of precursors~\citep{pigni2015nds}. Because the CASL depletion chain
+results in a significantly lower runtime compared to the full depletion chain
+while still retaining accuracy, it is recommended for general use by OpenMC
+users.
+
+We found that when comparing results on the SFR assembly model, larger
+differences in $k_\text{eff}$ and isotopic concentrations were observed when
+unresolved resonance region probability tables were turned on. These differences
+were found to be related to how the temperature dependence of cross sections is
+treated in the unresolved resonance region. In OpenMC, when the temperature of a
+material is in between two temperatures at which probability tables are
+tabulated, statistical interpolation between the tables is performed. Serpent,
+on the other hand, uses probability tables at the lower tabulated temperature.
+For the SFR model, this results in lower parasitic neutron absorption in
+$^{238}$U, lower production of $^{239}$Pu, and ultimately a lower
+$k_\text{eff}$. After depleting for 360 days, there was a 100 pcm difference
+between OpenMC and Serpent due to this effect. Although this is not a huge
+difference, users who are simulating problems with a non-uniform temperature
+distribution should be aware of this behavior.
+
+Finally, we remark that although we did observe some differences between OpenMC
+and Serpent, most of the differences were related to deficiencies in the nuclear
+data itself. As a whole, the results demonstrate that when the same nuclear data
+and same methods are used in two independently developed codes, it is possible
+to obtain results in a depletion simulation that match within 10s of pcm on
+$k_\text{eff}$ and less than 1\% on isotopic concentrations.
 
@@
 \section*{Acknowledgments}
+
+We gratefully acknowledge the OpenMC user community for providing early feedback
+and bug reports related to depletion that have helped improve the code. We wish
+to acknowledge Jaakko Lepp{\"{a}}nen (VTT) for providing the developmental
+version of Serpent 2.1.32 used in this work and for helping to shed light on the
+methods used in Serpent. We acknowledge Nicolas Stauff (ANL) for providing
+recommendations on a suitable SFR target problem.
 
@@
 contract DE-AC02-06CH11357.
+
+We gratefully acknowledge the computing resources provided on Bebop, a
+high-performance computing cluster operated by the Laboratory Computing Resource
+Center at Argonne National Laboratory. This research made use of the resources
+of the High Performance Computing Center at Idaho National Laboratory, which is
+supported by the Office of Nuclear Energy of the U.S. Department of Energy and
+the Nuclear Science User Facilities under Contract No. DE-AC07-05ID14517.
 

### 2020-06-23 manuscript.tex
Add missing details, add abstract

@@
 \begin{abstract}
-
+  A depletion solver has been implemented in OpenMC and is described herein. The
+  depletion solver is implemented in Python and interfaces with OpenMC's
+  transport solver through a C++ API, which enables an in-memory
+  depletion--transport coupling. Multiple integration methods for advancing in
+  time have been implemented and exhibit tradeoffs in cost, accuracy, and memory
+  use. For all time integration methods, evaluation of the matrix exponential is
+  performed using the incomplete partial fraction form of the Chebyshev rational
+  approximation method.
+
+  Simulations of a PWR pincell and an SFR assembly were carried out with OpenMC
+  and Serpent. For both problems, the use of a high-fidelity depletion chain
+  results in predictions of $k_\text{eff}$ that agree within 20--30 pcm between
+  OpenMC and Serpent. Predicted actinide concentrations were found to agree to a
+  fraction of a percent, and most fission product concentrations were found to
+  agree within 1\%. The few cases where larger differences were observed can be
+  attributed either to differences in how the energy dependence of fission
+  product yields is handled or deficiencies in the nuclear data used. OpenMC
+  simulations of the PWR and SFR problems using a simplified 228 nuclide
+  depletion chain demonstrate that it achieves accuracy close to that of the
+  full, high-fidelity depletion chain.
 \end{abstract}
@@
 \begin{keyword}
-  Monte Carlo, depletion
+  Monte Carlo, particle transport, depletion, transmutation
 \end{keyword}
@@
 \emph{depletion chain} file, which includes fundamental data gathered from ENDF
-incident neutron, decay, and fission product yield sublibraries. For each
+incident neutron, decay, and fission product yield (FPY) sublibraries. For each
 nuclide, this file includes:
@@
 Depletion chains can be automatically created in the Python API in the following
-manner. First, the user supplies a list of ENDF incident neutron, decay, and
-fission product yield sublibrary files. The incident neutron files are used to
-determine what nuclear reactions result in transmutation and what the
-corresponding Q values of the reactions are. The decay files are used to
-determine the half-life and possible decay modes of nuclides that are unstable.
-The fission product yield sublibraries are used to determine what fission
-products are produced when a fission reaction occurs. This information is then
-aggregated and written to an XML file that serves as the depletion chain file
-that is used at runtime to construct a burnup matrix.
+manner. First, the user supplies a list of ENDF incident neutron, decay, and FPY
+sublibrary files. The incident neutron files are used to determine what nuclear
+reactions result in transmutation and what the corresponding Q values of the
+reactions are. The decay files are used to determine the half-life and possible
+decay modes of nuclides that are unstable. The FPY sublibraries are used to
+determine what fission products are produced when a fission reaction occurs.
+This information is then aggregated and written to an XML file that serves as
+the depletion chain file that is used at runtime to construct a burnup matrix.
 
@@
 
-TODO: Discuss missing FPY~\cite{fensin2015pne}.
-
-Using all incident neutron, decay, and fission product yield sublibrary files
-from an ENDF-format library, such as ENDF/B-VIII.0, results in a \emph{full}
-depletion chain that typically includes thousands of nuclides. Because the use
-of a chain with $n$ nuclides will result in an $n\times n$ burnup matrix, there
-is an incentive to reduce the number of nuclides in the depletion chain to
-minimize both the size of the burnup matrix and the number of nuclides that are
-tracked during a transport solve. The depletion chain class in OpenMC's Python
-API includes a method for automatically reducing the size of the chain. Starting
-with a set of nuclides that must be included, the method will recursively go
-through all decay and reaction modes to determine what products may be produced.
-Using this method on the ENDF/B-VII.1 library starting from a list of all stable
-or very long lived (half-life greater than $10^{15}$ sec) nuclides reduces the
-total number of nuclides in the depletion chain from 3,820 to 1,693.
+Some nuclides that have fission reactions in ENDF/B-VII.1, such as $^{243}$Pu,
+do not have corresponding FPYs. For these nuclides, a decision has to be made on
+what set of FPY should be used when fission occurs. This issue was briefly
+studied by \citet{fensin2015pne}, who found that the proportion of fission
+reactions that occur in these nuclides is usually small enough that the choice
+of FPYs is not of great importance. For the depletion chains used in OpenMC, we
+have adopted the method used in Serpent, which is as follows:
+\begin{enumerate}
+  \item If the nuclide missing FPYs is a metastable state, use FPYs from the
+  ground state;
+  \item Use FPYs from an isotone if available;
+  \item Use FPYs from $^{235}$U.
+\end{enumerate}
+
+Using all incident neutron, decay, and FPY sublibrary files from an ENDF-format
+library, such as ENDF/B-VIII.0, results in a \emph{full} depletion chain that
+typically includes thousands of nuclides. Because the use of a chain with $n$
+nuclides will result in an $n\times n$ burnup matrix, there is an incentive to
+reduce the number of nuclides in the depletion chain to minimize both the size
+of the burnup matrix and the number of nuclides that are tracked during a
+transport solve. The depletion chain class in OpenMC's Python API includes a
+method for automatically reducing the size of the chain. Starting with a set of
+nuclides that must be included, the method will recursively go through all decay
+and reaction modes to determine what products may be produced. Using this method
+on the ENDF/B-VII.1 library starting from a list of all stable or very long
+lived (half-life greater than $10^{15}$ sec) nuclides reduces the total number
+of nuclides in the depletion chain from 3,820 to 1,693.
 
 In addition to the full and reduced depletion chains, a simplified depletion
-chain based on the VERA depletion benchmark specification~\citep{kim2015casl}
-was created and made available for users at \url{https://openmc.org}. This chain
-includes 223 nuclides and has adjusted decay branching ratios for some nuclides
+chain based on the VERA depletion benchmark specification~\citep{kim2015casl},
+which will henceforth be referred to as the \emph{CASL} depletion chain, was
+created and made available for users at \url{https://openmc.org}. This chain
+includes 228 nuclides and has adjusted decay branching ratios for some nuclides
 to preserve the accuracy of nuclide production rates. \Cref{fig:chains} shows
-which nuclides are included in each of the depletion chains.
+which nuclides are included in each of the depletion chains. To accurately model
+fast reactor problems (which may include Zr and Mo in the fuel), we have
+included $^{90}$Zr, $^{92}$Zr, $^{94}$Zr, $^{92}$Mo, and $^{94}$Mo in this
+depletion chain even though they were not included in the original
+specification.
 \begin{figure}
@@
 To validate the depletion solver in OpenMC, we have carried out simulations
-using the \texttt{openmc.deplete} module as well as Serpent 2.1.31 on two simple
-problems. The first problem is a pressurized water reactor (PWR) pincell and the
-second problem is a single assembly of a sodium-cooled fast reactor (SFR). These
-problems were selected to cover both thermal and fast neutron spectra.
+using the \texttt{openmc.deplete} module as well as a developmental version of
+Serpent 2.1.32 on two simple problems. The first problem is a pressurized water
+reactor (PWR) pincell and the second problem is a single assembly of a
+sodium-cooled fast reactor (SFR). These problems were selected to cover both
+thermal and fast neutron spectra. While we had originally attempted to use
+Serpent 2.1.31, early comparisons between OpenMC and this version of Serpent
+revealed significant differences in the predicted concentration of $^{97}$Mo.
+This was found to be due to $^{97\text{m}}$Y having a non-zero independent yield
+and a zero cumulative yield in the FPY sublibrary file for $^{235}$U fission in
+ENDF/B-VII.1. The zero cumulative yield caused Serpent 2.1.31 to ignore it
+altogether, which resulted in significantly less production of $^{97}$Mo. The
+developmental version of Serpent 2.1.32 no longer ignores this particular
+fission product despite the erroneous cumulative yield.
 
@@
   \item CASL depletion chain, thermal FPYs
-  \item CASL depletion chain, FPYs based on cutoff method
+  %\item CASL depletion chain, FPYs based on cutoff method
   \item CASL depletion chain, interpolated FPYs

### 2020-06-23 manuscript.tex
Changes from first read-through (need to expand conclusions)

@@
   depletion solver is implemented in Python and interfaces with OpenMC's
-  transport solver through a C++ API, which enables an in-memory
-  depletion--transport coupling. Multiple integration methods for advancing in
-  time have been implemented and exhibit tradeoffs in cost, accuracy, and memory
-  use. For all time integration methods, evaluation of the matrix exponential is
-  performed using the incomplete partial fraction form of the Chebyshev rational
-  approximation method.
+  transport solver through a C++ application programming interface, which
+  enables an in-memory transport--depletion coupling. Multiple integration
+  methods for advancing in time have been implemented and exhibit tradeoffs in
+  cost, accuracy, and memory use. For all time integration methods, evaluation
+  of the matrix exponential is performed using the incomplete partial fraction
+  form of the Chebyshev rational approximation method.
 
@@
 
-With continued advances in computing, Monte Carlo (MC) particle transport
-simulations are now widely used in reactor design and analysis in a variety of
-roles. The earliest efforts to couple MC particle transport with depletion date
-back to the 1990s~\citep{moore1995inel,trellue1998lanl}. These early efforts
-involved taking existing MC and depletion codes, such as
+Monte Carlo (MC) particle transport simulations are now widely used in reactor
+design and analysis in a variety of roles thanks to continued advances in
+computing. The earliest efforts to couple MC particle transport with depletion
+date back to the 1990s~\citep{moore1995inel,trellue1998lanl}. These early
+efforts involved taking existing MC and depletion codes, such as
 MCNP~\citep{goorley2012nt} and ORIGEN~\citep{croff1983nt}, and putting a
@@
 
-In the present work, we describe a new depletion solver in
-OpenMC~\citep{romano2015ane1}, an open source, community-developed MC particle
-transport code. To date, OpenMC has not had a built-in depletion solver. This
-has led to numerous efforts in the community toward developing coupled
-transport-depletion calculations with OpenMC~\citep{gul2017ane,
-lanversin2017icone, lanversin2019phd, liu2019nst, zhuang2020pne, zhao2020ned,
-zhao2020cpc, zhang2020ane}. While these efforts have proven the viability of
-performing such coupled calculations with OpenMC, none of them has resulted in a
-validated, openly-available, and generalized capability that OpenMC users can
-take advantage of. The capability that is described herein is part of the OpenMC
+In the present work, we describe a new depletion solver in OpenMC, an open
+source, community-developed MC particle transport code~\citep{romano2015ane1}.
+To date, OpenMC has not had a built-in depletion solver. This has led to
+numerous efforts in the community toward developing coupled transport--depletion
+calculations with OpenMC~\citep{gul2017ane, lanversin2017icone,
+lanversin2019phd, liu2019nst, zhuang2020pne, zhao2020ned, zhao2020cpc,
+zhang2020ane}. While these efforts have proven the viability of performing such
+coupled calculations with OpenMC, none of them has resulted in a validated,
+openly-available, and generalized capability that OpenMC users can take
+advantage of. The capability that is described herein is part of the OpenMC
 project and can be found in version 0.11~\citep{openmc-0110} of the code or
@@
 which material compositions change is sufficiently long such that the transport
-equation can be solved as a time-independent equation. Without loss of
-generality, let us write the burnup matrix solely as a function of $\vect{n}$
-so that \cref{eq:depletion-matrix-t} becomes
+equation can be solved as a steady state equation. Without loss of generality,
+let us write the burnup matrix solely as a function of $\vect{n}$ so that
+\cref{eq:depletion-matrix-t} becomes
 \begin{equation}
@@
 improved accuracy over the predictor method. The simplest of these methods is
-the CE/CM algorithm, which is the default algorithm used in
-MCNP6~\citep{fensin2006tans}. In this method, $\vect{A}$ is evaluated at the
-beginning of the step and used to deplete the composition over half the step.
-Then, $\vect{A}$ is reevaluated using the predicted middle-of-step composition
-and used to deplete over the full step. The algorithm can be expressed as
+the CE/CM (constant extrapolation, constant midpoint) algorithm, which is the
+default algorithm used in MCNP6~\citep{fensin2006tans}. In this method,
+$\vect{A}$ is evaluated at the beginning of the step and used to deplete the
+composition over half the step. Then, $\vect{A}$ is reevaluated using the
+predicted middle-of-step composition and used to deplete over the full step. The
+algorithm can be expressed as
 \begin{equation}
@@
 \end{equation}
-where $\vect{A}_1$ and $\vect{A}_2$ correspond $\hat{\vect{A}}$ evaluated at the
-nodes of a two-point Gauss-Legendre quadrature over the interval $[0,h]$:
+where $\vect{A}_1$ and $\vect{A}_2$ correspond to $\hat{\vect{A}}$ evaluated at
+the nodes of a two-point Gauss-Legendre quadrature over the interval $[0,h]$:
 \begin{equation}
@@
 requires two transport solutions and $m$ matrix exponentials. A comparison of
-these two approaches has shown that \cref{eq:celi-cfq4} achieves better accuracy
+these approaches has shown that \cref{eq:celi-cfq4} achieves better accuracy
 than the use of five substeps~\citep{josey2017phd}. OpenMC also includes an
@@
 \end{equation}
-This results in a non-linear equation for $\vect{n}_{i+1}$ that can't be solved
+This results in a non-linear equation for $\vect{n}_{i+1}$ that cannot be solved
 directly. However, it is possible to solve for $\vect{n}_{i+1}$ using a
 fixed-point iteration. \citet{dufek2013ane} utilized a stochastic approximation
-based on the Robbins-Monro algorithm~\citep{robbins1951ams} in order to
+based on the Robbins--Monro algorithm~\citep{robbins1951ams} in order to
 guarantee convergence of the fixed-point iteration while accounting for
@@
 approximation for CRAM requires solving $k/2$ sparse linear systems. OpenMC
-relies on functionality from scipy.sparse.linalg for solving the linear system.
+relies on functionality from scipy.sparse.linalg~\citep{virtanen2020nm} for
+solving the linear system.
 \begin{algorithm}[H]
@@
 In OpenMC, the burnup matrix is constructed based on data inside of a
-\emph{depletion chain} file, which includes fundamental data gathered from ENDF
+\emph{depletion chain} file that includes fundamental data gathered from ENDF
 incident neutron, decay, and fission product yield (FPY) sublibraries. For each
@@
 
-Depletion chains can be automatically created in the Python API in the following
-manner. First, the user supplies a list of ENDF incident neutron, decay, and FPY
-sublibrary files. The incident neutron files are used to determine what nuclear
-reactions result in transmutation and what the corresponding Q values of the
-reactions are. The decay files are used to determine the half-life and possible
-decay modes of nuclides that are unstable. The FPY sublibraries are used to
-determine what fission products are produced when a fission reaction occurs.
-This information is then aggregated and written to an XML file that serves as
-the depletion chain file that is used at runtime to construct a burnup matrix.
+Depletion chains can be automatically created in the Python application
+programming interface (API) in the following manner. First, the user supplies a
+list of ENDF incident neutron, decay, and FPY sublibrary files. The incident
+neutron files are used to determine what nuclear reactions result in
+transmutation and what the corresponding Q values of the reactions are. The
+decay files are used to determine the half-life and possible decay modes of
+nuclides that are unstable. The FPY sublibraries are used to determine what
+fission products are produced when a fission reaction occurs. This information
+is then aggregated and written to an XML file that serves as the depletion chain
+file that is used at runtime to construct a burnup matrix.
 
@@
 Some nuclides that have fission reactions in ENDF/B-VII.1, such as $^{243}$Pu,
-do not have corresponding FPYs. For these nuclides, a decision has to be made on
-what set of FPY should be used when fission occurs. This issue was briefly
-studied by \citet{fensin2015pne}, who found that the proportion of fission
-reactions that occur in these nuclides is usually small enough that the choice
-of FPYs is not of great importance. For the depletion chains used in OpenMC, we
-have adopted the method used in Serpent, which is as follows:
+do not have corresponding FPYs. For these nuclides, a decision must be made on
+what set of FPYs to use when fission occurs. This issue was briefly studied by
+\citet{fensin2015pne}, who found that the proportion of fission reactions that
+occur in these nuclides is usually small enough that the choice of FPYs is not
+of great importance. For the depletion chains used in OpenMC, we have adopted
+the method used in Serpent, which is as follows:
 \begin{enumerate}
@@
   \item Use FPYs from an isotone if available;
-  \item Use FPYs from $^{235}$U.
+  \item If all else fails, use FPYs from $^{235}$U.
 \end{enumerate}
@@
 
-Like capture branching ratios, fission product yields (FPYs) also depend on the
-incident neutron energy. ENDF fission product yield sublibraries typically
-include yields tabulated at two or three energies (thermal, fast, and fusion
-energies). Complicating matters is the fact that while these data are specified
-for a single energy, they actually represent the observed FPYs for a spectrum of
-energies (characterized by a single point). For example, FPYs are often given at
-\SI{0.0253}{\electronvolt}, but this actually represents the observed FPY under
-a thermal neutron spectrum that contains a continuum of energies. The ENDF-6
-formats manual~\citep{trkov2018bnl} recommends using linear-linear interpolation
-between energy points. However, at the end of a transport solve, we typically
-have a one-group or multigroup fission reaction rate---it's not clear how one
-should use this information to calculate an effective FPY.
+Like capture branching ratios, FPYs also depend on the incident neutron energy.
+ENDF FPY sublibraries typically include yields tabulated at two or three
+energies (thermal, fast, and fusion energies). Complicating matters is the fact
+that while these data are specified for a single energy, they actually represent
+the observed FPYs for a spectrum of energies (characterized by a single point).
+For example, FPYs are often given at \SI{0.0253}{\electronvolt}, but this
+actually represents the observed FPYs under a thermal neutron spectrum that
+contains a continuum of energies. The ENDF-6 formats manual~\citep{trkov2018bnl}
+recommends using linear-linear interpolation between energy points. However, at
+the end of a transport solve, we typically have a one-group or multigroup
+fission reaction rate---it's not clear how one should use this information to
+calculate an effective FPY.
 
@@
 Generally, to run OpenMC, a user writes a Python script that utilizes classes
-and functions defined as part of an application programming interface (API) to
-generate XML input files that are then read by the executable. On top of this,
-OpenMC is also built as a shared library with a C++ API that can be called by
-external programs. OpenMC's Python API includes bindings to the C++ API; this
-enables a single run to be broken up into individual components. For example, a
-function call from the Python bindings to the C++ API can initialize the solver
-(reading inputs, loading cross sections, etc.) without actually performing a
-simulation. A separate function call can run the transport simulation. Further
-function calls allow a power user to obtain information about tally results
-directly in memory from the C++ API.
+and functions defined as part of an API to generate XML input files that are
+then read by the executable. On top of this, OpenMC is also built as a shared
+library with a C++ API that can be called by external programs. OpenMC's Python
+API includes bindings to the C++ API; this enables a single run to be broken up
+into individual components. For example, a function call from the Python
+bindings to the C++ API can initialize the solver (reading inputs, loading cross
+sections, etc.) without actually performing a simulation. A separate function
+call can run the transport simulation. Further function calls allow a power user
+to obtain information about tally results directly in memory from the C++ API.
 
@@
 The abstract \texttt{Integrator} class represents a time integration method.
-This class effectively orchestrates the coupled transport-depletion simulation.
+This class effectively orchestrates the coupled transport--depletion simulation.
 Subclasses of \texttt{Integrator} are required to implement a method that takes
@@
 
-Six coupled transport-depletion simulations were carried out with OpenMC to test
+Five coupled transport--depletion simulations were carried out with OpenMC to test
 a variety of combinations of depletion chains and methods for handling FPY
@@
 Because Serpent doesn't have an option for changing the depletion chain or the
-method for handling FPY energy dependence, a single coupled transport-depletion
+method for handling FPY energy dependence, a single coupled transport--depletion
 simulation with Serpent was carried out. For all simulations, 20 inactive
@@
 
-Four coupled transport-depletion simulations were carried out with OpenMC for
+Four coupled transport--depletion simulations were carried out with OpenMC for
 the SFR assembly model:
@@
 included a case for the SFR assembly using the reduced depletion chain. Two
-coupled transport-depletion simulation with Serpent was carried out, one with
+coupled transport--depletion simulation with Serpent was carried out, one with
 probabilty tables turned on and one with them turned off. For all simulations,
@@
 on a single time integration method, OpenMC implements several time integration
-schemes including the predictor method, standard predictor-corrector methods,
+schemes including the predictor method, standard predictor--corrector methods,
 stochastic implicit Euler integrators, and several novel
@@
 
+Discuss limitations (not validated against experiment, only as good as data).
+Improvements in runtime are needed. Only tested for fission applications.
+Further development and testing needed for other application areas, such as
+transmutation in fusion reactors.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-06-24 manuscript.tex
More proofreading edits

@@
 \usepackage{algpseudocode}
-\usepackage{siunitx}
+\usepackage[per-mode=symbol]{siunitx}
 \usepackage{multirow}
@@
 thermal and fast neutron spectra. While we had originally attempted to use
-Serpent 2.1.31, early comparisons between OpenMC and this version of Serpent
-revealed significant differences in the predicted concentration of $^{97}$Mo.
-This was found to be due to $^{97\text{m}}$Y having a non-zero independent yield
-and a zero cumulative yield in the FPY sublibrary file for $^{235}$U fission in
-ENDF/B-VII.1. The zero cumulative yield caused Serpent 2.1.31 to ignore it
-altogether, which resulted in significantly less production of $^{97}$Mo. The
-developmental version of Serpent 2.1.32 no longer ignores this particular
-fission product despite the erroneous cumulative yield.
+Serpent 2.1.31, preliminary comparisons between OpenMC and this version of
+Serpent revealed significant differences in the predicted concentration of
+$^{97}$Mo. This difference was found to be due to $^{97\text{m}}$Y having a
+non-zero independent yield and a zero cumulative yield in the FPY sublibrary
+file for $^{235}$U fission in ENDF/B-VII.1. The zero cumulative yield caused
+Serpent 2.1.31 to ignore it altogether, which resulted in significantly less
+production of $^{97}$Mo. The developmental version of Serpent 2.1.32 no longer
+ignores this particular fission product despite the erroneous cumulative yield.
 
@@
 simulation, we must ensure that the incident neutron cross sections and
-secondary angle/energy distributions, radioactive decay data, and fission
-product yields are the same for both codes. In the simulations performed, all
-data was based on the ENDF/B-VII.1 nuclear data library~\citep{chadwick2011nds}.
-Incident neutron cross sections from the ENDF/B-VII.1 library processed by the
-LANL nuclear data team~\citep{conlin2013lanl}, known as endf71x, were used for
-transport simulations in both OpenMC and Serpent. Serpent is able to use the ACE
-files from the endf71x library directly while OpenMC uses an HDF5 format that
-requires conversion of the ACE files first. For the PWR pincell problem,
-$S(\alpha,\beta)$ data is also needed for hydrogen in light water; because
-Serpent does not yet handle the continuous representation used in the ENDF71SaB
-library, the light water data from ENDF/B-VII.0 was used instead. We note that
-the light water evaluation didn't actually change between ENDF/B-VII.0 and
-ENDF/B-VII.1.
+secondary angle/energy distributions, radioactive decay data, and FPYs are the
+same for both codes. In the simulations performed, all data was based on the
+ENDF/B-VII.1 nuclear data library~\citep{chadwick2011nds}. Incident neutron
+cross sections from the ENDF/B-VII.1 library processed by the LANL nuclear data
+team~\citep{conlin2013lanl}, known as endf71x, were used for transport
+simulations in both OpenMC and Serpent. Serpent is able to use the ACE files
+from the endf71x library directly while OpenMC uses an HDF5 format that requires
+conversion of the ACE files first. For the PWR pincell problem,
+$S(\alpha,\beta)$ thermal scattering data is also needed for hydrogen in light
+water; because Serpent does not yet handle the continuous representation used in
+the ENDF71SaB library, the light water data from ENDF/B-VII.0 was used instead.
+We note that the light water evaluation didn't actually change between
+ENDF/B-VII.0 and ENDF/B-VII.1.
 
@@
 manner, the default fission heating values from Serpent were used in OpenMC. By
-default, the energy release per fission in $^{235}U$ in Serpent is set to
-\SI{202.27}{\mega\electronvolt}, and the values for other actinides is scaled
+default, the energy release per fission in $^{235}$U in Serpent is set to
+\SI{202.27}{\mega\electronvolt}, and the values for other actinides are scaled
 based on the Q value for fission listed in MF=3 in the ENDF evaluation. Without
 the same heating values, the power normalization might result in OpenMC
-depleting materials too quickly or too slowly.
+depleting materials too quickly or too slowly relative to Serpent.
 
@@
 branching ratio treatment. Then, the energy-independent branching ratios were
-set in both Serpent and OpenMC. As discussed in \cref{sec:fpy}, OpenMC also has
-several methods for treating the energy dependence of FPYs. The OpenMC
-simulations were carried out with the three different methods to see what effect
-they have on resulting isotopic compositions as compared with Serpent.
+set in both Serpent and OpenMC.
 
@@
 method, the purpose of the comparison is to confirm that OpenMC and Serpent
-produce comparable results, irrespective of the absolute values.
+produce comparable results irrespective of the absolute values.
 
@@
   \item Full depletion chain, interpolated FPYs
-  \item Reduced depletion chain, thermal FPYs
+  \item Reduced depletion chain, interpolated FPYs
   \item CASL depletion chain, thermal FPYs
@@
 method for handling FPY energy dependence doesn't appear to have much of an
-affect on $k_\text{eff}$, which is primarily driven by the prediction of
+effect on $k_\text{eff}$, which is primarily driven by the prediction of
 actinide concentrations.
@@
 average fission energy to interpolate between two sets of FPYs, the systematic
-difference is actually expected when using thermal FPYs in OpenMC. As the
-incident energy of a neutron causing fission increases, the ``saddle'' and
-``wings'' in the FPY curve tend to increase significantly, and the peaks around
-$A=95$ and $A=135$ tend to slightly decrease. This corresponds to exactly what
-is observed in \cref{fig:pwr-fp-full-thermal}. Because thermal FPYs were used in
-OpenMC for this case, we expect that fission products near one of the two peaks
-of the FPY curve will be slightly overpredicted relative to Serpent, and fission
-products away from the peak will be underpredicted. This can be confirmed for
-particular nuclides. Take, for example, $^{109}$Ag; in the $A=109$ mass chain,
-the nuclides with the highest independent yields are $^{109}$Tc and $^{109}$Mo.
-Both these nuclides have independent yields that increase significantly with
-increasing neutron energy. The same situation occurs for $^{129}$I, whose
-predominant precursor is $^{129}$Sn. The independent yields for these three
-precursor nuclides are shown in \cref{tab:fp-yields}.
+difference is expected when using thermal FPYs in OpenMC. As the incident energy
+of a neutron causing fission increases, the ``saddle'' and ``wings'' in the FPY
+curve tend to increase significantly, and the peaks around $A=95$ and $A=135$
+tend to slightly decrease. This corresponds to exactly what is observed in
+\cref{fig:pwr-fp-full-thermal}. Because thermal FPYs were used in OpenMC for
+this case, we expect that fission products near one of the two peaks of the FPY
+curve will be slightly overpredicted relative to Serpent, and fission products
+away from the peak will be underpredicted. This can be confirmed for particular
+nuclides. Take, for example, $^{109}$Ag; in the $A=109$ mass chain, the nuclides
+with the highest independent yields are $^{109}$Tc and $^{109}$Mo. Both these
+nuclides have independent yields that increase significantly with increasing
+neutron energy. The same situation occurs for $^{129}$I, whose predominant
+precursor is $^{129}$Sn. The independent yields for these three precursor
+nuclides are shown in \cref{tab:fp-yields}.
 \begin{figure}[H]
@@
 the cumulative yield of $^{85}$Kr is known to be inconsistent with the
-independent yields of the $A=85$ mass chain~\citep{pigni2015nds}. Serpent relies
-solely on the independent yield in its depletion solver (as does the full
-depletion chain). Therefore, there is nothing inherently wrong with the CASL
-depletion chain or OpenMC; it is simply a matter of the underlying nuclear data
-being inconsistent. The 5\% underprediction of $^{85}$Kr is also likely
-responsible for the slightly higher difference in $k_\text{eff}$ for the CASL
-depletion chain shown in \cref{fig:pwr-keff}.
+independent yields of the $A=85$ mass chain in
+ENDF/B-VII.1~\citep{pigni2015nds}. Serpent relies solely on independent yields
+in its depletion solver (as does the full depletion chain). We can conclude,
+then, that there is nothing inherently wrong with the CASL depletion chain or
+OpenMC; it is simply a matter of the underlying nuclear data being inconsistent.
+The 5\% underprediction of $^{85}$Kr is also likely responsible for the slightly
+higher difference in $k_\text{eff}$ for the CASL depletion chain shown in
+\cref{fig:pwr-keff}.
 \begin{figure}[H]
@@
 Serpent uses an explicit temperature treatment known as target motion
-sampling~\cite{viitanen2012nse}. However, we note that for this problem, the
+sampling~\citep{viitanen2012nse}. However, we note that for this problem, the
 temperature dependence of unresolved resonance probability tables becomes
@@
 were used, which resulted in an uncertainty on the effective multiplication
-factor of about 3--4 pcm.
+factor of 3--4 pcm.
 
@@
 
-\Cref{fig:sfr-keff} shows the difference between $k_\text{eff}$ over time
-calculated by OpenMC and Serpent. The blue points correspond to results using
-the full depletion chain, and the green points correspond to results using the
-CASL chain. For each chain, simulations with probability tables on are denoted
-by circles and simulations without probability tables are denoted by squares.
-Unlike in the PWR pincell case, there is a statistically significant difference
-in $k_\text{eff}$ at zero days. Let us first consider the case with probability
-tables turned off. In this case, $k_\text{eff}$ calculated by OpenMC is about 20
-pcm higher than that of Serpent. When using the full depletion chain, there is
-no change in the difference between OpenMC and Serpent. When using the CASL
-depletion chain, the difference increases by about 20--30 pcm, similar to the
-behavior observed in \cref{fig:pwr-keff}. With probability tables turned on,
-however, the situation is quite different. At zero days, $k_\text{eff}$
-calculated by OpenMC is 20 pcm lower than the value calculated by Serpent. As
-the burnup increases, there is an increasing positive bias between OpenMC and
-Serpent. To understand the cause of this difference, it is instructive to look
-at the differences in isotopic compositions.
+\Cref{fig:sfr-keff} shows the difference between $k_\text{eff}$ calculated by
+OpenMC and Serpent as the assembly is depleted. The blue points correspond to
+results using the full depletion chain, and the green points correspond to
+results using the CASL chain. For each chain, simulations with probability
+tables on are denoted by circles and simulations without probability tables are
+denoted by squares. Unlike in the PWR pincell case, there is a statistically
+significant difference in $k_\text{eff}$ at zero days. Let us first consider the
+case with probability tables turned off. In this case, $k_\text{eff}$ calculated
+by OpenMC is about 20 pcm higher than that of Serpent. When using the full
+depletion chain, there is no change in the difference between OpenMC and Serpent
+with increasing burnup. When using the CASL depletion chain, the difference
+increases by about 20--30 pcm, similar to the behavior shown in
+\cref{fig:pwr-keff}. With probability tables turned on, however, the situation
+is quite different. At zero days, $k_\text{eff}$ calculated by OpenMC is 20 pcm
+lower than the value calculated by Serpent. As the burnup increases, there is an
+increasing positive bias between OpenMC and Serpent. To understand the cause of
+this difference, it is instructive to look at differences in isotopic
+compositions.
 \begin{figure}[H]
@@
 fission product concentrations are mostly the same as when probability tables
-were turned off, as shown in \cref{fig:sfr-fp-nop}. We note that because the
-difference in $^{238}$U and $^{239}$Pu was still less than 0.2\% in
-\cref{fig:sfr-actinides}, it is expected that the use of probability tables
-doesn't have a large effect on the fission product concentrations. Again, most
+were turned off, as shown in \cref{fig:sfr-fp-nop}. Because the difference in
+$^{238}$U and $^{239}$Pu concentrations was less than 0.2\% in
+\cref{fig:sfr-actinides}, it makes sense that the use of probability tables
+would not have a large effect on the fission product concentrations. Again, most
 fission product concentrations agree to within 1\%, with $^{109}$Ag and
@@
 through filesystem I/O. The design of the depletion solver relies on a set of
-abstract classes that provide modularity and separation between difference
+abstract classes that provide modularity and separation between different
 components, including the time integration methods, the matrix exponential
@@
 
-We found that when comparing results on the SFR assembly model, larger
-differences in $k_\text{eff}$ and isotopic concentrations were observed when
-unresolved resonance region probability tables were turned on. These differences
-were found to be related to how the temperature dependence of cross sections is
-treated in the unresolved resonance region. In OpenMC, when the temperature of a
-material is in between two temperatures at which probability tables are
-tabulated, statistical interpolation between the tables is performed. Serpent,
-on the other hand, uses probability tables at the lower tabulated temperature.
-For the SFR model, this results in lower parasitic neutron absorption in
-$^{238}$U, lower production of $^{239}$Pu, and ultimately a lower
-$k_\text{eff}$. After depleting for 360 days, there was a 100 pcm difference
-between OpenMC and Serpent due to this effect. Although this is not a huge
-difference, users who are simulating problems with a non-uniform temperature
-distribution should be aware of this behavior.
+When comparing results on the SFR assembly model, larger differences in
+$k_\text{eff}$ and isotopic concentrations were observed when unresolved
+resonance region probability tables were turned on. These differences were found
+to be related to how the temperature dependence of cross sections is treated in
+the unresolved resonance region. In OpenMC, when the temperature of a material
+is inbetween two temperatures at which probability tables are tabulated,
+statistical interpolation between the tables is performed. Serpent, on the other
+hand, uses probability tables at the lower tabulated temperature. For the SFR
+model, this results in lower parasitic neutron absorption in $^{238}$U, lower
+production of $^{239}$Pu, and ultimately a lower $k_\text{eff}$. After depleting
+for 360 days, there was a 100 pcm difference between OpenMC and Serpent due to
+this effect. Although this is not a huge difference, users who are simulating
+problems with a non-uniform temperature distribution should be aware of this
+behavior.
 
@@
 
-Discuss limitations (not validated against experiment, only as good as data).
-Improvements in runtime are needed. Only tested for fission applications.
-Further development and testing needed for other application areas, such as
-transmutation in fusion reactors.
+In the present work, we have only performed a cross-code comparison between
+OpenMC and Serpent. While the results are certainly encouraging, we would
+caution readers that no comparisons against experimental results have been
+performed to date and would be needed to ensure the fitness for a particular
+application. We also highlight the fact that the present comparisons are focused
+solely on fission reactor applications. Although the \texttt{openmc.deplete}
+module may eventually be of interest to other application areas, further changes
+and enhancements would be needed. For example, for use in fusion applications, a
+wider variety of transmutation reactions would need to be accounted for, and the
+ability to handle transmutation from a fixed, pulsed neutron source would be
+required.
 

### 2020-06-24 manuscript.tex
Spellcheck

@@
 used without sacrificing accuracy. For example, on a test problem based on a
-3$\times$3 fuel pin array with gadolinum as a burnable absorber, the CF4 method
+3$\times$3 fuel pin array with gadolinium as a burnable absorber, the CF4 method
 allowed 20 day timesteps while still achieving an accuracy of of 0.1\% on the
@@
 tableau, \citet{josey2016jcp} proposed using the coefficients for the classic
-fourth Runge--Kutta method:
+fourth-order Runge--Kutta method:
 \begin{equation}
@@
 single subclass of \texttt{TransportOperator} that interfaces with the Python
-bindings to OpenMC's C++ API to call a tranport solve and obtain reaction rates.
+bindings to OpenMC's C++ API to call a transport solve and obtain reaction rates.
 Once reaction rates are obtained, a separate depletion chain class (that is not
@@
   \includegraphics[width=4in]{figures/pwr/pwr_actinides_full.pdf}
-  \caption{Relative difference in actinide concetrations between OpenMC and
+  \caption{Relative difference in actinide concentrations between OpenMC and
   Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the full
@@
   \includegraphics[width=4in]{figures/pwr/pwr_actinides_casl.pdf}
-  \caption{Relative difference in actinide concetrations between OpenMC and
+  \caption{Relative difference in actinide concentrations between OpenMC and
   Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the CASL
@@
   \includegraphics[width=4in]{figures/pwr/pwr_fp_full_average.pdf}
-  \caption{Relative difference in fission product concetrations between OpenMC
+  \caption{Relative difference in fission product concentrations between OpenMC
   and Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the
@@
 
-\Cref{fig:pwr-fp-full-thermal} shows the predicted fission product concetrations
+\Cref{fig:pwr-fp-full-thermal} shows the predicted fission product concentrations
 when using the full depletion chain along with thermal FPYs. In this case,
@@
   \includegraphics[width=4in]{figures/pwr/pwr_fp_full_thermal.pdf}
-  \caption{Relative difference in fission product concetrations between OpenMC
+  \caption{Relative difference in fission product concentrations between OpenMC
   and Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the
@@
   \includegraphics[width=4in]{figures/pwr/pwr_fp_casl_average.pdf}
-  \caption{Relative difference in fission product concetrations between OpenMC
+  \caption{Relative difference in fission product concentrations between OpenMC
   and Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the
@@
   \includegraphics[width=4in]{figures/pwr/pwr_fp_casl_thermal.pdf}
-  \caption{Relative difference in fission product concetrations between OpenMC
+  \caption{Relative difference in fission product concentrations between OpenMC
   and Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the
@@
 coupled transport--depletion simulation with Serpent was carried out, one with
-probabilty tables turned on and one with them turned off. For all simulations,
+probability tables turned on and one with them turned off. For all simulations,
 20 inactive batches, 200 active batches, and \num{4.0e6} particles per batch
@@
   \end{subfigure}
-  \caption{Relative difference in actinide concetrations between OpenMC and
+  \caption{Relative difference in actinide concentrations between OpenMC and
   Serpent for the SFR assembly model at 80, 180, 260, and 360 days with
@@
 the predicted actinide concentrations in OpenMC and Serpent, as shown in
-\cref{fig:sfr-actinides-nop}. In this case, the actinide concetrations agree to
+\cref{fig:sfr-actinides-nop}. In this case, the actinide concentrations agree to
 within 0.01\% at all times during depletion for both the full and CASL depletion
@@
   \end{subfigure}
-  \caption{Relative difference in fission product concetrations between OpenMC
+  \caption{Relative difference in fission product concentrations between OpenMC
   and Serpent for the SFR assembly model at 80, 180, 260, and 360 days with
@@
 the unresolved resonance region. In OpenMC, when the temperature of a material
-is inbetween two temperatures at which probability tables are tabulated,
+is between two temperatures at which probability tables are tabulated,
 statistical interpolation between the tables is performed. Serpent, on the other

### 2020-06-24 manuscript.tex
Mention why FPs were chosen

@@
 actinide concentrations between OpenMC and Serpent at four different burnup
-values. We see that most of the differences are less than 0.1\%, and all are
-within 0.15\%. The differences in $^{235}$U and $^{239}$Pu concentrations
-increase slightly with increasing burnup, which most likely accounts for the 20
-pcm difference observed for the full chain in \cref{fig:pwr-keff}. While not
-shown, the predicted actinide concentrations for the reduced chain are nearly
-identical to those observed for the full chain in \cref{fig:pwr-actinides-full}.
-\Cref{fig:pwr-actinides-casl} shows the differences in actinide concentrations
-using the CASL depletion chain. Because the CASL depletion chain differs from
-the full depletion chain primarily with respect to how fission products are
-treated, the differences reported in \cref{fig:pwr-actinides-casl} closely
-mirror those in \cref{fig:pwr-actinides-full}.
+values. The fission products that are shown in \cref{fig:pwr-actinides-full}
+have been chosen because they are either strong neutron absorbers or significant
+contributors to decay heat and/or radiotoxicity. We see that most of the
+differences are less than 0.1\%, and all are within 0.15\%. The differences in
+$^{235}$U and $^{239}$Pu concentrations increase slightly with increasing
+burnup, which most likely accounts for the 20 pcm difference observed for the
+full chain in \cref{fig:pwr-keff}. While not shown, the predicted actinide
+concentrations for the reduced chain are nearly identical to those observed for
+the full chain in \cref{fig:pwr-actinides-full}. \Cref{fig:pwr-actinides-casl}
+shows the differences in actinide concentrations using the CASL depletion chain.
+Because the CASL depletion chain differs from the full depletion chain primarily
+with respect to how fission products are treated, the differences reported in
+\cref{fig:pwr-actinides-casl} closely mirror those in
+\cref{fig:pwr-actinides-full}.
 \begin{figure}[H]

### 2020-06-24 figures/plot_barcharts.py

@@ -52,14 +52,11 @@ actinides = [
 ]
 
 fission_products = [
-    'Kr85', 'Sr90', 'Y90', 'Mo95', 'Tc99', 'Ru101', 'Ru106', 'Rh103', 'Pd105', 'Pd107', 'Ag109',
-    'Te132', 'I129', 'Xe131', 'Xe135', 'Cs133', 'Cs134', 'Cs135', 'Cs137',
-    'La139', 'Ce140', 'Ce142', 'Ce144', 'Pm147', 'Nd142', 'Nd143',
-    'Nd144', 'Nd145', 'Nd146', 'Nd148', 'Nd150', 'Sm147', 'Sm148', 'Sm149',
-#    'Sm150', 'Sm151', 'Sm152', 'Sm154', 'Eu151', 'Eu153', 'Eu154', 'Eu155',
-#    'Gd154', 'Gd155', 'Gd156', 'Gd158', 'Gd160'
-    'Zr91', 'Zr93', 'Zr95', 'Zr96',
-    'Mo96', 'Mo97', 'Mo98', 'Mo99', 'Mo100'
+    'Kr85', 'Sr90', 'Y90', 'Zr93', 'Mo95', 'Mo97', 'Tc99', 'Ru101', 'Ru106',
+    'Rh103', 'Pd105', 'Pd107', 'Ag109', # 'Cd113_m1', 'Sn121_m1', 'Sn126',
+    'Te132', 'I129', 'I131', 'Xe131', 'Xe135', 'Cs133', 'Cs134', 'Cs135', 'Cs137',
+    'La139', 'Ce142', 'Nd143', 'Nd145', 'Pm147', 'Sm149', 'Sm151',
+    # 'Eu155', 'Gd158',
 ]
 
 def sup_label(name):
@@ -110,6 +107,8 @@ def isotope_bar_plot(results_openmc, results_serpent, nuclides, filename, decima
 
     locs = ind + height*(len(burn_indices)/2 - 0.5)
     ax.xaxis.set_major_formatter(mtick.PercentFormatter(1.0, decimals))
+    if decimals >= 3:
+        ax.xaxis.set_tick_params(rotation=30)
     ax.set_xlabel('(OpenMC - Serpent)/Serpent')
     ax.set_yticks(locs)
     ax.set_yticklabels([sup_label(x) for x in nuclides])

### 2020-06-24 manuscript.tex
Move figures

@@
   \centering
-  \includegraphics[width=4in]{figures/pwr/pwr_actinides_full.pdf}
+  \includegraphics[width=4in]{figures/pwr_actinides_full.pdf}
   \caption{Relative difference in actinide concentrations between OpenMC and
@@
   \centering
-  \includegraphics[width=4in]{figures/pwr/pwr_actinides_casl.pdf}
+  \includegraphics[width=4in]{figures/pwr_actinides_casl.pdf}
   \caption{Relative difference in actinide concentrations between OpenMC and
@@
   \centering
-  \includegraphics[width=4in]{figures/pwr/pwr_fp_full_average.pdf}
+  \includegraphics[width=4in]{figures/pwr_fp_full_average.pdf}
   \caption{Relative difference in fission product concentrations between OpenMC
@@
   \centering
-  \includegraphics[width=4in]{figures/pwr/pwr_fp_full_thermal.pdf}
+  \includegraphics[width=4in]{figures/pwr_fp_full_thermal.pdf}
   \caption{Relative difference in fission product concentrations between OpenMC
@@
   \centering
-  \includegraphics[width=4in]{figures/pwr/pwr_fp_casl_average.pdf}
+  \includegraphics[width=4in]{figures/pwr_fp_casl_average.pdf}
   \caption{Relative difference in fission product concentrations between OpenMC
@@
   \centering
-  \includegraphics[width=4in]{figures/pwr/pwr_fp_casl_thermal.pdf}
+  \includegraphics[width=4in]{figures/pwr_fp_casl_thermal.pdf}
   \caption{Relative difference in fission product concentrations between OpenMC
@@
   \begin{subfigure}[t]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{figures/sfr/sfr_actinides_full.pdf}
+    \includegraphics[width=\textwidth]{figures/sfr_actinides_full.pdf}
     \caption{Full depletion chain}
@@
   \begin{subfigure}[t]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{figures/sfr/sfr_actinides_casl.pdf}
+    \includegraphics[width=\textwidth]{figures/sfr_actinides_casl.pdf}
     \caption{CASL depletion chain}
@@
   \begin{subfigure}[t]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{figures/sfr/sfr_actinides_full_nop.pdf}
+    \includegraphics[width=\textwidth]{figures/sfr_actinides_full_nop.pdf}
     \caption{Full depletion chain}
@@
   \begin{subfigure}[t]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{figures/sfr/sfr_actinides_casl_nop.pdf}
+    \includegraphics[width=\textwidth]{figures/sfr_actinides_casl_nop.pdf}
     \caption{CASL depletion chain}
@@
   \begin{subfigure}[t]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{figures/sfr/sfr_fp_full_average.pdf}
+    \includegraphics[width=\textwidth]{figures/sfr_fp_full_average.pdf}
     \caption{Full depletion chain}
@@
   \begin{subfigure}[t]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{figures/sfr/sfr_fp_casl_average.pdf}
+    \includegraphics[width=\textwidth]{figures/sfr_fp_casl_average.pdf}
     \caption{CASL depletion chain}
@@
   \begin{subfigure}[t]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{figures/sfr/sfr_fp_full_average_nop.pdf}
+    \includegraphics[width=\textwidth]{figures/sfr_fp_full_average_nop.pdf}
     \caption{Full depletion chain}
@@
   \begin{subfigure}[t]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{figures/sfr/sfr_fp_casl_average_nop.pdf}
+    \includegraphics[width=\textwidth]{figures/sfr_fp_casl_average_nop.pdf}
     \caption{CASL depletion chain}

### 2020-06-24 figures/plot_barcharts.py

@@ -23,11 +23,11 @@ rcParams['text.latex.preamble'] = [
 
 print('Loading PWR data...')
 results_full_thermal = ResultsList.from_hdf5("pwr/depletion_results_full_thermal.h5")
-results_full_cutoff = ResultsList.from_hdf5("pwr/depletion_results_full_cutoff.h5")
+#results_full_cutoff = ResultsList.from_hdf5("pwr/depletion_results_full_cutoff.h5")
 results_full_average = ResultsList.from_hdf5("pwr/depletion_results_full_average.h5")
 results_reduced = ResultsList.from_hdf5("pwr/depletion_results_reduced_average.h5")
 results_casl_thermal = ResultsList.from_hdf5("pwr/depletion_results_casl_thermal.h5")
-results_casl_cutoff = ResultsList.from_hdf5("pwr/depletion_results_casl_cutoff.h5")
+#results_casl_cutoff = ResultsList.from_hdf5("pwr/depletion_results_casl_cutoff.h5")
 results_casl_average = ResultsList.from_hdf5("pwr/depletion_results_casl_average.h5")
 fuel = serpentTools.read('pwr/serpent_input_dep.m')['fuel']
 
@@ -119,25 +119,25 @@ def isotope_bar_plot(results_openmc, results_serpent, nuclides, filename, decima
     plt.savefig(filename)
 
 ext = 'pdf'
-isotope_bar_plot(results_full_average, fuel, actinides, f'pwr/pwr_actinides_full.{ext}')
-isotope_bar_plot(results_reduced, fuel, actinides, f'pwr/pwr_actinides_reduced.{ext}')
-isotope_bar_plot(results_casl_average, fuel, actinides, f'pwr/pwr_actinides_casl.{ext}')
+isotope_bar_plot(results_full_average, fuel, actinides, f'pwr_actinides_full.{ext}')
+isotope_bar_plot(results_reduced, fuel, actinides, f'pwr_actinides_reduced.{ext}')
+isotope_bar_plot(results_casl_average, fuel, actinides, f'pwr_actinides_casl.{ext}')
 
-isotope_bar_plot(results_sfr_full, fuel_sfr, actinides, f'sfr/sfr_actinides_full.{ext}')
[…]

### 2020-06-24 figures/show_yields.py

@@ -1,0 +1,11 @@
+import openmc.data
+from tabulate import tabulate
+
+nfy = openmc.data.FissionProductYields('/opt/data/endf/endf-b-vii.1/nfy/nfy-092_U_235.endf')
+energies = nfy.energies
+tc109 = [yi['Tc109'].n for yi in nfy.independent]
+mo109 = [yi['Mo109'].n for yi in nfy.independent]
+sn129 = [yi['Sn129'].n for yi in nfy.independent]
+
+print(tabulate(list(zip(energies, tc109, mo109, sn129)),
+               headers=['Energy [eV]', 'Tc109', 'Mo109', 'Sn129']))

### 2020-07-01 manuscript.tex
Address @drewejohnson and @liangjg feedback

@@
 including MCNPX~\citep{waters2007aip} (which has since been merged into
-MCNP6~\citep{goorley2012nt}), Serpent~\citep{leppanen2015ane}, and
-MC21~\citep{griesheimer2015ane}.
+MCNP6~\citep{goorley2012nt}), Serpent~\citep{leppanen2015ane},
+MC21~\citep{griesheimer2015ane}, and RMC~\citep{wang2015ane}.
 
@@
 reactor model, this may entail using a unique material in each fuel pin that is
-depleted separately. Ultimately, the user is responsible for deciding what level
-of spatial discretization is adequate for the problem at hand.
+depleted separately.\footnote{OpenMC includes functions for automatically
+differentiating materials in cells that are repeated throughout a geometry.}
+Ultimately, the user is responsible for deciding what level of spatial
+discretization is adequate for the problem at hand.
 
@@
 nuclides that must be included, the method will recursively go through all decay
-and reaction modes to determine what products may be produced. Using this method
-on the ENDF/B-VII.1 library starting from a list of all stable or very long
-lived (half-life greater than $10^{15}$ sec) nuclides reduces the total number
-of nuclides in the depletion chain from 3,820 to 1,693.
+and reaction modes to determine what products may be produced. This can help
+minimize both the size of the depletion matrix and the number of nuclides for
+which reaction rates must be tallied, thereby reducing the time spent in the
+depletion and transport solvers. Using this method on the ENDF/B-VII.1 library
+starting from a list of all stable or very long lived (half-life greater than
+$10^{15}$ sec) nuclides reduces the total number of nuclides in the depletion
+chain from 3,820 to 1,693.
 
@@
 of the system, carrying away their energy with them. The latter effect may be
-particularly important for small, leaky systems.
+particularly important for small, high-leakage systems.
 

### 2020-07-01 figures/plot_chains.py

@@ -27,12 +27,12 @@ chain_reduced = chain_full.reduce(stable)
 
 def get_patch(z, a, **kwargs):
     n = a - z
-    return Rectangle((n-0.5, z-0.5), 1, 1, **kwargs)
+    return Rectangle((n-0.5, z-0.5), 1, 1, linewidth=0.001, **kwargs)
 
 fig, ax = plt.subplots()
 
-style_full = {'alpha': 0.3}
-style_reduced = {}
+style_full = {'color': '#bcd6e9'}
+style_reduced = {'color': '#1f77b4'}
 style_casl = {'color': 'orange'}
 
 for nuc in chain_full.nuclides:

### 2020-07-01 manuscript.tex
Mention runtime in conclusions

@@
 
+While the focus in this work has primarily been on correctness and accuracy, the
+performance of a coupled transport--depletion simulation is also an important
+consideration that directly impacts usability. The need to determine reaction
+rates for hundreds of nuclides leads to a significant performance degredation
+during a transport simulation. Research on methods to reduce the computational
+expense of obtaining transmutation reaction rates~\cite{salcedoperez2019mc} show
+promise but have yet to be incorporated in a release of OpenMC.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-07-10 manuscript.tex
Incorporate feedback on section 2 from @cjosey

@@
   \begin{split}
-  \frac{dN_i(t)}{dt} = &\sum\limits_j \left [ f_{j \rightarrow i}
-  \int_0^\infty dE \; \sigma_j (E, t) \phi(E,t) + \lambda_{j\rightarrow i}
+  \frac{dN_i(t)}{dt} = &\sum\limits_j \left [ \int_0^\infty dE \;
+  f_{j \rightarrow i}(E) \sigma_j (E, t) \phi(E,t) + \lambda_{j\rightarrow i}
   \right ] N_j(t) \\ &- \left [\int_0^\infty dE \; \sigma_i (E,t) \phi(E,t) +
@@
   \phi(E,t) &\equiv \text{neutron flux at energy $E$ and time $t$} \\
-  f_{j \rightarrow i} &\equiv \text{fraction of transmutation reactions in nuclide $j$ that produce nuclide $i$} \\
+  f_{j \rightarrow i}(E) &\equiv \text{fraction of transmutation reactions in nuclide $j$ that produce nuclide $i$} \\
   \lambda_{j \rightarrow i} &\equiv \text{decay constant for decay modes in nuclide $j$ that produce nuclide $i$} \\
@@
 
-Although the neutron transport equation is time-dependent, the timescale over
-which material compositions change is sufficiently long such that the transport
-equation can be solved as a steady state equation. Without loss of generality,
-let us write the burnup matrix solely as a function of $\vect{n}$ so that
-\cref{eq:depletion-matrix-t} becomes
+Evaluating $\vect{A}$ for a given $\vect{n}$ and $t$ is equivalent to solving
+the neutron transport equation to obtain transmutation reaction rates and
+combining that information with decay and fission product yield data to
+construct the matrix of coefficients. Although the transport equation is
+time-dependent, the timescale over which material compositions change is
+sufficiently long such that the transport equation can be solved as a steady
+state equation. This implies that the only ``input'' that is required by the
+transport solver is $\vect{n}$, as opposed to $\vect{n}$ and $t$. Without loss of
+generality, let us write the burnup matrix solely as a function of $\vect{n}$ so
+that \cref{eq:depletion-matrix-t} becomes
 \begin{equation}
@@
 \end{equation}
-This better reflects the fact that the time, $t$, is generally not an input to
-the transport equation, and hence the burnup matrix is uniquely determined by
-the value of $\vect{n}$.
+This better reflects the fact that the burnup matrix is uniquely determined by
+the value of $\vect{n}$. To be clear, $\vect{A}$ is still a function of time but
+only transitively through the dependence on $\vect{n}(t)$.
 
@@
 outline the available methods in OpenMC, several of which have not been
-previously described in the literature.
+fully described in the literature. An exhaustive analysis of these time
+integration methods and their merits can be found in the thesis of
+\citet{josey2017phd}.
 
@@
 is called commutator-free because it does not require computing any matrix
-commutators like those that appear in \cref{eq:magnus-terms}. Other methods
-based on the Magnus expansion do require computing matrix commutators, some of
-which were explored by \citet{josey2017phd}.} exponential
+commutators like those that appear in \cref{eq:magnus-terms}.} exponential
 integrator~\citep{blanes2006anm} that is derived from the Magnus expansion in
@@
   \label{eq:implicit}
-  \vect{n}_{i=1} = \exp \left( h\vect{A}(\vect{n}_{i+1}) \right) \vect{n}_i.
+  \vect{n}_{i+1} = \exp \left( h\vect{A}(\vect{n}_{i+1}) \right) \vect{n}_i.
 \end{equation}
@@
 $\vect{n}_{i+1}^p$ used in \cref{eq:linear-interp}, obtaining progressively
-better estimates that are then used to solve \cref{eq:linear-ode}. A full
-discussion of the algorithm and its performance and tradeoffs can be found in
-\citep{josey2017phd}.
+better estimates that are then used to solve \cref{eq:linear-ode}.
 
@@
 CE/CM method requires two evaluations and is thus twice as expensive as the
-predictor method, but achieves an error of $\mathcal{O}(h^2)$. An exhaustive
-description of time integration methods and their merits can be found in the
-thesis of \citet{josey2017phd}.
+predictor method, but achieves an error of $\mathcal{O}(h^2)$.
 
@@
 difference during depletion. The most well-known example is capture in
-$^{241}$Am, which can produce either $^{242}$Am or $^{242\text{m}}$Am. Because the
-metastable state has a significantly longer half-life than the ground state, it
-is important to account for the branching of the capture reaction in $^{241}$Am
-to accurately predict the production of higher actinides~\citep{haeck2012nse}.
-This is complicated by the fact that the branching ratio may depend on the
-incident neutron energy causing capture.
+$^{241}$Am, which can produce either $^{242}$Am or $^{242\text{m}}$Am. Because
+the metastable state has a significantly longer half-life than the ground state,
+it is important to account for the branching of the capture reaction in
+$^{241}$Am to accurately predict the production of higher
+actinides~\citep{haeck2012nse}. This is complicated by the fact that the
+branching ratio may depend on the incident neutron energy causing capture (hence
+why $f_{j\rightarrow i}$ in \cref{eq:depletion} depends on $E$).
 
 OpenMC does not currently implement a method to handle energy-dependent capture
-branching ratios directly. However, the depletion chain file does allow a
-reaction to be listed multiple times with different branching ratios that result
-in different products. Spectrum-averaged capture branching ratios have been
-computed in LWR and SFR spectra and are available at
-\url{https://openmc.org/depletion-chains}. Within the Python API, the class
-representing a depletion chain allows branching ratios to be easily retrieved
-and changed. We note that while this treatment doesn't accurately account for
-the change in the capture branching ratio with increasing burnup (due to
-spectral changes), these changes are generally much smaller than the differences
-in capture branching ratios between nuclear data evaluations.
+branching ratios directly and treats $f_{j\rightarrow i}$ as a constant.
+However, the depletion chain file does allow a reaction to be listed multiple
+times with different branching ratios that result in different products.
+Spectrum-averaged capture branching ratios have been computed in LWR and SFR
+spectra and are available at \url{https://openmc.org/depletion-chains}. Within
+the Python API, the class representing a depletion chain allows branching ratios
+to be easily retrieved and changed. We note that while this treatment doesn't
+accurately account for the change in the capture branching ratio with increasing
+burnup (due to spectral changes), these changes are generally much smaller than
+the differences in capture branching ratios between nuclear data evaluations.
 

### 2020-07-13 manuscript.tex
More suggested changes from @cjosey

@@
 differentiation formula rather than a method based on an exponential
-integrator~\citep{carpenter2009mc,hykes2017mc,sublet2017nds}.
+integrator~\citep{carpenter2009mc,hykes2017mc,sublet2017nds}. The next section
+shows one such example that is used in OpenMC.
 
@@
 OpenMC and Serpent. The blue points correspond to results using the full
-depletion chain, the red points correspond to points using the reduced chain,
-and the green points correspond to results using the CASL chain. For each chain,
-the use of interpolated FPYs is denoted by circles and thermal FPYs by x's. At
-zero days, the difference in $k_\text{eff}$ between OpenMC and Serpent is not
-statistically significant. As the burnup increases, the difference for the full
-chain cases rises to about 20 pcm. The reduced chain case exhibits the exact
-same behavior as the full chain as one would expect. Use of the CASL chain
+depletion chain, the gray points correspond to points using the reduced chain,
+and the orange points correspond to results using the CASL chain. For each
+chain, the use of interpolated FPYs is denoted by circles and thermal FPYs by
+x's. At zero days, the difference in $k_\text{eff}$ between OpenMC and Serpent
+is not statistically significant. As the burnup increases, the difference for
+the full chain cases rises to about 20 pcm. The reduced chain case exhibits the
+exact same behavior as the full chain as one would expect. Use of the CASL chain
 results in a slightly larger difference between OpenMC and Serpent at 50 MWd/kg,
@@
   and Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the
-  full depletion chain and interpolated FPYs.}
+  full depletion chain and interpolated FPYs in OpenMC.}
   \label{fig:pwr-fp-full-average}
@@
   and Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the
-  full depletion chain and thermal FPYs.}
+  full depletion chain and thermal FPYs in OpenMC.}
   \label{fig:pwr-fp-full-thermal}
@@
   and Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the
-  CASL depletion chain and interpolated FPYs.}
+  CASL depletion chain and interpolated FPYs in OpenMC.}
   \label{fig:pwr-fp-casl-average}
@@
   and Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the
-  CASL depletion chain and thermal FPYs.}
+  CASL depletion chain and thermal FPYs in OpenMC.}
   \label{fig:pwr-fp-casl-thermal}
@@
 OpenMC and Serpent as the assembly is depleted. The blue points correspond to
-results using the full depletion chain, and the green points correspond to
+results using the full depletion chain, and the orange points correspond to
 results using the CASL chain. For each chain, simulations with probability
@@
 problems with a non-uniform temperature distribution should be aware of this
-behavior.
+behavior. This difference could be mitigated by using nuclear data processed on
+a finer temperature grid---in our results, cross sections and probability table
+data were being interpolated over a rather large \SI{300}{\kelvin} interval.
 
@@
 during a transport simulation. Research on methods to reduce the computational
-expense of obtaining transmutation reaction rates~\cite{salcedoperez2019mc} show
-promise but have yet to be incorporated in a release of OpenMC.
+expense of obtaining transmutation reaction rates~\citep{salcedoperez2019mc}
+show promise but have yet to be incorporated in a release of OpenMC.
 

### 2020-07-13 figures/plot_barcharts.py

@@ -21,6 +21,8 @@ rcParams['text.latex.preamble'] = [
     r'\usepackage{amsmath}',
 ]
 
+plt.style.use('tableau-colorblind10')
+
 print('Loading PWR data...')
 results_full_thermal = ResultsList.from_hdf5("pwr/depletion_results_full_thermal.h5")
 #results_full_cutoff = ResultsList.from_hdf5("pwr/depletion_results_full_cutoff.h5")

### 2020-07-13 figures/plot_keff.py

@@ -17,6 +17,8 @@ rcParams['text.latex.preamble'] = [
     r'\usepackage{amsmath}',
 ]
 
+plt.style.use('tableau-colorblind10')
+
 # Read Serpent results --------------------------------------------------------
 
 def get_serpent_keff(filename):
@@ -57,11 +59,11 @@ burnup = [
 ]
 
 fig, ax = plt.subplots()
-ax.plot(burnup, difference(results_full_average, serpent_pwr), 'bo', label='Full, Interpolated FPY')
-ax.plot(burnup, difference(results_full_thermal, serpent_pwr), 'bx', label='Full, Thermal FPY')
-ax.plot(burnup, difference(results_reduced, serpent_pwr), 'ro', label='Reduced, Interpolated FPY')
-ax.plot(burnup, difference(results_casl_average, serpent_pwr), 'go', label='CASL, Interpolated FPY')
-ax.plot(burnup, difference(results_casl_thermal, serpent_pwr), 'gx', label='CASL, Thermal FPY')
+ax.plot(burnup, difference(results_full_average, serpent_pwr), 'o', color='C0', label='Full, Interpolated FPY')
+ax.plot(burnup, difference(results_full_thermal, serpent_pwr), 'x', color='C0', label='Full, Thermal FPY')
+ax.plot(burnup, difference(results_reduced, serpent_pwr), 'o', color='C2', label='Reduced, Interpolated FPY')
+ax.plot(burnup, difference(results_casl_average, serpent_pwr), 'go', color='C1', label='CASL, Interpolated FPY')
+ax.plot(burnup, difference(results_casl_thermal, serpent_pwr), 'gx', color='C1', label='CASL, Thermal FPY')
 ax.plot(color='k', linestyle='--')
 ax.set_xlabel("Burnup [MWd/kg]")
 ax.set_ylabel(r"$k_\text{OpenMC} - k_\text{Serpent}$ [pcm]")
@@ -75,10 +77,10 @@ plt.close()
 days = np.linspace(0.0, 360.0, 19)
 
 fig, ax = plt.subplots()
-ax.plot(days, difference(results_sfr_full, serpent_sfr), 'bo', label='Full, ptables')
-ax.plot(days, difference(results_sfr_casl, serpent_sfr), 'go', label='CASL, ptables')
[…]

### 2020-07-14 manuscript.tex
Move footnote into caption of algorithm

@@
 tabulated for orders up to $k=48$. Rather than computing the full approximation
-and then multiplying it by a vector, \cref{alg:cram}\footnote{The original
-description of the algorithm presented by \citet{pusa2016nse} contains a typo.}
-is used to incrementally apply the terms within the product. The $k$th order
-approximation for CRAM requires solving $k/2$ sparse linear systems. OpenMC
-relies on functionality from scipy.sparse.linalg~\citep{virtanen2020nm} for
-solving the linear system.
+and then multiplying it by a vector, \cref{alg:cram} is used to incrementally
+apply the terms within the product. The $k$th order approximation for CRAM
+requires solving $k/2$ sparse linear systems. OpenMC relies on functionality
+from scipy.sparse.linalg~\citep{virtanen2020nm} for solving the linear system.
 \begin{algorithm}[H]
-  \caption{Incomplete partial fraction form of CRAM.}
+  \caption{Incomplete partial fraction form of CRAM. Note that the original
+  description of the algorithm presented by \citet{pusa2016nse} contains a
+  typo.}
   \label{alg:cram}
-  \begin{algorithmic}[1]
+  \begin{algorithmic}
     \State $\vect{n} \gets \vect{n_0}$

### 2020-07-15 manuscript.tex
Edits from Gail Pieper

@@
 \address[lanl]{Los Alamos National Laboratory, P.O. Box 1663, Los Alamos, NM 87545, United States}
-\address[gatech]{Georgia Institute of Technology, 770 State St NW, Atlanta, GA 30318, United States}
+\address[gatech]{Georgia Institute of Technology, 770 State St. NW, Atlanta, GA 30318, United States}
 \address[tsinghua]{Institute of Nuclear and New Energy Technology, Tsinghua University, Beijing, China}
@@
   cost, accuracy, and memory use. For all time integration methods, evaluation
-  of the matrix exponential is performed using the incomplete partial fraction
+  of the matrix exponential is performed by using the incomplete partial fraction
   form of the Chebyshev rational approximation method.
 
-  Simulations of a PWR pincell and an SFR assembly were carried out with OpenMC
+  Simulations of a pressurized water reactor (PWR) pincell and a sodium-cooled fast reactor (SFR) assembly were carried out with OpenMC
   and Serpent. For both problems, the use of a high-fidelity depletion chain
@@
   product yields is handled or deficiencies in the nuclear data used. OpenMC
-  simulations of the PWR and SFR problems using a simplified 228 nuclide
+  simulations of the PWR and SFR problems using a simplified 228-nuclide
   depletion chain demonstrate that it achieves accuracy close to that of the
@@
 When a material in a system is subject to irradiation over a long period of
-time, nuclides within the material will transmute due to nuclear reactions and
+time, nuclides within the material will transmute because of nuclear reactions and
 spontaneous radioactive decay. The time-dependent process by which nuclides
@@
 the transport equation. As will be shown in \cref{sec:methods}, this results in
-a non-linear coupling between transport and depletion. The implication is that
+a nonlinear coupling between transport and depletion. The implication is that
 depletion codes are rarely used in isolation---normally, they are directly
@@
 code to the other. Thus, the actual transport/depletion codes themselves
-required no source code changes. In the last decade, many MC code developers
+required no source code changes. In the past decade, many MC code developers
 have chosen to directly include a depletion solver as a first-rate feature,
@@
 source, community-developed MC particle transport code~\citep{romano2015ane1}.
-To date, OpenMC has not had a built-in depletion solver. This has led to
-numerous efforts in the community toward developing coupled transport--depletion
+To date, OpenMC has not had a built-in depletion solver. This lack has led to
+numerous efforts in the community to develop coupled transport--depletion
 calculations with OpenMC~\citep{gul2017ane, lanversin2017icone,
@@
 coupled calculations with OpenMC, none of them has resulted in a validated,
-openly-available, and generalized capability that OpenMC users can take
+openly available, and generalized capability that OpenMC users can take
 advantage of. The capability that is described herein is part of the OpenMC
@@
 within OpenMC, and how nuclear data is used to create depletion chains. Then, in
-\cref{sec:results}, we present results on two problems that demonstrate the
-accuracy of OpenMC when compared to Serpent~\citep{leppanen2015ane}. In
+\cref{sec:results}, we present results for two problems that demonstrate the
+accuracy of OpenMC when compared with Serpent~\citep{leppanen2015ane}. In
 \cref{sec:conclusions}, we summarize the findings, discuss limitations, and
@@
 The equation that governs the transmutation of nuclides by nuclear reactions and
-radioactive decay inside of an irradiated environment can be written as
+radioactive decay inside an irradiated environment can be written as
 \begin{equation}
@@
   \right ] N_j(t) \\ &- \left [\int_0^\infty dE \; \sigma_i (E,t) \phi(E,t) +
-  \sum\limits_j \lambda_{i\rightarrow j} \right ] N_i(t), \quad i=1,\dots,n
+  \sum\limits_j \lambda_{i\rightarrow j} \right ] N_i(t), \quad i=1,\dots,n ,
   \end{split}
@@
 sections. For \cref{eq:depletion} to be accurate, the spatial variation of the
-reaction rates over the region which is being depleted should be small. For a
+reaction rates over the region that is being depleted should be small. For a
 reactor model, this may entail using a unique material in each fuel pin that is
@@
 on the density of other nuclides, $\{N_j \,|\, j \ne i\}$, \cref{eq:depletion}
-represents a system of first-order ordinary differential equations (ODEs). To
+represents a system of first-order ordinary differential equations (ODEs). In order to
 form a proper initial value problem, the nuclide densities at time 0 must be
@@
   \frac{d\vect{n}}{dt} = \vect{A}(\vect{n},t)\vect{n}, \quad \vect{n}(0) =
-  \vect{n}_0
+  \vect{n}_0,
 \end{equation}
@@
 
-The principle difficulty of solving \cref{eq:depletion-matrix} is the fact that
+The principal difficulty in solving \cref{eq:depletion-matrix} is that
 the elements of $\vect{A}$ are not constant with respect to time. If
@@
 \begin{equation}
-  \exp(\vect{X}) = \sum\limits_{k=0}^\infty \frac{\vect{X}^k}{k!}
+  \exp(\vect{X}) = \sum\limits_{k=0}^\infty \frac{\vect{X}^k}{k!},
 \end{equation}
@@
 \begin{enumerate}
-  \item Use of a numerical method to integrate \cref{eq:depletion-matrix}
+  \item Using a numerical method to integrate \cref{eq:depletion-matrix}
   forward in time, usually in a form that involves taking one or more matrix
-  exponentials, and
+  exponentials. 
   \item Evaluating the matrix exponential or, equivalently, the action of a
-  matrix exponential on a vector.
+  matrix exponential on a vector
 \end{enumerate}
 We will henceforth refer to the first component as \emph{time integration}. Let
-us now consider how each of these components are treated in OpenMC.
+us now consider how each of these components is treated in OpenMC.
 
@@
 one must resort to using a numerical method that generates approximate solutions
-in discrete increments. The problem we need to solve then is: given $\vect{n}_i
+in discrete increments. The problem we need to solve then is: Given $\vect{n}_i
 \equiv \vect{n}(t_i)$, what is the value of $\vect{n}_{i+1} \equiv
 \vect{n}(t_{i+1}) = \vect{n}(t_i + h)$, with $h$ being the step size? Without
-loss of generality, we shall choose $t_i=0$ as it makes the presentation of
+loss of generality, we  choose $t_i=0$ because it makes the presentation of
 various methods simpler.
 
-OpenMC doesn't rely on a single time integration method but instead has several
+OpenMC does not rely on a single time integration method but instead has several
 methods available that the user can choose. In the following sections, we
-outline the available methods in OpenMC, several of which have not been
+outline these methods in OpenMC, several of which have not been
 fully described in the literature. An exhaustive analysis of these time
@@
 The simplest numerical method for solving \cref{eq:depletion-matrix} over a
-timestep is known as the \emph{predictor} method. It proceeds by assuming that
+timestep is known as the \emph{predictor} method. It assumes that
 $\vect{A}$ is constant over the timestep and assumes the value at the
@@
 $\vect{A}$ is evaluated at the beginning of the step and used to deplete the
-composition over half the step. Then, $\vect{A}$ is reevaluated using the
-predicted middle-of-step composition and used to deplete over the full step. The
+composition over half the step. Then, $\vect{A}$ is reevaluated by using the
+predicted middle-of-step composition and is used to deplete over the full step. The
 algorithm can be expressed as
@@
 linear interpolation) method, the end-of-step composition is first predicted
-using $\vect{A}(\vect{n}_i)$:
+by using $\vect{A}(\vect{n}_i)$:
 \begin{equation}
@@
   \label{eq:linear-ode}
-  \frac{d\vect{n}}{dt} = \hat{\vect{A}}(t) \vect{n}
+  \frac{d\vect{n}}{dt} = \hat{\vect{A}}(t) \vect{n},
 \end{equation}
@@
 \end{equation}
-We see that $\hat{\vect{A}}$ no longer depends on $\vect{n}$ thereby making
+We see that $\hat{\vect{A}}$ no longer depends on $\vect{n}$, thereby making
 \cref{eq:linear-ode} a linear nonautonomous ODE. Because $\hat{\vect{A}}$ is a
@@
 While using substeps is a reasonable approach for solving \cref{eq:linear-ode},
-there are other methods as well. For example, some codes use a backward
+other methods are also available. For example, some codes use a backward
 differentiation formula rather than a method based on an exponential
@@
 
-\subsubsection{Magnus expansion integrators}
+\subsubsection{Magnus Expansion Integrators}
 
@@
     \vect{\Omega}_3(t) &= \int_0^t dt_1 \int_0^{t_1} dt_2 \int_0^{t_2} dt_3
-      \left ( [\vect{A}(t_1), [\vect{A}(t_2), \vect{A}(t_3)]] + [\vect{A}(t_3), [\vect{A}(t_2), \vect{A}(t_1)]] \right )
+      \left ( [\vect{A}(t_1), [\vect{A}(t_2), \vect{A}(t_3)]] + [\vect{A}(t_3), [\vect{A}(t_2), \vect{A}(t_1)]] \right ),
   \end{split}
@@
 is called commutator-free because it does not require computing any matrix
-commutators like those that appear in \cref{eq:magnus-terms}.} exponential
+commutators as do those that appear in \cref{eq:magnus-terms}.} exponential
 integrator~\citep{blanes2006anm} that is derived from the Magnus expansion in
@@
   \vect{n}_{i+1} = \exp (a_2 \vect{A}_1 + a_1 \vect{A}_2)
-  \exp (a_1 \vect{A}_1 + a_2 \vect{A}_2) \vect{n}_i
+  \exp (a_1 \vect{A}_1 + a_2 \vect{A}_2) \vect{n}_i ,
 \end{equation}
@@
 these approaches has shown that \cref{eq:celi-cfq4} achieves better accuracy
-than the use of five substeps~\citep{josey2017phd}. OpenMC also includes an
+than does the use of five substeps~\citep{josey2017phd}. OpenMC also includes an
 integrator based on the LE/QI (linear extrapolation, quadratic interpolation)
-method~\citep{isotalo2011ane2} where the integration of \cref{eq:linear-ode} (in
+method~\citep{isotalo2011ane2}, where the integration of \cref{eq:linear-ode} (in
 which case $\hat{\vect{A}}(t)$ is represented by a quadratic polynomial in $t$)
-is performed using the fourth-order commutator-free Magnus integrator based on
+is performed by using the fourth-order commutator-free Magnus integrator based on
 \cref{eq:compose4}.
@@
 This method requires four transport solutions and five matrix exponentials per
-timestep. However, because it is fourth-order, it allows large timesteps to be
+timestep. Because it is fourth-order, however, it allows large timesteps to be
 used without sacrificing accuracy. For example, on a test problem based on a
 3$\times$3 fuel pin array with gadolinium as a burnable absorber, the CF4 method
-allowed 20 day timesteps while still achieving an accuracy of of 0.1\% on the
+allowed 20-day timesteps while still achieving an accuracy of 0.1\% on the
 $^{157}$Gd number density~\citep{josey2017phd}.
@@
     \hat{\vect{n}}_k &= \exp \left( h \sum_{j=1}^{k-1} a_{ij} \vect{A}(\hat{\vect{n}}_j) \right) \vect{n}_i \\
-    \vect{n}_{i+1} &= \exp \left( h \sum_{j=1}^{s} b_{j} \vect{A}(\hat{\vect{n}}_j) \right) \vect{n}_i
+    \vect{n}_{i+1} &= \exp \left( h \sum_{j=1}^{s} b_{j} \vect{A}(\hat{\vect{n}}_j) \right) \vect{n}_i .
   \end{split}
@@
 All of the integration methods discussed thus far are explicit methods that are
-only conditionally stable. There are many well-documented cases where
+only conditionally stable. In many well-documented cases 
 oscillatory behavior and diverging solutions have been observed. To address
-this, \citet{dufek2013ane} proposed an implicit-Euler-like method that is
+this situation, \citet{dufek2013ane} proposed an implicit-Euler-like method that is
 unconditionally stable. The main idea is to replace $\vect{A}(\vect{n}_i)$ with
@@
 \end{equation}
-This results in a non-linear equation for $\vect{n}_{i+1}$ that cannot be solved
-directly. However, it is possible to solve for $\vect{n}_{i+1}$ using a
+This results in a nonlinear equation for $\vect{n}_{i+1}$ that cannot be solved
+directly. It is possible, however, to solve for $\vect{n}_{i+1}$ by using a
 fixed-point iteration. \citet{dufek2013ane} utilized a stochastic approximation
@@
 between the accuracy of the method, the computational expense, and the memory
-requirements. When a MC code is used, the expense is dominated by the time
-to compute the transport solution, i.e., to evaluate $\vect{A}$ for a given
+requirements. When an MC code is used, the expense is dominated by the time
+to compute the transport solution, in other words, to evaluate $\vect{A}$ for a given
 $\vect{n}$. Thus, the cost of a method scales with the number of $\vect{A}$
@@
 require more evaluations generally achieve higher accuracy. The predictor method
-only requires one evaluation and its error converges as $\mathcal{O}(h)$. The
+requires only one evaluation, and its error converges as $\mathcal{O}(h)$. The
 CE/CM method requires two evaluations and is thus twice as expensive as the
@@
     \vect{I} + 2 \text{Re} \left ( \widetilde{\alpha}_\ell \left (\vect{A}t
-    - \theta_\ell \vect{I} \right )^{-1} \right ) \right )
+    - \theta_\ell \vect{I} \right )^{-1} \right ) \right ),
 \end{equation}
@@
 In principle, using CRAM for evaluating matrix exponentials is fairly simple:
-construct the burnup matrix at various times and solve a set of sparse linear
+Construct the burnup matrix at various times, and solve a set of sparse linear
 systems. However, constructing the burnup matrix, $\vect{A}$, involves not only
-solving the transport equation to estimate nuclear reaction rates but also a
+solving the transport equation to estimate nuclear reaction rates but also making a
 series of choices about what data to include.
@@
 
-In OpenMC, the burnup matrix is constructed based on data inside of a
+In OpenMC, the burnup matrix is constructed based on data inside a
 \emph{depletion chain} file that includes fundamental data gathered from ENDF
 incident neutron, decay, and fission product yield (FPY) sublibraries. For each
-nuclide, this file includes:
+nuclide, this file includes the following:
 \begin{itemize}
-  \item What transmutation reactions are possible, their Q values, and their products;
+  \item What transmutation reactions are possible, their Q values, and their products
   \item If a nuclide is not stable, what decay modes are possible, their
-  branching ratios, and their products; and
+  branching ratios, and their products  
   \item If a nuclide is fissionable, the fission products yields at any number
-  of incident neutron energies.
+  of incident neutron energies
 \end{itemize}
@@
 Depletion chains can be automatically created in the Python application
-programming interface (API) in the following manner. First, the user supplies a
+programming interface in the following manner. First, the user supplies a
 list of ENDF incident neutron, decay, and FPY sublibrary files. The incident
@@
 nuclide will undergo a $\upbeta^-$ decay. Once the replacement nuclide has been
-determined, OpenMC will check again if it exists in the decay sublibrary files.
+determined, OpenMC will check again whether it exists in the decay sublibrary files.
 If not, the process is repeated.
@@
   \item If the nuclide missing FPYs is a metastable state, use FPYs from the
-  ground state;
-  \item Use FPYs from an isotone if available;
+  ground state.
+  \item Use FPYs from an isotone if available.
   \item If all else fails, use FPYs from $^{235}$U.
@@
 nuclides will result in an $n\times n$ burnup matrix, there is an incentive to
-reduce the number of nuclides in the depletion chain to minimize both the size
+reduce the number of nuclides in the depletion chain in order to minimize both the size
 of the burnup matrix and the number of nuclides that are tracked during a
@@
 nuclides that must be included, the method will recursively go through all decay
-and reaction modes to determine what products may be produced. This can help
+and reaction modes to determine what products may be produced. This process can help
 minimize both the size of the depletion matrix and the number of nuclides for
@@
 depletion and transport solvers. Using this method on the ENDF/B-VII.1 library
-starting from a list of all stable or very long lived (half-life greater than
+starting from a list of all stable or very long-lived (half-life greater than
 $10^{15}$ sec) nuclides reduces the total number of nuclides in the depletion
@@
 
-OpenMC will setup tallies in a problem based on what transmutation reactions are
+OpenMC will set up tallies in a problem based on what transmutation reactions are
 available in a depletion chain file, so any arbitrary number of transmutation
@@
 reactions: fission, (n,$\gamma$), (n,2n), (n,3n), (n,4n), (n,p), and
-(n,$\alpha$). While this is suitable for fission applications, fusion
+(n,$\alpha$). While this list is suitable for fission applications, fusion
 applications may require a wider set of transmutation reactions to be tracked.
@@
 $^{241}$Am to accurately predict the production of higher
-actinides~\citep{haeck2012nse}. This is complicated by the fact that the
+actinides~\citep{haeck2012nse}. Doing so is complicated, however, by the fact that the
 branching ratio may depend on the incident neutron energy causing capture (hence
@@
 
-OpenMC does not currently implement a method to handle energy-dependent capture
+OpenMC does not currently implement a method for handling energy-dependent capture
 branching ratios directly and treats $f_{j\rightarrow i}$ as a constant.
@@
 Spectrum-averaged capture branching ratios have been computed in LWR and SFR
-spectra and are available at \url{https://openmc.org/depletion-chains}. Within
+spectra and are available at \url{https://openmc.org/depletion-chains}. In
 the Python API, the class representing a depletion chain allows branching ratios
-to be easily retrieved and changed. We note that while this treatment doesn't
+to be easily retrieved and changed. We note that while this treatment does not
 accurately account for the change in the capture branching ratio with increasing
@@
 contains a continuum of energies. The ENDF-6 formats manual~\citep{trkov2018bnl}
-recommends using linear-linear interpolation between energy points. However, at
-the end of a transport solve, we typically have a one-group or multigroup
-fission reaction rate---it's not clear how one should use this information to
+recommends using linear-linear interpolation between energy points. At
+the end of a transport solve, however, we typically have a one-group or multigroup
+fission reaction rate---and it is not clear how this information should be used  to
 calculate an effective FPY.
 
-Different codes have adopted different approaches in how to handle the energy
+Different codes have adopted different approaches for handling the energy
 dependence of FPYs. In ORIGEN, the average neutron energy causing fission is
@@
 The reaction rates calculated by OpenMC are given in units of reactions per
-source particle. For depletion, it is necessary to compute an absolute reaction
+source particle. For depletion, one must compute an absolute reaction
 rate in reactions per second. To do so, the total observed heating in the system
@@
 to estimate the total heating. This method is particularly useful for making
-inter-code comparisons, where one must ensure that the same Q values for fission
+intercode comparisons, where one must ensure that the same Q values for fission
 are applied. In the second method, the actual heating rate is tallied over the
-entire system (separate from the tally for transmutation reaction rates). This
-method makes it possible to account for the energy dependence of the fission
+entire system (separate from the tally for transmutation reaction rates). With this
+method one can account for the energy dependence of the fission
 energy release data as well as the fact that neutrons and photons may leak out
@@
 library with a C++ API that can be called by external programs. OpenMC's Python
-API includes bindings to the C++ API; this enables a single run to be broken up
+API includes bindings to the C++ API; these enable a single run to be broken up
 into individual components. For example, a function call from the Python
@@
 advantages. First, the time necessary to load cross sections is amortized
-because it only has to be performed once, as opposed to once per transport
+because it has to be performed only once, as opposed to once per transport
 simulation. Second, using the Python bindings to the C++ API allows results to
-be read directly in memory rather than using filesystem I/O, which helps to
-minimizing overhead, particularly for large simulations.
+be read directly in memory rather than using filesystem I/O, thereby helping 
+minimize overhead, particularly for large simulations.
 
@@
 that represent the major components of the depletion solver. Each of the
-abstract classes themselves defines an interface, that is, a set of methods that
+abstract classes defines an interface, that is, a set of methods that
 a subclass must implement. First, the abstract \texttt{TransportOperator} class
 represents a particle transport solver. Subclasses of \texttt{TransportOperator}
-are required to implement a method that takes a list of nuclide densities to be
+are required in order to implement a method that takes a list of nuclide densities to be
 used in the problem and a system power and returns the $k$-eigenvalue and the
@@
 evaluate \cref{eq:constant-A}). Subclasses of \texttt{DepSystemSolver} are
-required to implement a method that takes the burnup matrix $\vect{A}$, an
-initial density vector $\vect{n}_0$, and a timestep size, and returns the
+required in order to implement a method that takes the burnup matrix $\vect{A}$, an
+initial density vector $\vect{n}_0$ and a timestep size and returns the
 resulting nuclide densities at the end of the step. The only subclass of
@@
 This class effectively orchestrates the coupled transport--depletion simulation.
-Subclasses of \texttt{Integrator} are required to implement a method that takes
+Subclasses of \texttt{Integrator} are required in order to implement a method that takes
 an initial vector of nuclide densities, reaction rates returned from an
-operator, a time interval, and a power, and returns the resulting nuclide
+operator, a time interval, and a power and returns the resulting nuclide
 densities at the end of the step. Thus, \texttt{Integrator} utilizes
-\texttt{TransportOperator} in order to obtain reaction rates given $\vect{n}$,
+\texttt{TransportOperator} in order to obtain reaction rates given $\vect{n}$;
 and once the burnup matrix is constructed, it utilizes \texttt{DepSystemSolver}
@@
 
-Lastly, we note that when multiple depletable materials are present in a problem
+We note that when multiple depletable materials are present in a problem
 (thereby requiring multiple independent solutions of
@@
 distribute the material compositions among the participating MPI processes.
-Then, within each process, Python's \texttt{multiprocessing} standard-library
+Then, within each process, Python's \texttt{multiprocessing} standard library
 module is used to distribute materials among threads within a single MPI
@@
 using the \texttt{openmc.deplete} module as well as a developmental version of
-Serpent 2.1.32 on two simple problems. The first problem is a pressurized water
-reactor (PWR) pincell and the second problem is a single assembly of a
-sodium-cooled fast reactor (SFR). These problems were selected to cover both
+Serpent 2.1.32 on two simple problems. The first problem is a 
+PWR pincell, and the second problem is a single assembly of an
+SFR. These problems were selected to cover both
 thermal and fast neutron spectra. While we had originally attempted to use
@@
 $^{97}$Mo. This difference was found to be due to $^{97\text{m}}$Y having a
-non-zero independent yield and a zero cumulative yield in the FPY sublibrary
+nonzero independent yield and a zero cumulative yield in the FPY sublibrary
 file for $^{235}$U fission in ENDF/B-VII.1. The zero cumulative yield caused
@@
 fundamental nuclear data are being used in both codes, one expects to obtain the
-same answer. Thus, even very minute differences may be indicative of a true
+same answer. Thus, even minute differences may be indicative of a true
 problem in the method implementation in one of the codes.
@@
 
-Given the remarks above, it is of utmost importance to ensure that the same
+Given the remarks above, one must take utmost care to ensure that the same
 nuclear data are being used in a cross-code comparison. For a depletion
@@
 secondary angle/energy distributions, radioactive decay data, and FPYs are the
-same for both codes. In the simulations performed, all data was based on the
+same for both codes. In the simulations performed, all data were based on the
 ENDF/B-VII.1 nuclear data library~\citep{chadwick2011nds}. Incident neutron
@@
 simulations in both OpenMC and Serpent. Serpent is able to use the ACE files
-from the endf71x library directly while OpenMC uses an HDF5 format that requires
+from the endf71x library directly, whereas OpenMC uses an HDF5 format that requires
 conversion of the ACE files first. For the PWR pincell problem,
-$S(\alpha,\beta)$ thermal scattering data is also needed for hydrogen in light
+$S(\alpha,\beta)$ thermal scattering data are also needed for hydrogen in light
 water; because Serpent does not yet handle the continuous representation used in
-the ENDF71SaB library, the light water data from ENDF/B-VII.0 was used instead.
-We note that the light water evaluation didn't actually change between
+the ENDF71SaB library, the light water data from ENDF/B-VII.0 were used instead.
+We note that the light water evaluation did not actually change between
 ENDF/B-VII.0 and ENDF/B-VII.1.
@@
 \begin{itemize}
-  \item A single 0.1 MWd/kg, 0.4 MWd/kg, and 0.5 MWd/kg timesteps;
-  \item 1.0 MWd/kg timesteps out to a cumulative burnup of 10.0 MWd/kg;
-  \item 2.5 MWd/kg timesteps out to a cumulative burnup of 50.0 MWd/kg.
+  \item A single 0.1 MWd/kg, 0.4 MWd/kg, and 0.5 MWd/kg timesteps
+  \item 1.0 MWd/kg timesteps out to a cumulative burnup of 10.0 MWd/kg
+  \item 2.5 MWd/kg timesteps out to a cumulative burnup of 50.0 MWd/kg
 \end{itemize}
@@
 Five coupled transport--depletion simulations were carried out with OpenMC to test
-a variety of combinations of depletion chains and methods for handling FPY
+various combinations of depletion chains and methods for handling FPY
 energy dependence:
@@
 \end{enumerate}
-Because Serpent doesn't have an option for changing the depletion chain or the
+Because Serpent does not have an option for changing the depletion chain or the
 method for handling FPY energy dependence, a single coupled transport--depletion
@@
 the full chain cases rises to about 20 pcm. The reduced chain case exhibits the
-exact same behavior as the full chain as one would expect. Use of the CASL chain
+exact same behavior as the full chain does, as one would expect. Use of the CASL chain
 results in a slightly larger difference between OpenMC and Serpent at 50 MWd/kg,
 about 40 pcm. We also see that for both the full and CASL chains, changing the
-method for handling FPY energy dependence doesn't appear to have much of an
-effect on $k_\text{eff}$, which is primarily driven by the prediction of
+method for handling FPY energy dependence does not appear to have much
+effect on $k_\text{eff}$, which is driven primarily by the prediction of
 actinide concentrations.
@@
 \Cref{fig:pwr-actinides-full} shows the relative difference in predicted
-actinide concentrations between OpenMC and Serpent at four different burnup
-values. The fission products that are shown in \cref{fig:pwr-actinides-full}
+actinide concentrations between OpenMC and Serpent at four burnup
+values. The fission products shown in \cref{fig:pwr-actinides-full}
 have been chosen because they are either strong neutron absorbers or significant
@@
 We begin with results for the full depletion chain and interpolated FPYs, which
-most closely matches the method used in Serpent; \cref{fig:pwr-fp-full-average}
+most closely match the method used in Serpent; \cref{fig:pwr-fp-full-average}
 shows the relative difference in fission product concentrations between OpenMC
@@
 curve tend to increase significantly, and the peaks around $A=95$ and $A=135$
-tend to slightly decrease. This corresponds to exactly what is observed in
+tend to slightly decrease. These results correspond to exactly what is observed in
 \cref{fig:pwr-fp-full-thermal}. Because thermal FPYs were used in OpenMC for
 this case, we expect that fission products near one of the two peaks of the FPY
-curve will be slightly overpredicted relative to Serpent, and fission products
-away from the peak will be underpredicted. This can be confirmed for particular
+curve will be slightly overpredicted relative to Serpent and that fission products
+away from the peak will be underpredicted. These results can be confirmed for particular
 nuclides. Take, for example, $^{109}$Ag; in the $A=109$ mass chain, the nuclides
@@
 almost all fission products are predicted to well within 1\% at 50 MWd/kg.
-However, in this case there is one clear outlier: $^{85}$Kr. Upon further
-investigation, it was determined that the root cause for this difference is the
-fact that the CASL depletion chain uses a cumulative yield for $^{85}$Kr, and
+In this case, however, we find one clear outlier: $^{85}$Kr. Upon further
+investigation, we determined that the root cause for this difference is 
+that the CASL depletion chain uses a cumulative yield for $^{85}$Kr, and
 the cumulative yield of $^{85}$Kr is known to be inconsistent with the
@@
 fission product concentrations when using the CASL depletion chain along with
-thermal FPYs. The results are quite close to those shown
+thermal FPYs. The results are close to those shown
 \cref{fig:pwr-fp-full-thermal}, with the exception of $^{85}$Kr for the
@@
 The SFR assembly model is based on an OECD/NEA benchmark for SFR
-analysis~\citep{nsc2015}. Namely, a 2D model of a driver assembly for the 1000
+analysis~\citep{nsc2015}. Specifically, a 2D model of a driver assembly for the 1000
 MWth metallic core (MET-1000) was used. The fuel composition corresponded to the
-lowest axial region (the 17.16 cm column in Table 2.23 from ~\citep{nsc2015}).
+lowest axial region (the 17.16 cm column in Table 2.23 from~\citep{nsc2015}).
 All dimensions and material compositions can be found in the original OECD/NEA
@@
 pin~\citep{cahalan2007anl} was used for power normalization. The assembly was
-depleted for 360 days in 20 day timesteps. As in the PWR pincell case, the
+depleted for 360 days in 20-day timesteps. As in the PWR pincell case, the
 predictor method was used in both codes for time integration.
 
-Unlike the PWR pincell model, which only contains materials at a single
+Unlike the PWR pincell model, which contains materials only at a single
 temperature, the SFR assembly model uses a higher temperature in the fuel than
-in other materials. In OpenMC, the temperature-dependence of neutron cross
+in other materials. In OpenMC, the temperature dependence of neutron cross
 sections was handled by interpolating data between neighboring temperatures.
 Serpent uses an explicit temperature treatment known as target motion
-sampling~\citep{viitanen2012nse}. However, we note that for this problem, the
+sampling~\citep{viitanen2012nse}. We note, however, that for this problem the
 temperature dependence of unresolved resonance probability tables becomes
@@
 included a case for the SFR assembly using the reduced depletion chain. Two
-coupled transport--depletion simulation with Serpent was carried out, one with
+coupled transport--depletion simulation with Serpent was carried out: one with
 probability tables turned on and one with them turned off. For all simulations,
@@
 results using the CASL chain. For each chain, simulations with probability
-tables on are denoted by circles and simulations without probability tables are
-denoted by squares. Unlike in the PWR pincell case, there is a statistically
+tables on are denoted by circles, and simulations without probability tables are
+denoted by squares. Unlike in the PWR pincell case, we see a statistically
 significant difference in $k_\text{eff}$ at zero days. Let us first consider the
@@
 increasing positive bias between OpenMC and Serpent. To understand the cause of
-this difference, it is instructive to look at differences in isotopic
+this difference, let us look at differences in isotopic
 compositions.
@@
 
-When probability tables are turned off, very close agreement is observed between
+When probability tables are turned off, close agreement is observed between
 the predicted actinide concentrations in OpenMC and Serpent, as shown in
@@
 Most differences between OpenMC and Serpent are below 1\%. As with the PWR
-pincell model, OpenMC tends to produce a lower concentration of $^{109}$Ag most
+pincell model, OpenMC tends to produce a lower concentration of $^{109}$Ag, most
 likely because of differences in the treatment of FPY energy dependence. The
@@
 on a single time integration method, OpenMC implements several time integration
-schemes including the predictor method, standard predictor--corrector methods,
-stochastic implicit Euler integrators, and several novel
+schemes, including the predictor method, standard predictor--corrector methods,
+and stochastic implicit Euler integrators.
 methods~\citep{josey2017phd, josey2017jcp}. To evaluate the matrix exponential,
@@
 
-To assess the accuracy of the new depletion solver, comparisons were made
-between OpenMC and Serpent on a PWR pincell model and an SFR assembly model.
+To assess the accuracy of the new depletion solver, we compared 
+OpenMC and Serpent on a PWR pincell model and an SFR assembly model.
 When the full depletion chain was used in OpenMC, the predictions of
 $k_\text{eff}$ agreed to within 20--30 pcm of that calculated by Serpent at all
-times in life for both the PWR and SFR model. Actinide concentrations were found
+times for both the PWR and SFR models. Actinide concentrations were found
 to agree to a fraction of a percent, and most fission product concentrations
@@
 that there is no consensus in the community as to what is the ``correct'' way of
-handling this energy dependence, so it is not possible to opine on whether one
+handling this energy dependence, so it is not possible to opine about whether one
 set of results is right or wrong. When the simplified CASL depletion chain was
@@
 yields of precursors~\citep{pigni2015nds}. Because the CASL depletion chain
-results in a significantly lower runtime compared to the full depletion chain
+results in a significantly lower runtime compared with the full depletion chain
 while still retaining accuracy, it is recommended for general use by OpenMC
@@
 this effect. Although this is not a huge difference, users who are simulating
-problems with a non-uniform temperature distribution should be aware of this
+problems with a nonuniform temperature distribution should be aware of this
 behavior. This difference could be mitigated by using nuclear data processed on
@@
 
-Finally, we remark that although we did observe some differences between OpenMC
+Although we did observe some differences between OpenMC
 and Serpent, most of the differences were related to deficiencies in the nuclear
-data itself. As a whole, the results demonstrate that when the same nuclear data
-and same methods are used in two independently developed codes, it is possible
-to obtain results in a depletion simulation that match within 10s of pcm on
+data. As a whole, the results demonstrate that when the same nuclear data
+and same methods are used in two independently developed codes, one can 
+obtain results in a depletion simulation that match within 10s of pcm on
 $k_\text{eff}$ and less than 1\% on isotopic concentrations.
 
-In the present work, we have only performed a cross-code comparison between
-OpenMC and Serpent. While the results are certainly encouraging, we would
+In the present work, we have performed a cross-code comparison between
+OpenMC and Serpent. While the results are certainly encouraging, we 
 caution readers that no comparisons against experimental results have been
-performed to date and would be needed to ensure the fitness for a particular
+performed to date and would be needed in order to ensure the fitness for a particular
 application. We also highlight the fact that the present comparisons are focused
@@
 We gratefully acknowledge the OpenMC user community for providing early feedback
-and bug reports related to depletion that have helped improve the code. We wish
-to acknowledge Jaakko Lepp{\"{a}}nen (VTT) for providing the developmental
-version of Serpent 2.1.32 used in this work and for helping to shed light on the
+and bug reports related to depletion that have helped improve the code. We 
+acknowledge Jaakko Lepp{\"{a}}nen (VTT) for providing the developmental
+version of Serpent 2.1.32 used in this work and for helping shed light on the
 methods used in Serpent. We acknowledge Nicolas Stauff (ANL) for providing
@@
 collaborative effort of the U.S. Department of Energy Office of Science and the
-National Nuclear Security Administration. The submitted manuscript has been
-created by UChicago Argonne, LLC, operator of Argonne National Laboratory under
-contract DE-AC02-06CH11357.
-
+National Nuclear Security Administration. The submitted manuscript
+was also suported in part by the U.S. Department of Energy, Office of Science, under contrct DE-AC02-06CH11357.
 We gratefully acknowledge the computing resources provided on Bebop, a

### 2020-07-15 manuscript.tex
Move lines around (no actual changes)

@@
   cost, accuracy, and memory use. For all time integration methods, evaluation
-  of the matrix exponential is performed by using the incomplete partial fraction
-  form of the Chebyshev rational approximation method.
-
-  Simulations of a pressurized water reactor (PWR) pincell and a sodium-cooled fast reactor (SFR) assembly were carried out with OpenMC
-  and Serpent. For both problems, the use of a high-fidelity depletion chain
-  results in predictions of $k_\text{eff}$ that agree within 20--30 pcm between
-  OpenMC and Serpent. Predicted actinide concentrations were found to agree to a
-  fraction of a percent, and most fission product concentrations were found to
-  agree within 1\%. The few cases where larger differences were observed can be
-  attributed either to differences in how the energy dependence of fission
-  product yields is handled or deficiencies in the nuclear data used. OpenMC
-  simulations of the PWR and SFR problems using a simplified 228-nuclide
-  depletion chain demonstrate that it achieves accuracy close to that of the
-  full, high-fidelity depletion chain.
+  of the matrix exponential is performed by using the incomplete partial
+  fraction form of the Chebyshev rational approximation method.
+
+  Simulations of a pressurized water reactor (PWR) pincell and a sodium-cooled
+  fast reactor (SFR) assembly were carried out with OpenMC and Serpent. For both
+  problems, the use of a high-fidelity depletion chain results in predictions of
+  $k_\text{eff}$ that agree within 20--30 pcm between OpenMC and Serpent.
+  Predicted actinide concentrations were found to agree to a fraction of a
+  percent, and most fission product concentrations were found to agree within
+  1\%. The few cases where larger differences were observed can be attributed
+  either to differences in how the energy dependence of fission product yields
+  is handled or deficiencies in the nuclear data used. OpenMC simulations of the
+  PWR and SFR problems using a simplified 228-nuclide depletion chain
+  demonstrate that it achieves accuracy close to that of the full, high-fidelity
+  depletion chain.
 \end{abstract}
@@
 When a material in a system is subject to irradiation over a long period of
-time, nuclides within the material will transmute because of nuclear reactions and
-spontaneous radioactive decay. The time-dependent process by which nuclides
+time, nuclides within the material will transmute because of nuclear reactions
+and spontaneous radioactive decay. The time-dependent process by which nuclides
 transmute under irradiation is known as \emph{depletion}, \emph{burnup}, or
@@
 on the density of other nuclides, $\{N_j \,|\, j \ne i\}$, \cref{eq:depletion}
-represents a system of first-order ordinary differential equations (ODEs). In order to
-form a proper initial value problem, the nuclide densities at time 0 must be
-specified:
+represents a system of first-order ordinary differential equations (ODEs). In
+order to form a proper initial value problem, the nuclide densities at time 0
+must be specified:
 \begin{equation}
@@
 state equation. This implies that the only ``input'' that is required by the
-transport solver is $\vect{n}$, as opposed to $\vect{n}$ and $t$. Without loss of
-generality, let us write the burnup matrix solely as a function of $\vect{n}$ so
-that \cref{eq:depletion-matrix-t} becomes
+transport solver is $\vect{n}$, as opposed to $\vect{n}$ and $t$. Without loss
+of generality, let us write the burnup matrix solely as a function of $\vect{n}$
+so that \cref{eq:depletion-matrix-t} becomes
 \begin{equation}
@@
 methods available that the user can choose. In the following sections, we
-outline these methods in OpenMC, several of which have not been
-fully described in the literature. An exhaustive analysis of these time
-integration methods and their merits can be found in the thesis of
-\citet{josey2017phd}.
+outline these methods in OpenMC, several of which have not been fully described
+in the literature. An exhaustive analysis of these time integration methods and
+their merits can be found in the thesis of \citet{josey2017phd}.
 
@@
 composition over half the step. Then, $\vect{A}$ is reevaluated by using the
-predicted middle-of-step composition and is used to deplete over the full step. The
-algorithm can be expressed as
+predicted middle-of-step composition and is used to deplete over the full step.
+The algorithm can be expressed as
 \begin{equation}
@@
 integrator based on the LE/QI (linear extrapolation, quadratic interpolation)
-method~\citep{isotalo2011ane2}, where the integration of \cref{eq:linear-ode} (in
-which case $\hat{\vect{A}}(t)$ is represented by a quadratic polynomial in $t$)
-is performed by using the fourth-order commutator-free Magnus integrator based on
-\cref{eq:compose4}.
+method~\citep{isotalo2011ane2}, where the integration of \cref{eq:linear-ode}
+(in which case $\hat{\vect{A}}(t)$ is represented by a quadratic polynomial in
+$t$) is performed by using the fourth-order commutator-free Magnus integrator
+based on \cref{eq:compose4}.
 
@@
 All of the integration methods discussed thus far are explicit methods that are
-only conditionally stable. In many well-documented cases 
-oscillatory behavior and diverging solutions have been observed. To address
-this situation, \citet{dufek2013ane} proposed an implicit-Euler-like method that is
+only conditionally stable. In many well-documented cases oscillatory behavior
+and diverging solutions have been observed. To address this situation,
+\citet{dufek2013ane} proposed an implicit-Euler-like method that is
 unconditionally stable. The main idea is to replace $\vect{A}(\vect{n}_i)$ with
@@
 between the accuracy of the method, the computational expense, and the memory
-requirements. When an MC code is used, the expense is dominated by the time
-to compute the transport solution, in other words, to evaluate $\vect{A}$ for a given
-$\vect{n}$. Thus, the cost of a method scales with the number of $\vect{A}$
-evaluations that are performed per timestep. On the other hand, methods that
-require more evaluations generally achieve higher accuracy. The predictor method
-requires only one evaluation, and its error converges as $\mathcal{O}(h)$. The
-CE/CM method requires two evaluations and is thus twice as expensive as the
-predictor method, but achieves an error of $\mathcal{O}(h^2)$.
+requirements. When an MC code is used, the expense is dominated by the time to
+compute the transport solution, in other words, to evaluate $\vect{A}$ for a
+given $\vect{n}$. Thus, the cost of a method scales with the number of
+$\vect{A}$ evaluations that are performed per timestep. On the other hand,
+methods that require more evaluations generally achieve higher accuracy. The
+predictor method requires only one evaluation, and its error converges as
+$\mathcal{O}(h)$. The CE/CM method requires two evaluations and is thus twice as
+expensive as the predictor method, but achieves an error of $\mathcal{O}(h^2)$.
 
@@
 systems. However, constructing the burnup matrix, $\vect{A}$, involves not only
-solving the transport equation to estimate nuclear reaction rates but also making a
-series of choices about what data to include.
+solving the transport equation to estimate nuclear reaction rates but also
+making a series of choices about what data to include.
 
@@
 \begin{itemize}
-  \item What transmutation reactions are possible, their Q values, and their products
+  \item What transmutation reactions are possible, their Q values, and their
+  products
   \item If a nuclide is not stable, what decay modes are possible, their
@@
 Depletion chains can be automatically created in the Python application
-programming interface in the following manner. First, the user supplies a
-list of ENDF incident neutron, decay, and FPY sublibrary files. The incident
-neutron files are used to determine what nuclear reactions result in
-transmutation and what the corresponding Q values of the reactions are. The
-decay files are used to determine the half-life and possible decay modes of
-nuclides that are unstable. The FPY sublibraries are used to determine what
-fission products are produced when a fission reaction occurs. This information
-is then aggregated and written to an XML file that serves as the depletion chain
-file that is used at runtime to construct a burnup matrix.
+programming interface in the following manner. First, the user supplies a list
+of ENDF incident neutron, decay, and FPY sublibrary files. The incident neutron
+files are used to determine what nuclear reactions result in transmutation and
+what the corresponding Q values of the reactions are. The decay files are used
+to determine the half-life and possible decay modes of nuclides that are
+unstable. The FPY sublibraries are used to determine what fission products are
+produced when a fission reaction occurs. This information is then aggregated and
+written to an XML file that serves as the depletion chain file that is used at
+runtime to construct a burnup matrix.
 
@@
 nuclide will undergo a $\upbeta^-$ decay. Once the replacement nuclide has been
-determined, OpenMC will check again whether it exists in the decay sublibrary files.
-If not, the process is repeated.
+determined, OpenMC will check again whether it exists in the decay sublibrary
+files. If not, the process is repeated.
 
@@
 nuclides will result in an $n\times n$ burnup matrix, there is an incentive to
-reduce the number of nuclides in the depletion chain in order to minimize both the size
-of the burnup matrix and the number of nuclides that are tracked during a
-transport solve. The depletion chain class in OpenMC's Python API includes a
+reduce the number of nuclides in the depletion chain in order to minimize both
+the size of the burnup matrix and the number of nuclides that are tracked during
+a transport solve. The depletion chain class in OpenMC's Python API includes a
 method for automatically reducing the size of the chain. Starting with a set of
 nuclides that must be included, the method will recursively go through all decay
-and reaction modes to determine what products may be produced. This process can help
-minimize both the size of the depletion matrix and the number of nuclides for
-which reaction rates must be tallied, thereby reducing the time spent in the
+and reaction modes to determine what products may be produced. This process can
+help minimize both the size of the depletion matrix and the number of nuclides
+for which reaction rates must be tallied, thereby reducing the time spent in the
 depletion and transport solvers. Using this method on the ENDF/B-VII.1 library
@@
 
-OpenMC will set up tallies in a problem based on what transmutation reactions are
-available in a depletion chain file, so any arbitrary number of transmutation
-reactions can be tracked. By default, the automatic depletion chain generation
-described in \cref{sec:depletion-chain} will include the following transmutation
-reactions: fission, (n,$\gamma$), (n,2n), (n,3n), (n,4n), (n,p), and
-(n,$\alpha$). While this list is suitable for fission applications, fusion
-applications may require a wider set of transmutation reactions to be tracked.
-Currently, reaction rates are determined by directly tallying the reaction rates
-using a tracklength estimator.
+OpenMC will set up tallies in a problem based on what transmutation reactions
+are available in a depletion chain file, so any arbitrary number of
+transmutation reactions can be tracked. By default, the automatic depletion
+chain generation described in \cref{sec:depletion-chain} will include the
+following transmutation reactions: fission, (n,$\gamma$), (n,2n), (n,3n),
+(n,4n), (n,p), and (n,$\alpha$). While this list is suitable for fission
+applications, fusion applications may require a wider set of transmutation
+reactions to be tracked. Currently, reaction rates are determined by directly
+tallying the reaction rates using a tracklength estimator.
 
@@
 $^{241}$Am to accurately predict the production of higher
-actinides~\citep{haeck2012nse}. Doing so is complicated, however, by the fact that the
-branching ratio may depend on the incident neutron energy causing capture (hence
-why $f_{j\rightarrow i}$ in \cref{eq:depletion} depends on $E$).
-
-OpenMC does not currently implement a method for handling energy-dependent capture
-branching ratios directly and treats $f_{j\rightarrow i}$ as a constant.
+actinides~\citep{haeck2012nse}. Doing so is complicated, however, by the fact
+that the branching ratio may depend on the incident neutron energy causing
+capture (hence why $f_{j\rightarrow i}$ in \cref{eq:depletion} depends on $E$).
+
+OpenMC does not currently implement a method for handling energy-dependent
+capture branching ratios directly and treats $f_{j\rightarrow i}$ as a constant.
 However, the depletion chain file does allow a reaction to be listed multiple
@@
 Spectrum-averaged capture branching ratios have been computed in LWR and SFR
-spectra and are available at \url{https://openmc.org/depletion-chains}. In
-the Python API, the class representing a depletion chain allows branching ratios
-to be easily retrieved and changed. We note that while this treatment does not
+spectra and are available at \url{https://openmc.org/depletion-chains}. In the
+Python API, the class representing a depletion chain allows branching ratios to
+be easily retrieved and changed. We note that while this treatment does not
 accurately account for the change in the capture branching ratio with increasing
@@
 contains a continuum of energies. The ENDF-6 formats manual~\citep{trkov2018bnl}
-recommends using linear-linear interpolation between energy points. At
-the end of a transport solve, however, we typically have a one-group or multigroup
-fission reaction rate---and it is not clear how this information should be used  to
-calculate an effective FPY.
+recommends using linear-linear interpolation between energy points. At the end
+of a transport solve, however, we typically have a one-group or multigroup
+fission reaction rate---and it is not clear how this information should be used
+to calculate an effective FPY.
 
@@
 The reaction rates calculated by OpenMC are given in units of reactions per
-source particle. For depletion, one must compute an absolute reaction
-rate in reactions per second. To do so, the total observed heating in the system
-is determined, and then the ratio of the system power to the observed heating is
+source particle. For depletion, one must compute an absolute reaction rate in
+reactions per second. To do so, the total observed heating in the system is
+determined, and then the ratio of the system power to the observed heating is
 used as a normalization factor on the reaction rates. OpenMC allows the total
@@
 are applied. In the second method, the actual heating rate is tallied over the
-entire system (separate from the tally for transmutation reaction rates). With this
-method one can account for the energy dependence of the fission
-energy release data as well as the fact that neutrons and photons may leak out
-of the system, carrying away their energy with them. The latter effect may be
+entire system (separate from the tally for transmutation reaction rates). With
+this method one can account for the energy dependence of the fission energy
+release data as well as the fact that neutrons and photons may leak out of the
+system, carrying away their energy with them. The latter effect may be
 particularly important for small, high-leakage systems.
@@
 that represent the major components of the depletion solver. Each of the
-abstract classes defines an interface, that is, a set of methods that
-a subclass must implement. First, the abstract \texttt{TransportOperator} class
-represents a particle transport solver. Subclasses of \texttt{TransportOperator}
-are required in order to implement a method that takes a list of nuclide densities to be
-used in the problem and a system power and returns the $k$-eigenvalue and the
-reaction rates resulting from the transport operator. Currently, there is a
+abstract classes defines an interface, that is, a set of methods that a subclass
+must implement. First, the abstract \texttt{TransportOperator} class represents
+a particle transport solver. Subclasses of \texttt{TransportOperator} are
+required in order to implement a method that takes a list of nuclide densities
+to be used in the problem and a system power and returns the $k$-eigenvalue and
+the reaction rates resulting from the transport operator. Currently, there is a
 single subclass of \texttt{TransportOperator} that interfaces with the Python
-bindings to OpenMC's C++ API to call a transport solve and obtain reaction rates.
-Once reaction rates are obtained, a separate depletion chain class (that is not
-abstract) uses them to construct the burnup matrix, $\vect{A}$, based on the
-data present in the depletion chain.
+bindings to OpenMC's C++ API to call a transport solve and obtain reaction
+rates. Once reaction rates are obtained, a separate depletion chain class (that
+is not abstract) uses them to construct the burnup matrix, $\vect{A}$, based on
+the data present in the depletion chain.
 
@@
 evaluate \cref{eq:constant-A}). Subclasses of \texttt{DepSystemSolver} are
-required in order to implement a method that takes the burnup matrix $\vect{A}$, an
-initial density vector $\vect{n}_0$ and a timestep size and returns the
+required in order to implement a method that takes the burnup matrix $\vect{A}$,
+an initial density vector $\vect{n}_0$ and a timestep size and returns the
 resulting nuclide densities at the end of the step. The only subclass of
@@
 This class effectively orchestrates the coupled transport--depletion simulation.
-Subclasses of \texttt{Integrator} are required in order to implement a method that takes
-an initial vector of nuclide densities, reaction rates returned from an
-operator, a time interval, and a power and returns the resulting nuclide
+Subclasses of \texttt{Integrator} are required in order to implement a method
+that takes an initial vector of nuclide densities, reaction rates returned from
+an operator, a time interval, and a power and returns the resulting nuclide
 densities at the end of the step. Thus, \texttt{Integrator} utilizes
@@
 using the \texttt{openmc.deplete} module as well as a developmental version of
-Serpent 2.1.32 on two simple problems. The first problem is a 
-PWR pincell, and the second problem is a single assembly of an
-SFR. These problems were selected to cover both
-thermal and fast neutron spectra. While we had originally attempted to use
-Serpent 2.1.31, preliminary comparisons between OpenMC and this version of
-Serpent revealed significant differences in the predicted concentration of
-$^{97}$Mo. This difference was found to be due to $^{97\text{m}}$Y having a
-nonzero independent yield and a zero cumulative yield in the FPY sublibrary
-file for $^{235}$U fission in ENDF/B-VII.1. The zero cumulative yield caused
-Serpent 2.1.31 to ignore it altogether, which resulted in significantly less
-production of $^{97}$Mo. The developmental version of Serpent 2.1.32 no longer
-ignores this particular fission product despite the erroneous cumulative yield.
+Serpent 2.1.32 on two simple problems. The first problem is a PWR pincell, and
+the second problem is a single assembly of an SFR. These problems were selected
+to cover both thermal and fast neutron spectra. While we had originally
+attempted to use Serpent 2.1.31, preliminary comparisons between OpenMC and this
+version of Serpent revealed significant differences in the predicted
+concentration of $^{97}$Mo. This difference was found to be due to
+$^{97\text{m}}$Y having a nonzero independent yield and a zero cumulative yield
+in the FPY sublibrary file for $^{235}$U fission in ENDF/B-VII.1. The zero
+cumulative yield caused Serpent 2.1.31 to ignore it altogether, which resulted
+in significantly less production of $^{97}$Mo. The developmental version of
+Serpent 2.1.32 no longer ignores this particular fission product despite the
+erroneous cumulative yield.
 
@@
 fundamental nuclear data are being used in both codes, one expects to obtain the
-same answer. Thus, even minute differences may be indicative of a true
-problem in the method implementation in one of the codes.
+same answer. Thus, even minute differences may be indicative of a true problem
+in the method implementation in one of the codes.
 
@@
 simulations in both OpenMC and Serpent. Serpent is able to use the ACE files
-from the endf71x library directly, whereas OpenMC uses an HDF5 format that requires
-conversion of the ACE files first. For the PWR pincell problem,
+from the endf71x library directly, whereas OpenMC uses an HDF5 format that
+requires conversion of the ACE files first. For the PWR pincell problem,
 $S(\alpha,\beta)$ thermal scattering data are also needed for hydrogen in light
@@
 the full chain cases rises to about 20 pcm. The reduced chain case exhibits the
-exact same behavior as the full chain does, as one would expect. Use of the CASL chain
-results in a slightly larger difference between OpenMC and Serpent at 50 MWd/kg,
-about 40 pcm. We also see that for both the full and CASL chains, changing the
-method for handling FPY energy dependence does not appear to have much
-effect on $k_\text{eff}$, which is driven primarily by the prediction of
+exact same behavior as the full chain does, as one would expect. Use of the CASL
+chain results in a slightly larger difference between OpenMC and Serpent at 50
+MWd/kg, about 40 pcm. We also see that for both the full and CASL chains,
+changing the method for handling FPY energy dependence does not appear to have
+much effect on $k_\text{eff}$, which is driven primarily by the prediction of
 actinide concentrations.
@@
 \Cref{fig:pwr-actinides-full} shows the relative difference in predicted
-actinide concentrations between OpenMC and Serpent at four burnup
-values. The fission products shown in \cref{fig:pwr-actinides-full}
-have been chosen because they are either strong neutron absorbers or significant
-contributors to decay heat and/or radiotoxicity. We see that most of the
-differences are less than 0.1\%, and all are within 0.15\%. The differences in
-$^{235}$U and $^{239}$Pu concentrations increase slightly with increasing
-burnup, which most likely accounts for the 20 pcm difference observed for the
-full chain in \cref{fig:pwr-keff}. While not shown, the predicted actinide
-concentrations for the reduced chain are nearly identical to those observed for
-the full chain in \cref{fig:pwr-actinides-full}. \Cref{fig:pwr-actinides-casl}
-shows the differences in actinide concentrations using the CASL depletion chain.
-Because the CASL depletion chain differs from the full depletion chain primarily
-with respect to how fission products are treated, the differences reported in
+actinide concentrations between OpenMC and Serpent at four burnup values. The
+fission products shown in \cref{fig:pwr-actinides-full} have been chosen because
+they are either strong neutron absorbers or significant contributors to decay
+heat and/or radiotoxicity. We see that most of the differences are less than
+0.1\%, and all are within 0.15\%. The differences in $^{235}$U and $^{239}$Pu
+concentrations increase slightly with increasing burnup, which most likely
+accounts for the 20 pcm difference observed for the full chain in
+\cref{fig:pwr-keff}. While not shown, the predicted actinide concentrations for
+the reduced chain are nearly identical to those observed for the full chain in
+\cref{fig:pwr-actinides-full}. \Cref{fig:pwr-actinides-casl} shows the
+differences in actinide concentrations using the CASL depletion chain. Because
+the CASL depletion chain differs from the full depletion chain primarily with
+respect to how fission products are treated, the differences reported in
 \cref{fig:pwr-actinides-casl} closely mirror those in
@@
 
-\Cref{fig:pwr-fp-full-thermal} shows the predicted fission product concentrations
-when using the full depletion chain along with thermal FPYs. In this case,
-although most fission products are within 1\% of Serpent, they are
+\Cref{fig:pwr-fp-full-thermal} shows the predicted fission product
+concentrations when using the full depletion chain along with thermal FPYs. In
+this case, although most fission products are within 1\% of Serpent, they are
 systematically overpredicted other than a few nuclides. Because Serpent uses the
@@
 curve tend to increase significantly, and the peaks around $A=95$ and $A=135$
-tend to slightly decrease. These results correspond to exactly what is observed in
-\cref{fig:pwr-fp-full-thermal}. Because thermal FPYs were used in OpenMC for
+tend to slightly decrease. These results correspond to exactly what is observed
+in \cref{fig:pwr-fp-full-thermal}. Because thermal FPYs were used in OpenMC for
 this case, we expect that fission products near one of the two peaks of the FPY
-curve will be slightly overpredicted relative to Serpent and that fission products
-away from the peak will be underpredicted. These results can be confirmed for particular
-nuclides. Take, for example, $^{109}$Ag; in the $A=109$ mass chain, the nuclides
-with the highest independent yields are $^{109}$Tc and $^{109}$Mo. Both these
-nuclides have independent yields that increase significantly with increasing
-neutron energy. The same situation occurs for $^{129}$I, whose predominant
-precursor is $^{129}$Sn. The independent yields for these three precursor
-nuclides are shown in \cref{tab:fp-yields}.
+curve will be slightly overpredicted relative to Serpent and that fission
+products away from the peak will be underpredicted. These results can be
+confirmed for particular nuclides. Take, for example, $^{109}$Ag; in the $A=109$
+mass chain, the nuclides with the highest independent yields are $^{109}$Tc and
+$^{109}$Mo. Both these nuclides have independent yields that increase
+significantly with increasing neutron energy. The same situation occurs for
+$^{129}$I, whose predominant precursor is $^{129}$Sn. The independent yields for
+these three precursor nuclides are shown in \cref{tab:fp-yields}.
 \begin{figure}[H]
@@
 depletion chain and interpolated FPYs. As in \cref{fig:pwr-fp-full-average},
-almost all fission products are predicted to well within 1\% at 50 MWd/kg.
-In this case, however, we find one clear outlier: $^{85}$Kr. Upon further
-investigation, we determined that the root cause for this difference is 
-that the CASL depletion chain uses a cumulative yield for $^{85}$Kr, and
-the cumulative yield of $^{85}$Kr is known to be inconsistent with the
-independent yields of the $A=85$ mass chain in
-ENDF/B-VII.1~\citep{pigni2015nds}. Serpent relies solely on independent yields
-in its depletion solver (as does the full depletion chain). We can conclude,
-then, that there is nothing inherently wrong with the CASL depletion chain or
-OpenMC; it is simply a matter of the underlying nuclear data being inconsistent.
-The 5\% underprediction of $^{85}$Kr is also likely responsible for the slightly
-higher difference in $k_\text{eff}$ for the CASL depletion chain shown in
-\cref{fig:pwr-keff}.
+almost all fission products are predicted to well within 1\% at 50 MWd/kg. In
+this case, however, we find one clear outlier: $^{85}$Kr. Upon further
+investigation, we determined that the root cause for this difference is that the
+CASL depletion chain uses a cumulative yield for $^{85}$Kr, and the cumulative
+yield of $^{85}$Kr is known to be inconsistent with the independent yields of
+the $A=85$ mass chain in ENDF/B-VII.1~\citep{pigni2015nds}. Serpent relies
+solely on independent yields in its depletion solver (as does the full depletion
+chain). We can conclude, then, that there is nothing inherently wrong with the
+CASL depletion chain or OpenMC; it is simply a matter of the underlying nuclear
+data being inconsistent. The 5\% underprediction of $^{85}$Kr is also likely
+responsible for the slightly higher difference in $k_\text{eff}$ for the CASL
+depletion chain shown in \cref{fig:pwr-keff}.
 \begin{figure}[H]
@@
 The SFR assembly model is based on an OECD/NEA benchmark for SFR
-analysis~\citep{nsc2015}. Specifically, a 2D model of a driver assembly for the 1000
-MWth metallic core (MET-1000) was used. The fuel composition corresponded to the
-lowest axial region (the 17.16 cm column in Table 2.23 from~\citep{nsc2015}).
-All dimensions and material compositions can be found in the original OECD/NEA
-report. The assembly model extends to positive/negative infinity in the $z$
-direction and has periodic boundary conditions on the outer hexagonal surfaces.
-A linear heating rate of \SI{23.3}{\kilo\watt\per\meter} per
+analysis~\citep{nsc2015}. Specifically, a 2D model of a driver assembly for the
+1000 MWth metallic core (MET-1000) was used. The fuel composition corresponded
+to the lowest axial region (the 17.16 cm column in Table 2.23
+from~\citep{nsc2015}). All dimensions and material compositions can be found in
+the original OECD/NEA report. The assembly model extends to positive/negative
+infinity in the $z$ direction and has periodic boundary conditions on the outer
+hexagonal surfaces. A linear heating rate of \SI{23.3}{\kilo\watt\per\meter} per
 pin~\citep{cahalan2007anl} was used for power normalization. The assembly was
@@
 
-% Also, we
-% have only included a single case with thermal FPYs to compare with the use of
-% interpolated FPYs using the same depletion chain.
-
 \Cref{fig:sfr-keff} shows the difference between $k_\text{eff}$ calculated by
@@
 increasing positive bias between OpenMC and Serpent. To understand the cause of
-this difference, let us look at differences in isotopic
-compositions.
+this difference, let us look at differences in isotopic compositions.
 \begin{figure}[H]
@@
 
-To assess the accuracy of the new depletion solver, we compared 
-OpenMC and Serpent on a PWR pincell model and an SFR assembly model.
-When the full depletion chain was used in OpenMC, the predictions of
-$k_\text{eff}$ agreed to within 20--30 pcm of that calculated by Serpent at all
-times for both the PWR and SFR models. Actinide concentrations were found
-to agree to a fraction of a percent, and most fission product concentrations
-were found to agree within 1\%. The few cases where fission product
-concentrations differed by a larger amount were found to be related to
-differences in the treatment of the energy dependence of FPYs. It is our sense
-that there is no consensus in the community as to what is the ``correct'' way of
-handling this energy dependence, so it is not possible to opine about whether one
-set of results is right or wrong. When the simplified CASL depletion chain was
-used in OpenMC, the accuracy of actinide and fission product concentrations did
-not change appreciably. The only nuclide for which a significant difference was
-observed was $^{85}$Kr; this difference can be explained by a cumulative yield
-in the ENDF/B-VII.1 FPY sublibrary that is inconsistent with the independent
-yields of precursors~\citep{pigni2015nds}. Because the CASL depletion chain
-results in a significantly lower runtime compared with the full depletion chain
-while still retaining accuracy, it is recommended for general use by OpenMC
-users.
+To assess the accuracy of the new depletion solver, we compared OpenMC and
+Serpent on a PWR pincell model and an SFR assembly model. When the full
+depletion chain was used in OpenMC, the predictions of $k_\text{eff}$ agreed to
+within 20--30 pcm of that calculated by Serpent at all times for both the PWR
+and SFR models. Actinide concentrations were found to agree to a fraction of a
+percent, and most fission product concentrations were found to agree within 1\%.
+The few cases where fission product concentrations differed by a larger amount
+were found to be related to differences in the treatment of the energy
+dependence of FPYs. It is our sense that there is no consensus in the community
+as to what is the ``correct'' way of handling this energy dependence, so it is
+not possible to opine about whether one set of results is right or wrong. When
+the simplified CASL depletion chain was used in OpenMC, the accuracy of actinide
+and fission product concentrations did not change appreciably. The only nuclide
+for which a significant difference was observed was $^{85}$Kr; this difference
+can be explained by a cumulative yield in the ENDF/B-VII.1 FPY sublibrary that
+is inconsistent with the independent yields of precursors~\citep{pigni2015nds}.
+Because the CASL depletion chain results in a significantly lower runtime
+compared with the full depletion chain while still retaining accuracy, it is
+recommended for general use by OpenMC users.
 
@@
 
-Although we did observe some differences between OpenMC
-and Serpent, most of the differences were related to deficiencies in the nuclear
-data. As a whole, the results demonstrate that when the same nuclear data
-and same methods are used in two independently developed codes, one can 
-obtain results in a depletion simulation that match within 10s of pcm on
-$k_\text{eff}$ and less than 1\% on isotopic concentrations.
-
-In the present work, we have performed a cross-code comparison between
-OpenMC and Serpent. While the results are certainly encouraging, we 
-caution readers that no comparisons against experimental results have been
-performed to date and would be needed in order to ensure the fitness for a particular
-application. We also highlight the fact that the present comparisons are focused
-solely on fission reactor applications. Although the \texttt{openmc.deplete}
-module may eventually be of interest to other application areas, further changes
-and enhancements would be needed. For example, for use in fusion applications, a
+Although we did observe some differences between OpenMC and Serpent, most of the
+differences were related to deficiencies in the nuclear data. As a whole, the
+results demonstrate that when the same nuclear data and same methods are used in
+two independently developed codes, one can obtain results in a depletion
+simulation that match within 10s of pcm on $k_\text{eff}$ and less than 1\% on
+isotopic concentrations.
+
+In the present work, we have performed a cross-code comparison between OpenMC
+and Serpent. While the results are certainly encouraging, we caution readers
+that no comparisons against experimental results have been performed to date and
+would be needed in order to ensure the fitness for a particular application. We
+also highlight the fact that the present comparisons are focused solely on
+fission reactor applications. Although the \texttt{openmc.deplete} module may
+eventually be of interest to other application areas, further changes and
+enhancements would be needed. For example, for use in fusion applications, a
 wider variety of transmutation reactions would need to be accounted for, and the
@@
 collaborative effort of the U.S. Department of Energy Office of Science and the
-National Nuclear Security Administration. The submitted manuscript
-was also suported in part by the U.S. Department of Energy, Office of Science, under contrct DE-AC02-06CH11357.
-We gratefully acknowledge the computing resources provided on Bebop, a
-high-performance computing cluster operated by the Laboratory Computing Resource
-Center at Argonne National Laboratory. This research made use of the resources
-of the High Performance Computing Center at Idaho National Laboratory, which is
-supported by the Office of Nuclear Energy of the U.S. Department of Energy and
-the Nuclear Science User Facilities under Contract No. DE-AC07-05ID14517.
+National Nuclear Security Administration. The submitted manuscript was also
+suported in part by the U.S. Department of Energy, Office of Science, under
+contrct DE-AC02-06CH11357. We gratefully acknowledge the computing resources
+provided on Bebop, a high-performance computing cluster operated by the
+Laboratory Computing Resource Center at Argonne National Laboratory. This
+research made use of the resources of the High Performance Computing Center at
+Idaho National Laboratory, which is supported by the Office of Nuclear Energy of
+the U.S. Department of Energy and the Nuclear Science User Facilities under
+Contract No. DE-AC07-05ID14517.
 

### 2020-09-29 manuscript.tex
Fix sentence in conclusions
Fix typos in acknowledgments
Add middle initial for Colin
Address reviewer 1 comments

@@
 
-\author[lanl]{Colin Josey}
+\author[lanl]{Colin J. Josey}
 \ead{cjosey@lanl.gov}
@@
 MCNP6~\citep{goorley2012nt}), Serpent~\citep{leppanen2015ane},
-MC21~\citep{griesheimer2015ane}, and RMC~\citep{wang2015ane}.
+MC21~\citep{griesheimer2015ane}, RMC~\citep{wang2015ane}, and
+Shift~\citep{davidson2016ane}.
 
@@
   \begin{split}
-  \frac{dN_i(t)}{dt} = &\sum\limits_j \left [ \int_0^\infty dE \;
+  \frac{dN_i(t)}{dt} = &\sum\limits_{j=1}^n \left [ \int_0^\infty dE \;
   f_{j \rightarrow i}(E) \sigma_j (E, t) \phi(E,t) + \lambda_{j\rightarrow i}
   \right ] N_j(t) \\ &- \left [\int_0^\infty dE \; \sigma_i (E,t) \phi(E,t) +
-  \sum\limits_j \lambda_{i\rightarrow j} \right ] N_i(t), \quad i=1,\dots,n ,
+  \sum\limits_{j=1}^n \lambda_{i\rightarrow j} \right ] N_i(t), \quad i=1,\dots,n ,
   \end{split}
@@
 \end{equation*}
-Note that we have not included the spatial dependence of the flux or cross
+By definition, $f_{i\rightarrow i}$ and $\lambda_{i\rightarrow i}$ are zero so
+the summations in \cref{eq:depletion} are effectively over all nuclides except
+$i$. Note that we have not included the spatial dependence of the flux or cross
 sections. For \cref{eq:depletion} to be accurate, the spatial variation of the
@@
 
-\subsubsection{Predictor Method}
+\subsubsection{Constant Extrapolation Method}
 
 The simplest numerical method for solving \cref{eq:depletion-matrix} over a
-timestep is known as the \emph{predictor} method. It assumes that
-$\vect{A}$ is constant over the timestep and assumes the value at the
-beginning of the step. The solution then follows directly from
-\cref{eq:constant-A}:
-\begin{equation}
-  \label{eq:predictor}
+timestep is known as the \emph{constant extrapolation} (CE) method\footnote{In
+the literature, this algorithm is also sometimes called the fully-explicit
+method or the predictor method.}. It assumes that $\vect{A}$ is constant over
+the timestep and assumes the value at the beginning of the step. The solution
+then follows directly from \cref{eq:constant-A}:
+\begin{equation}
+  \label{eq:constant_extrapolation}
   \vect{n}_{i+1} = \exp\left(h\vect{A}(\vect{n}_i) \right) \vect{n}_i.
@@
 A series of \emph{predictor--corrector} methods that use multiple stages offer
-improved accuracy over the predictor method. The simplest of these methods is
-the CE/CM (constant extrapolation, constant midpoint) algorithm, which is the
-default algorithm used in MCNP6~\citep{fensin2006tans}. In this method,
-$\vect{A}$ is evaluated at the beginning of the step and used to deplete the
-composition over half the step. Then, $\vect{A}$ is reevaluated by using the
+improved accuracy over the CE method. The simplest of these
+methods is the CE/CM (constant extrapolation, constant midpoint) algorithm,
+which is the default algorithm used in MCNP6~\citep{fensin2006tans}. In this
+method, $\vect{A}$ is evaluated at the beginning of the step and used to deplete
+the composition over half the step. Then, $\vect{A}$ is reevaluated by using the
 predicted middle-of-step composition and is used to deplete over the full step.
@@
 This method requires two transport solutions and two matrix exponential solves
-per timestep (twice as expensive as the predictor method) and is second-order
+per timestep (twice as expensive as the CE method) and is second-order
 accurate.
@@
 unconditionally stable. The main idea is to replace $\vect{A}(\vect{n}_i)$ with
-$\vect{A}(\vect{n}_{i+1})$ in \cref{eq:predictor}:
+$\vect{A}(\vect{n}_{i+1})$ in \cref{eq:constant_extrapolation}:
 \begin{equation}
@@
 The major downside of an algorithm based on \cref{eq:implicit} is that, like the
-predictor method, it is only first-order accurate. Implicit methods have also
-been applied to higher-order
+CE method, it is only first-order accurate. Implicit methods have also been
+applied to higher-order
 algorithms~\citep{kotlyar2014ane,kotlyar2016ane,cosgrove2020ane}, thereby
@@
 $\vect{A}$ evaluations that are performed per timestep. On the other hand,
-methods that require more evaluations generally achieve higher accuracy. The
-predictor method requires only one evaluation, and its error converges as
+methods that require more evaluations generally achieve higher accuracy. The CE
+method requires only one evaluation, and its error converges as
 $\mathcal{O}(h)$. The CE/CM method requires two evaluations and is thus twice as
-expensive as the predictor method, but achieves an error of $\mathcal{O}(h^2)$.
+expensive as the CE method, but achieves an error of $\mathcal{O}(h^2)$.
 
@@
 The short timesteps at the beginning are used to accurately capture the buildup
-of xenon isotopes. In both codes, the predictor method was used for time
-integration. While the predictor method is the least accurate time integration
+of xenon isotopes. In both codes, the CE method was used for time
+integration. While the CE method is the least accurate time integration
 method, the purpose of the comparison is to confirm that OpenMC and Serpent
@@
 \Cref{fig:pwr-actinides-full} shows the relative difference in predicted
-actinide concentrations between OpenMC and Serpent at four burnup values. The
-fission products shown in \cref{fig:pwr-actinides-full} have been chosen because
-they are either strong neutron absorbers or significant contributors to decay
-heat and/or radiotoxicity. We see that most of the differences are less than
+actinide concentrations between OpenMC and Serpent at four burnup values using
+the full depletion chain. We see that most of the differences are less than
 0.1\%, and all are within 0.15\%. The differences in $^{235}$U and $^{239}$Pu
@@
 shows the relative difference in fission product concentrations between OpenMC
-and Serpent for this case. Most of the differences are less than 0.2\%, and only
-a few fission products show differences up to 0.5\%.
+and Serpent for this case. The fission products shown in
+\cref{fig:pwr-fp-full-average} have been chosen because they are either strong
+neutron absorbers or significant contributors to decay heat and/or
+radiotoxicity. Most of the differences are less than 0.2\%, and only a few
+fission products show differences up to 0.5\%.
 \begin{figure}[H]
@@
 pin~\citep{cahalan2007anl} was used for power normalization. The assembly was
-depleted for 360 days in 20-day timesteps. As in the PWR pincell case, the
-predictor method was used in both codes for time integration.
+depleted for 360 days in 20-day timesteps. As in the PWR pincell case, the CE
+method was used in both codes for time integration.
 
@@
 on a single time integration method, OpenMC implements several time integration
-schemes, including the predictor method, standard predictor--corrector methods,
-and stochastic implicit Euler integrators.
-methods~\citep{josey2017phd, josey2017jcp}. To evaluate the matrix exponential,
-OpenMC relies on the incomplete partial fraction form of CRAM by default.
+schemes, including the constant extrapolation method, standard
+predictor--corrector methods, and stochastic implicit Euler integrators. To
+evaluate the matrix exponential, OpenMC relies on the incomplete partial
+fraction form of CRAM by default.
 
@@
 National Nuclear Security Administration. The submitted manuscript was also
-suported in part by the U.S. Department of Energy, Office of Science, under
-contrct DE-AC02-06CH11357. We gratefully acknowledge the computing resources
-provided on Bebop, a high-performance computing cluster operated by the
-Laboratory Computing Resource Center at Argonne National Laboratory. This
+supported in part by the U.S. Department of Energy, Office of Science, under
+contract No. \mbox{DE-AC02-06CH11357}. We gratefully acknowledge the computing
+resources provided on Bebop, a high-performance computing cluster operated by
+the Laboratory Computing Resource Center at Argonne National Laboratory. This
 research made use of the resources of the High Performance Computing Center at
@@
 the U.S. Department of Energy and the Nuclear Science User Facilities under
-Contract No. DE-AC07-05ID14517.
+Contract No. \mbox{DE-AC07-05ID14517}.
 

### 2020-09-30 manuscript.tex
Address reviewer 2 comments

@@
   demonstrate that it achieves accuracy close to that of the full, high-fidelity
-  depletion chain.
+  depletion chain with respect to the studied responses.
 \end{abstract}
@@
 \cref{sec:results}, we present results for two problems that demonstrate the
-accuracy of OpenMC when compared with Serpent~\citep{leppanen2015ane}. In
-\cref{sec:conclusions}, we summarize the findings, discuss limitations, and
-provide commentary on future avenues of inquiry that may be promising.
+accuracy of OpenMC when compared with Serpent. In \cref{sec:conclusions}, we
+summarize the findings, discuss limitations, and provide commentary on future
+avenues of inquiry that may be promising.
 
@@
 the summations in \cref{eq:depletion} are effectively over all nuclides except
-$i$. Note that we have not included the spatial dependence of the flux or cross
+$i$. Also note that the transmutation cross section $\sigma_i$ represents the
+sum of the cross sections for all reaction channels that result in
+transmutation. We have not included the spatial dependence of the flux or cross
 sections. For \cref{eq:depletion} to be accurate, the spatial variation of the
@@
 \begin{equation}
-  \bar{E} = \frac{\int_0^\infty dE \; E \sigma_f(E)\phi(E)}{\int_0^\infty \sigma_f(E) \phi(E)},
+  \bar{E} = \frac{\int_0^\infty dE \; E \sigma_f(E)\phi(E)}{\int_0^\infty dE \; \sigma_f(E) \phi(E)},
 \end{equation}
@@
 erroneous cumulative yield.
+
+Although both OpenMC and Serpent use CRAM for matrix exponential solutions, they
+employ different forms of CRAM. OpenMC relies on the IPF form of CRAM whereas
+Serpent relies on the partial fraction decomposition (PFD)
+form~\citep{pusa2016nse}. Although the IPF form has been shown to be superior
+for long timesteps in decay calculations, both the PFD and IPF form are very
+accurate for burnup calculations, and we do not expect differences between the
+OpenMC and Serpent results to be attributable to the form of CRAM used. At the
+time of writing, OpenMC uses IPF CRAM of order 48 by default and Serpent uses
+PFD CRAM of order 14.
 
@@
 both the full and CASL depletion chains, the relative differences are all within
-0.2\%. While this may seem close, the observed difference in $^{238}$U and
+1\%. While this may seem close, the observed difference in $^{238}$U and
 $^{239}$Pu concentrations merits special mention. In this model, the fuel
@@
 probability tables at the exact fuel temperature, it statistically interpolates
-between the probability tables at \SI{600}{\kelvin} and \SI{900}{\kelvin}.
+between the probability tables at \SI{600}{\kelvin} and \SI{900}{\kelvin}. Also
+note that $^{239}$U and $^{239}$Np, which are also produced as a result of
+neutron capture in $^{238}$U, have a higher relative difference because of their
+short half-lives and correspondingly small concentrations.
 \begin{figure}[H]
@@
 \cref{fig:sfr-actinides-nop}. In this case, the actinide concentrations agree to
-within 0.01\% at all times during depletion for both the full and CASL depletion
+within 0.04\% at all times during depletion for both the full and CASL depletion
 chains.
@@
 
-While the focus in this work has primarily been on correctness and accuracy, the
-performance of a coupled transport--depletion simulation is also an important
-consideration that directly impacts usability. The need to determine reaction
-rates for hundreds of nuclides leads to a significant performance degredation
-during a transport simulation. Research on methods to reduce the computational
-expense of obtaining transmutation reaction rates~\citep{salcedoperez2019mc}
-show promise but have yet to be incorporated in a release of OpenMC.
+While the focus in this work has primarily been on correctness, the performance
+of a coupled transport--depletion simulation is also an important consideration
+that directly impacts usability. The need to determine reaction rates for
+hundreds of nuclides leads to a significant performance degredation during a
+transport simulation. Research on methods to reduce the computational expense of
+obtaining transmutation reaction rates~\citep{salcedoperez2019mc} show promise
+but have yet to be incorporated in a release of OpenMC.
 

### 2020-09-30 figures/plot_barcharts.py

@@ -47,7 +47,7 @@ day = 24*60*60
 colors = rcParams['axes.prop_cycle'].by_key()['color']
 
 actinides = [
-    'U234', 'U235', 'U236', 'U238',
+    'U234', 'U235', 'U236', 'U238', 'U239', 'Np239',
     'Pu238', 'Pu239', 'Pu240', 'Pu241', 'Pu242',
     'Am241', 'Am242', 'Am242_m1', 'Am243', 'Am244',
     'Cm242', 'Cm243', 'Cm244', 'Cm245', 'Cm246',
@@ -125,10 +125,10 @@ isotope_bar_plot(results_full_average, fuel, actinides, f'pwr_actinides_full.{ex
 isotope_bar_plot(results_reduced, fuel, actinides, f'pwr_actinides_reduced.{ext}')
 isotope_bar_plot(results_casl_average, fuel, actinides, f'pwr_actinides_casl.{ext}')
 
-isotope_bar_plot(results_sfr_full, fuel_sfr, actinides, f'sfr_actinides_full.{ext}')
-isotope_bar_plot(results_sfr_casl, fuel_sfr, actinides, f'sfr_actinides_casl.{ext}')
-isotope_bar_plot(results_sfr_full_nop, fuel_sfr_nop, actinides, f'sfr_actinides_full_nop.{ext}', 3)
-isotope_bar_plot(results_sfr_casl_nop, fuel_sfr_nop, actinides, f'sfr_actinides_casl_nop.{ext}', 4)
+isotope_bar_plot(results_sfr_full, fuel_sfr, actinides, f'sfr_actinides_full.{ext}', 1)
+isotope_bar_plot(results_sfr_casl, fuel_sfr, actinides, f'sfr_actinides_casl.{ext}', 1)
+isotope_bar_plot(results_sfr_full_nop, fuel_sfr_nop, actinides, f'sfr_actinides_full_nop.{ext}')
+isotope_bar_plot(results_sfr_casl_nop, fuel_sfr_nop, actinides, f'sfr_actinides_casl_nop.{ext}')
 
 kwargs = {'figsize': (6.4, 10.0)}
 isotope_bar_plot(results_full_thermal, fuel, fission_products, f'pwr_fp_full_thermal.{ext}', 0, **kwargs)
